---
title: "Chong LHC MR"
author: "Brian Alvarez"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

knitr::opts_knit$set(root.dir = 'C:/Users/brian/OneDrive/Documents/GitHub/Casual_Estimate_mtDNAcn_on_Alzheimer')

 
```


## Libraries

Loads the appropriate packages for MR analysis

``` {r eval=FALSE}
library(tidyverse) # Data wrangling 

library (usethis)

library(devtools)

library(ggman)

library(TwoSampleMR) # MR 

library(R.utils) #To read in Gz data? 
#For the read_outcome_data step

library(RadialMR) # Radial MR sensitivity analysis 

library(phenoscanner)

library(LDlinkR) # LD and proxy snps

library(lhcMR)

library(janitor)
```


## Importing and formatting Exposure

This step imports our GWAS Exposure data sets and formats them up for MR analysis


```{r eval=FALSE}
# Define column types for summary statistics GWAS
coltypes = cols(
  ID = col_character(),
  CHROM = col_double(),
  POS = col_double(),
  REF = col_character(),
  ALT = col_character(),
  AF = col_double(),
  TRAIT = col_character(),
  BETA = col_double(),
  SE = col_double(),
  Z = col_double(),
  P = col_double(),
  N = col_double(),
  OR = col_double(),
  OR_L95 = col_double(),
  OR_U95 = col_double(),
  DIR = col_character(),
  G1000_ID = col_character(),
  G1000_VARIANT = col_character(),
  DBSNP_ID = col_character(),
  DBSNP_VARIANT = col_character(),
  OLD_ID = col_character(),
  OLD_VARIANT = col_character()
)



```

## Step 1: Importing Exposure and Outcome Data
```{r ld}
## File paths needed for the analysis
LD.filepath = "Resources/LDscores_filtered.csv" # LD scores
rho.filepath = "Resources/LD_GM2_2prm.csv" # local/SNP-specific LD scores
ld = "Resources/eur_w_ld_chr/"
hm3 = "Resources/w_hm3.snplist"
```

## Exposure Import
Mitochondrial DNA count Number (X variable)

```{r read_exposure_Chong}

exposure_path = "Resources/Chong2022mtdnacn_eur.chrall.CPRA_b37.tsv.gz"
exposure_ss_Chong <- read_tsv(exposure_path, comment = "##", col_types = coltypes, 
                        col_select = c(DBSNP_ID, CHROM, POS, REF, ALT, AF, BETA, SE, Z, P, N)) %>%
  rename(SNP = DBSNP_ID)



```

```{r read_exposure_Longchamps}

Longchamps_path = "Resources/Longchamps2021mtdnacn.chrall.CPRA_b37.tsv.gz"
exposure_ss_Longchamps <- read_tsv(Longchamps_path, comment = "##", col_types = coltypes, 
                        col_select = c(DBSNP_ID, CHROM, POS, REF, ALT, AF, BETA, SE, Z, P, N)) %>%
  rename(SNP = DBSNP_ID)
```

```{r read_exposure_Hagg}

Hagg_path = "Resources/Hagg2020mtdnacn.chrall.CPRA_b37.tsv.gz"
exposure_ss_Hagg <- read_tsv(Hagg_path, comment = "##", col_types = coltypes, 
                        col_select = c(DBSNP_ID, CHROM, POS, REF, ALT, AF, BETA, SE, Z, P, N)) %>%
  rename(SNP = DBSNP_ID)
```

## Outcome Import 
Alzheimer's Disease (Y variable)

```{r read_outcome_Kunkle}
outcome_path = "Resources/Kunkle2019load_stage123.chrall.CPRA_b37.tsv.gz"
outcome_ss_Kunkle <- read_tsv(outcome_path, comment = "##",  col_types = coltypes, 
                        col_select = c(DBSNP_ID, CHROM, POS, REF, ALT, AF, BETA, SE, Z, P, N))  %>%
  rename(SNP = DBSNP_ID)
```

```{r read_outcome_Bellenguez}

Bellenguez_path = "Resources/Bellenguez2022load.chrall.CPRA_b37.tsv.gz"
outcome_ss_Bellenguez <- read_tsv(Bellenguez_path, comment = "##",  col_types = coltypes, 
                        col_select = c(DBSNP_ID, CHROM, POS, REF, ALT, AF, BETA, SE, Z, P, N))  %>%
  rename(SNP = DBSNP_ID)

```

## Filter APOE Region.
Removal of the Apoe region from both the Exposure and Outcome datasets. Creates
a new Binary variable apoe_region and filters dataset based on that column.

``` {r filter_exposures}

# For Exposure Chong

chong_apoe <- filter(exposure_ss_Chong , !(CHROM == 19 & between(POS, 44912079, 45912079)))

# For Exposure Longchamps

longchamps_apoe <- filter(exposure_ss_Longchamps , !(CHROM == 19 & between(POS, 44912079, 45912079)))

# For Exposure Hagg

hagg_apoe <- filter(exposure_ss_Hagg , !(CHROM == 19 & between(POS, 44912079, 45912079)))
```

``` {r filter_outcomes}

# For Outcome Kunkle.

kunkle_apoe <- filter(outcome_ss_Kunkle , !(CHROM == 19 & between(POS, 44912079, 45912079)))

# For Outcome Bellenguez.

bellenguez_apoe <- filter(outcome_ss_Bellenguez , !(CHROM == 19 & between(POS, 44912079, 45912079)))

```


## Merging Exposure and outcome datasets (X and Y) and perform LHC-MR

Will perform Multiple merging steps, as well as mulitple LHC-MR Analysis.
Warning! : be sure to move all LHC results to a different folder before performing
another LHC-MR paired analysis, as it will overwrite the previous ones results.
Please check your directory to see where the files were placed and place them 
in a new folder for safe keeping. 
Perform one LHC-MR at a time.
Running each LHC-MR pair will take approximately 6-12 hours depending on your machine
and the amount of cores specified in "ncores" (more cores speeds up the process).



## Chong-Kunkle-LHCMR

```{r merge_Chong_Kunkle}
## Step 1
trait.names=c("mtDNAcn_Apoe_removed","AD")
input.files = list(chong_apoe, kunkle_apoe)
df = merge_sumstats(input.files, trait.names, LD.filepath, rho.filepath)
```


## Step 2: Calculating smart starting points for the likelihood optimisation

```{r read_merge}
SP_list = calculate_SP(df,trait.names,run_ldsc=TRUE,run_MR=TRUE,hm3=hm3,ld="Resources/eur_w_ld_chr",nStep = 2,
                       SP_single=3,SP_pair=50,SNP_filter=10)
```

## Step 3: Running the likelihood optimisation to estimate the parameters, followed by a block-jackknife procedure to calculate parameter-SE

```{r read_merge_LHCMR}
## Step 3
res_Chong_Kunkle  = lhc_mr(SP_list, trait.names, paral_method="lapply", nCores = 1, nBlock=200)
res_Chong_Kunkle  %>% as_tibble() %>% write_csv(., "results/LHC_MR_Results_Chong_Kunkle/lhcmr_Chong_Kunkle_apoe_removed.csv")


view(res_Chong_Kunkle)



```

## Chong-Bellenguez-LHCMR

```{r merge_Chong_Bellenguez}
## Step 1
trait.names=c("mtDNAcn_Apoe_removed_chong","AD_bellenguez")
input.files = list(chong_apoe, bellenguez_apoe)
df = merge_sumstats(input.files, trait.names, LD.filepath, rho.filepath)
```


## Step 2: Calculating smart starting points for the likelihood optimisation

```{r read_merge}
SP_list = calculate_SP(df,trait.names,run_ldsc=TRUE,run_MR=TRUE,hm3=hm3,ld="Resources/eur_w_ld_chr",nStep = 2,
                       SP_single=3,SP_pair=50,SNP_filter=10)
```

## Step 3: Running the likelihood optimisation to estimate the parameters, followed by a block-jackknife procedure to calculate parameter-SE

```{r read_merge_LHCMR}
## Step 3
res_Chong_Bellenguez = lhc_mr(SP_list, trait.names, paral_method="lapply", nCores = 1, nBlock=200)
res_Chong_Bellenguez %>% as_tibble() %>% write_csv(., "results/LHC_MR_Results_Chong_Bellenguez/lhcmr_Chong_bellenguez_apoe_removed.csv")


view(res_Chong_Bellenguez)


```

## Longchamps-Bellenguez-LHCMR

```{r merge_longchamps_Bellenguez}
## Step 1
trait.names=c("mtDNAcn_Apoe_omit_longchamps","AD_bellenguez")
input.files = list(longchamps_apoe, bellenguez_apoe)
df = merge_sumstats(input.files, trait.names, LD.filepath, rho.filepath)
```


## Step 2: Calculating smart starting points for the likelihood optimisation

```{r read_merge}
SP_list = calculate_SP(df,trait.names,run_ldsc=TRUE,run_MR=TRUE,hm3=hm3,ld="Resources/eur_w_ld_chr",nStep = 2,
                       SP_single=3,SP_pair=50,SNP_filter=10)
```

## Step 3: Running the likelihood optimisation to estimate the parameters, followed by a block-jackknife procedure to calculate parameter-SE

```{r read_merge_LHCMR}
## Step 3
res_longchamps_Bellenguez = lhc_mr(SP_list, trait.names, paral_method="lapply", nCores = 1, nBlock=200)
res_longchamps_Bellenguez %>% as_tibble() %>% write_csv(., "results/LHC_MR_Results_Longchamps_Bellenguez/lhcmr_Longchamps_bellenguez_apoe_omit.csv")


view(res_longchamps_Bellenguez)


```


## Hagg-Bellenguez-LHCMR


```{r merge_Hagg_Bellenguez}
## Step 1
trait.names=c("mtDNAcn_Apoe_omit_hagg","AD_bellenguez")
input.files = list(hagg_apoe, bellenguez_apoe)
df = merge_sumstats(input.files, trait.names, LD.filepath, rho.filepath)
```


## Step 2: Calculating smart starting points for the likelihood optimisation

```{r read_merge}
SP_list = calculate_SP(df,trait.names,run_ldsc=TRUE,run_MR=TRUE,hm3=hm3,ld="Resources/eur_w_ld_chr",nStep = 2,
                       SP_single=3,SP_pair=50,SNP_filter=10)
```

## Step 3: Running the likelihood optimisation to estimate the parameters, followed by a block-jackknife procedure to calculate parameter-SE

```{r read_merge_LHCMR}
## Step 3
res_Hagg_Bellenguez = lhc_mr(SP_list, trait.names, paral_method="lapply", nCores = 1, nBlock=200)
res_Hagg_Bellenguez %>% as_tibble() %>% write_csv(., "results/LHC_MR_Results_Hagg_Bellenguez/lhcmr_Hagg_bellenguez_apoe_omit.csv")


view(res_Hagg_Bellenguez)


```


## MR Metaanalysis Forest Plots (Direction X > Y and Y > X)


``` {r apoe_omitted_meta_plot}
#Creating new data frame based on LHC MR results(Syntax of LHCMR results are
# Row1:betas/logodds, row2:se, row3:pvalue)
View(res_Chong_Bellenguez)

View(res_longchamps_Bellenguez)

View(res_Hagg_Bellenguez)



(All Directions plot)

LHCMR_forest_apoe_removed <- tribble(
  ~b, ~se, ~p_value, ~dataset, ~Causal_direction, ~Confounder_b, ~Confounder_se, ~Confounder_pvalue,
  #--|--|----
  -0.1755506, 0.03822413, 0.000004376, "chong_bellenguez", "mtDNAcn_on_AD", 0.0177505572686241, 0.0266442980004017, 0.505280252106757,
  -0.117470737, 0.040403246, 0.003643794, "longchamps_bellenguez", "mtDNAcn_on_AD", 0.0562304589342615, 0.0211505225955438, 0.00784695888183813, 
  -0.1750155, .03901719, 0.00000727, "hagg_bellenguez", "mtDNAcn_on_AD", 0.0545978599098101, 0.0268700849036229, 0.0421617667978208,
  -0.1792789, 0.2746141, 0.5138598, "chong_bellenguez", "AD_on_mtDNAcn", -0.0488227464177553, 0.0168226370957442, 0.00370545624198806,
  0.09668307, 0.04646173, 0.03744136, "longchamps_bellenguez", "AD_on_mtDNAcn", -0.0411582309529671, 0.015494269317608, 0.00789911504532021,
  -0.1409825, 0.1444253, 0.3289844, "hagg_bellenguez", "AD_on_mtDNAcn", 0.0350185170177856, 0.0737222034913936, 0.634782425737017,
)


view(LHCMR_forest_apoe_removed)

LHC_OR_apoe_removed <- generate_odds_ratios(LHCMR_forest_apoe_removed)

view(LHC_OR_apoe_removed)


LHCMR_apoe_removed_forest_mod <- ggplot(data=LHC_OR_apoe_removed, aes(x=dataset, y=or, ymin=or_lci95, ymax=or_uci95)) +
   facet_grid(cols = vars(Causal_direction)) + 
geom_point(shape = 15, size = 2) +
geom_pointrange()+ 
  geom_hline(yintercept=1, lty=2, size =1) +  
  geom_errorbar(aes(ymin=or_lci95, ymax=or_uci95), width=0.5, cex=1)+ 
  coord_flip() + 
  geom_point(shape = 15, size = 2) +
  ggtitle("Apoe removed LHC_MR Odds ratio by causal direction.") +
  xlab("Method")  +
  ylab("Odds Ratio AD due to mtDNAcn  (95% CI)")  
  


  
 

LHCMR_apoe_removed_forest_mod

# Forward direction plot only (Reverse direction is binary exposure on continous outcome, not a odds ratio)

LHCMR_forest_apoe_removed_mtDNA_on_AD <- tribble(
  ~b, ~se, ~p_value, ~dataset, ~Causal_direction, ~Confounder_b, ~Confounder_se, ~Confounder_pvalue,
  #--|--|----
  -0.1755506, 0.03822413, 0.000004376, "Chong", "mtDNAcn_on_AD", 0.0177505572686241, 0.0266442980004017, 0.505280252106757,
  -0.117470737, 0.040403246, 0.003643794, "Longchamps", "mtDNAcn_on_AD", 0.0562304589342615, 0.0211505225955438, 0.00784695888183813, 
  -0.1750155, .03901719, 0.00000727, "Hagg", "mtDNAcn_on_AD", 0.0545978599098101, 0.0268700849036229, 0.0421617667978208,
  
)

LHC_OR_apoe_removed_mtDNA_on_AD <- generate_odds_ratios(LHCMR_forest_apoe_removed_mtDNA_on_AD)

view(LHC_OR_apoe_removed_mtDNA_on_AD)

LHCMR_apoe_removed_forest_mod_mtDNA_on_AD <- ggplot(data=LHC_OR_apoe_removed_mtDNA_on_AD, aes(x=dataset, y=or, ymin=or_lci95, ymax=or_uci95)) +
   facet_grid(cols = vars(Causal_direction)) + 
geom_point(shape = 15, size = 2) +
geom_pointrange()+ 
  geom_hline(yintercept=1, lty=2, size =1) +  
  geom_errorbar(aes(ymin=or_lci95, ymax=or_uci95), width=0.5, cex=1)+ 
  coord_flip() + 
  geom_point(shape = 15, size = 2) +
  ggtitle("Apoe removed LHC_MR Odds ratio by causal direction.") +
  xlab("Method")  +
  ylab("Odds Ratio AD due to mtDNAcn  (95% CI)")  


# CI Plot for reverse causal direction, new table.

LHCMR_forest_apoe_removed_AD_on_mtDNA <- tribble(
  ~b, ~dataset, ~Causal_direction, ~lci95, ~uci95, 
   #--|--|----
  -0.1792789, "Chong", "AD_on_mtDNAcn", -0.719759043303374, 0.356728264230965,
	 0.09668307, "Longchamps", "AD_on_mtDNAcn", 0.0108396654302798, 0.192969635729239,
  -0.1409825, "Hagg", "AD_on_mtDNAcn", -0.42279866490016, 0.143348700286467
)

LHCMR_apoe_removed_forest_mod_AD_on_mtDNA <- ggplot(data=LHCMR_forest_apoe_removed_AD_on_mtDNA, aes(x=dataset, y=b, ymin=lci95, ymax=uci95)) +
   facet_grid(cols = vars(Causal_direction)) + 
geom_point(shape = 15, size = 2) +
geom_pointrange()+ 
  geom_hline(yintercept=0, lty=2, size =1) +  
  geom_errorbar(aes(ymin=lci95, ymax=uci95), width=0.5, cex=1)+ 
  coord_flip() + 
  geom_point(shape = 15, size = 2) +
  ggtitle("Causal estimate of AD on mtDNAcn,  APOE omitted (95% CI).") +
  xlab("Method")  +
  ylab("Causal estimate of AD on mtDNAcn (95% CI)")  

LHCMR_apoe_removed_forest_mod_AD_on_mtDNA

```

# Export forest plots of LHC_MR

``` {r forest_plots_combine_export}

ggsave('Casual_Estimate_mtDNAcn_on_Alzheimer/results/MR_Results_all_datasets/LHCmr_omit_apoe_forestplot.png', plot = LHCMR_apoe_removed_forest_mod)

ggsave('results/MR_Results_all_datasets/LHCmr_omit_apoe_forestplot_mtDNA_on_AD.png', plot = LHCMR_apoe_removed_forest_mod_mtDNA_on_AD)

ggsave('results/MR_Results_all_datasets/LHCmr_omit_apoe_forestplot_AD_on_mtDNA.png', plot = LHCMR_apoe_removed_forest_mod_AD_on_mtDNA)
```


