---
title: "6 Pair MVMR"
author: "Brian Alvarez"
date: "2023-02-20"
output: html_document
---

```{r setup, include=FALSE}

# Note: Run this code block manually by highlighting and running the individual lines.
knitr::opts_chunk$set(echo = TRUE)

knitr::opts_knit$set(root.dir = 'C:/Users/brian/OneDrive/Documents/GitHub/Casual_Estimate_mtDNAcn_on_Alzheimer')


```



## Libraries

Loads the appropriate packages for MR analysis

``` {r eval=FALSE}
library(tidyverse) # Data wrangling 

library(cowplot)

library(ggman)

library(TwoSampleMR) # MR 

library(RadialMR) # Radial MR sensitivity analysis 

library(phenoscanner)

library(LDlinkR) # LD and proxy snps

library(janitor)

library(MVMR)         # Multivarble MR

library(RMVMR)


# Define column types for summary statistics (MDNA and AD GWAS)
coltypes = cols(
  ID = col_character(),
  CHROM = col_double(),
  POS = col_double(),
  REF = col_character(),
  ALT = col_character(),
  AF = col_double(),
  TRAIT = col_character(),
  BETA = col_double(),
  SE = col_double(),
  Z = col_double(),
  P = col_double(),
  N = col_double(),
  OR = col_double(),
  OR_L95 = col_double(),
  OR_U95 = col_double(),
  DIR = col_character(),
  G1000_ID = col_character(),
  G1000_VARIANT = col_character(),
  DBSNP_ID = col_character(),
  DBSNP_VARIANT = col_character(),
  OLD_ID = col_character(),
  OLD_VARIANT = col_character()
)




```


``` {r read_in_chong}

exposure_path_chong = "Resources/Chong2022mtdnacn_eur.chrall.CPRA_b37.tsv.gz"

Exposure_mtDNAcn_chong <- read_tsv(exposure_path_chong,comment="#")


# To format the GWAS results for MR analysis Exposure. 

# Rename right-side variables to match column names of your GWAS dataset 
Format_mtDNAcn_chong <- Exposure_mtDNAcn_chong %>% format_data(.,
  type = "exposure",
  snps = NULL,
  header = TRUE,
  phenotype_col = "TRAIT",
  snp_col = "DBSNP_ID",
  beta_col = "BETA",
  se_col = "SE",
  eaf_col = "AF",
  effect_allele_col = "REF",
  other_allele_col = "ALT",
  pval_col = "P",
  units_col = "units",
  ncase_col = "N_CASES",
  ncontrol_col = "N_CTRLS",
  samplesize_col = "N",
  gene_col = "gene",
  min_pval = 1e-200,
  z_col = "Z",
  info_col = "info",
  chr_col = "CHROM",
  pos_col = "POS",
  log_pval = FALSE
)  %>%
  as_tibble()

Format_mtDNAcn_chong <- Format_mtDNAcn_chong %>%
mutate(exposure = "mtDNA_chong")



# Chong Filter and Clumping
mtDNAcn_filter_chong <- filter(Format_mtDNAcn_chong , pval.exposure < 5e-8)


mtDNAcn_clump_chong <- clump_data(mtDNAcn_filter_chong, clump_kb = 10000,
                            clump_r2 = 0.001,
                            clump_p1 = 1,
                            clump_p2 = 1,
                            pop = "EUR"
)%>% 
  as_tibble()

```


``` {r read_in_longchamps}

exposure_path_longchamps = "Resources/Longchamps2021mtdnacn.chrall.CPRA_b37.tsv.gz"

Exposure_mtDNAcn_longchamps <- read_tsv(exposure_path_longchamps,comment="#")


# To format the GWAS results for MR analysis Exposure. 
# Rename right-side variables to match column names of your GWAS dataset.
Format_mtDNAcn_longchamps <- Exposure_mtDNAcn_longchamps %>% format_data(.,
  type = "exposure",
  snps = NULL,
  header = TRUE,
  phenotype_col = "TRAIT",
  snp_col = "DBSNP_ID",
  beta_col = "BETA",
  se_col = "SE",
  eaf_col = "AF",
  effect_allele_col = "REF",
  other_allele_col = "ALT",
  pval_col = "P",
  units_col = "units",
  ncase_col = "N_CASES",
  ncontrol_col = "N_CTRLS",
  samplesize_col = "N",
  gene_col = "gene",
  min_pval = 1e-200,
  z_col = "Z",
  info_col = "info",
  chr_col = "CHROM",
  pos_col = "POS",
  log_pval = FALSE
)  %>%
  as_tibble()


Format_mtDNAcn_longchamps <- Format_mtDNAcn_longchamps %>%
mutate(exposure = "mtDNA_longchamps")


# longchamps Filter and Clumping

mtDNAcn_filter_longchamps <- filter(Format_mtDNAcn_longchamps, pval.exposure < 5e-8)


mtDNAcn_clump_longchamps <- clump_data(mtDNAcn_filter_longchamps, clump_kb = 10000,
                            clump_r2 = 0.001,
                            clump_p1 = 1,
                            clump_p2 = 1,
                            pop = "EUR"
)%>% 
  as_tibble()

```



``` {r read_in_hagg}

exposure_path_hagg = "Resources/Hagg2020mtdnacn.chrall.CPRA_b37.tsv.gz"

Exposure_mtDNAcn_hagg <- read_tsv(exposure_path_hagg,comment="#")


# Format the GWAS results for MR analysis Exposure. 

# Rename right-side variables to match column names of your GWAS dataset.
Format_mtDNAcn_hagg <- Exposure_mtDNAcn_hagg %>% format_data(.,
  type = "exposure",
  snps = NULL,
  header = TRUE,
  phenotype_col = "TRAIT",
  snp_col = "DBSNP_ID",
  beta_col = "BETA",
  se_col = "SE",
  eaf_col = "AF",
  effect_allele_col = "REF",
  other_allele_col = "ALT",
  pval_col = "P",
  units_col = "units",
  ncase_col = "N_CASES",
  ncontrol_col = "N_CTRLS",
  samplesize_col = "N",
  gene_col = "gene",
  min_pval = 1e-200,
  z_col = "Z",
  info_col = "info",
  chr_col = "CHROM",
  pos_col = "POS",
  log_pval = FALSE
)  %>%
  as_tibble()

Format_mtDNAcn_hagg <- Format_mtDNAcn_hagg %>%
mutate(exposure = "mtDNA_hagg")

# hagg Filter and Clumping
mtDNAcn_filter_hagg <- filter(Format_mtDNAcn_hagg, pval.exposure < 5e-8)


mtDNAcn_clump_hagg <- clump_data(mtDNAcn_filter_hagg, clump_kb = 10000,
                            clump_r2 = 0.001,
                            clump_p1 = 1,
                            clump_p2 = 1,
                            pop = "EUR"
)%>% 
  as_tibble()

```


## Exposure 2: Blood Cell Composition 

This will be adjusting for differences in relative mtDNA count number in different
cell types in our exposures.

```{r exposure2}

exposure2_path_mod = "Resources/Astle2016blood_platelets.h.tsv.gz"

exposure2_intermediate <- read_tsv(exposure2_path_mod, comment = "##")

exposure2_intermediate_2 <- exposure2_intermediate %>%
rename(CHROM = chromosome, 
        POS = hm_pos,
        REF = effect_allele,
        ALT = other_allele, 
        BETA = beta, 
        SE = standard_error,
        P = p_value, 
        AF = effect_allele_frequency,
        OR = odds_ratio, 
        OR_L95 = ci_lower, 
        OR_U95 = ci_upper,
        DBSNP_ID = hm_rsid  )
           
exposure2_intermediate_3 <- filter(exposure2_intermediate_2, P < 5e-8)

exposure2_path <- filter(exposure2_intermediate_3, !(CHROM == 19 & between(POS, 44912079, 45912079)))

         
# Format data to TwoSampleMR format
exposure2 <- exposure2_path %>%
  format_data(.,
    type = "exposure",
    snps = NULL,
    header = TRUE,
    snp_col = "DBSNP_ID",
    beta_col = "BETA",
    se_col = "SE",
    eaf_col = "AF",
    effect_allele_col = "ALT",
    other_allele_col = "REF",
    pval_col = "P",
    chr_col = "CHROM",
    pos_col = "POS",
    log_pval = FALSE
) %>%
  as_tibble()

exposure2 <- exposure2 %>%
mutate(exposure = "Platlet_cn")


# Perform LD clumping on SNP data, filter SNPs to make it run faster
exposure2_clump <- exposure2 %>% 
  filter(pval.exposure < 0.01) %>%
  clump_data(.,
  clump_kb = 10000,
  clump_r2 = 0.001,
  clump_p1 = 1,
  clump_p2 = 1,
  pop = "EUR"
)

# Filter and Removal of Apoe Region SNPs.
exposure2_clump_inter <- filter(exposure2_clump, pval.exposure < 5e-8)

exposure2_dat <- filter(exposure2_clump_inter, !(chr.exposure == 19 & between(pos.exposure, 44912079, 45912079)))

# Export Clumped and Filtered dataset.

write_csv(exposure2_dat, 'Resources/platelets_clumped_dataset.csv')

```

 


## Importing and formatting Outcome

This step imports our GWAS Outcome data sets and formats them up for MR analysis

``` {r read_in_Bellenguez}

outcome_path_bellenguez = "Resources/Bellenguez2022load.chrall.CPRA_b37.tsv.gz"

# Import Code for outcome GWAS 
OutcomeAD_bellenguez <- read_tsv(outcome_path_bellenguez,comment="#")

# Rename right-side variables to match column names of your GWAS dataset 

outcome_AD_Format_bellenguez <- OutcomeAD_bellenguez %>%
  format_data(.,
    type = "outcome",
    snps = NULL,
    header = TRUE,
    phenotype_col = "Phenotype",
    snp_col = "DBSNP_ID",
    beta_col = "BETA",
    se_col = "SE",
    eaf_col = "AF",
    effect_allele_col = "ALT",
    other_allele_col = "REF",
    pval_col = "P",
    samplesize_col = "N",
    z_col = "Z",
    chr_col = "CHROM",
    pos_col = "POS",
    log_pval = FALSE
) %>%
  as_tibble()

outcome_AD_Format_bellenguez <- outcome_AD_Format_bellenguez %>%
mutate(outcome = "bellenguez")



```



``` {r read_in_Kunkle}


outcome_path_kunkle = "Resources/Kunkle2019load_stage123.chrall.CPRA_b37.tsv.gz"

# Import Code for outcome GWAS 
OutcomeAD_kunkle <- read_tsv(outcome_path_kunkle,comment="#")

# Rename right-side variables to match column names of your GWAS dataset 

outcome_AD_Format_kunkle <- OutcomeAD_kunkle %>%
  format_data(.,
    type = "outcome",
    snps = NULL,
    header = TRUE,
    phenotype_col = "Phenotype",
    snp_col = "DBSNP_ID",
    beta_col = "BETA",
    se_col = "SE",
    eaf_col = "AF",
    effect_allele_col = "ALT",
    other_allele_col = "REF",
    pval_col = "P",
    samplesize_col = "N",
    z_col = "Z",
    chr_col = "CHROM",
    pos_col = "POS",
    log_pval = FALSE
) %>%
  as_tibble()

outcome_AD_Format_kunkle <- outcome_AD_Format_kunkle %>%
mutate(outcome = "kunkle")

```

## Filter APOE Region. 
Removal of the Apoe region from both the Exposure and Outcome datasets.

``` {r filter_exposures}


# These Apoe SNP's will be removed in the subsequent MR analysis.

# For Chong 


mtDNAcn_clump_chong <- filter(mtDNAcn_clump_chong, !(chr.exposure == 19 & between(pos.exposure, 44912079, 45912079)))

view(mtDNAcn_clump_chong)
# For Longchamps 


mtDNAcn_clump_longchamps <- filter(mtDNAcn_clump_longchamps, !(chr.exposure == 19 & between(pos.exposure, 44912079, 45912079)))
view(mtDNAcn_clump_longchamps)


# For Hagg 

mtDNAcn_clump_hagg <- filter(mtDNAcn_clump_hagg, !(chr.exposure == 19 & between(pos.exposure, 44912079, 45912079)))

view(mtDNAcn_clump_hagg)









```



## Run MVMR on all exposure datasets via trios (mtDNA/Bloodtype/Outcome) 

Run the combine 3 pairs with ld proxy twice, one for each outcome (bellenguez and kunkle)
Chong_exposure2_kunkle
longchamps_exposure2_kunkle
hagg_exposure2_kunkle

Chong_exposure2_bellenguez
longchamps_exposure2_bellenguez
hagg_exposure2_bellenguez

MVMR is run 1 at a time for each trio


## MVMR Chong_exposure2_kunkle

# Harmonization 

Involves two rounds of LD proxies, with 1 round of exposure 1 and exposure 2 harmonization
in between.
1st. LD Proxies exposure 1 with exposure 2 as "outcome"
2nd. Exposure 1 and 2 harmonization.
3rd. LD proxies exposure1_2 combined with outcome AD Kunkle

# Step 1: LD Proxies exposure 1 with exposure 2 as "outcome"
``` {r chong_exposure2_LDproxy}

# extract exposure SNPs present in outcome (outcome datasets,then exposure dataset)
outcome_clump <- semi_join(
  exposure2_dat, mtDNAcn_clump_chong, by = "SNP"
)

# Exposure SNPs not present in outcome (exposure dataset, then outcome dataset)
exp_snps_wo <- anti_join(
  mtDNAcn_clump_chong, exposure2_dat, by = "SNP"
)

# My LD_Proxy code 7dbc1afe7454



# Use LDLinkR to identify proxy snps based on SNP dataset above
# that presents SNPs with no corresponding match in the other.
LDproxy_batch(exp_snps_wo$SNP, 
        pop = "CEU", 
        r2d = "r2", 
        token = '7dbc1afe7454', 
        append = TRUE,
        genome_build = "grch37")
system("mv combined_query_snp_list_grch37.txt data/exposure_outcome_proxy_snps_chong_plate.txt")


# Munge proxy snp file
proxy_snps <- read_tsv("data/exposure_outcome_proxy_snps_chong_plate.txt", skip = 1, col_names = F) %>%
  rename(id = X1, func = X2, proxy_snp = X3, coord = X4, alleles = X5, maf = X6, 
         distance = X7, dprime = X8, rsq = X9, correlated_alleles = X10, FORGEdb = X11, RegulomeDB = X12) %>%
  separate(coord, c('chr', 'pos'), sep = ":") %>%
  mutate(snp = ifelse(id == 1, proxy_snp, NA), 
         chr = str_replace(chr, 'chr', ""), 
         chr = as.numeric(chr), 
         pos = as.numeric(pos)) %>%
  fill(snp, .direction = 'down') %>%
  relocate(snp, .before = proxy_snp) %>%
  dplyr::select(-id, -func, -FORGEdb, -RegulomeDB) %>%
  filter(rsq >= 0.8)

# Munge proxy snp and outcome data 
proxy_outcome <- left_join(
  proxy_snps, exposure2_dat, by = c("proxy_snp" = "SNP")
) %>%
  separate(correlated_alleles, c("target_a1.outcome", "proxy_a1.outcome", 
                                 "target_a2.outcome", "proxy_a2.outcome"), sep = ",|=") %>%
  filter(!is.na(chr.outcome)) %>%
  arrange(snp, -rsq, abs(distance)) %>%
  group_by(snp) %>%
  slice(1) %>%
  ungroup() %>%
  mutate(
     proxy.outcome = TRUE,
     target_snp.outcome = snp,
     proxy_snp.outcome = proxy_snp, 
  ) %>% 
  mutate(
       new_effect_allele.outcome = case_when(
        proxy_a1.outcome == effect_allele.outcome & proxy_a2.outcome == other_allele.outcome ~ target_a1.outcome,
        proxy_a2.outcome == effect_allele.outcome & proxy_a1.outcome == other_allele.outcome ~ target_a2.outcome,
        TRUE ~ NA_character_
     ), 
      new_other_allele.outcome = case_when(
        proxy_a1.outcome == effect_allele.outcome & proxy_a2.outcome == other_allele.outcome ~ target_a2.outcome,
        proxy_a2.outcome == effect_allele.outcome & proxy_a1.outcome == other_allele.outcome ~ target_a1.outcome,
        TRUE ~ NA_character_
     ), 
     effect_allele.outcome = new_effect_allele.outcome, 
     other_allele.outcome = new_other_allele.outcome
  ) %>%
  dplyr::select(-proxy_snp, -chr, -pos, -alleles, -maf, -distance, -rsq, -dprime,  
         -new_effect_allele.outcome, -new_other_allele.outcome) %>%
  relocate(target_a1.outcome, proxy_a1.outcome, target_a2.outcome, proxy_a2.outcome, .after = proxy_snp.outcome) %>%
  rename(SNP = snp) %>%
  relocate(SNP, .after = samplesize.outcome)


# Merge outcome and proxy outcomes
chong_plate_dat_proxy <- bind_rows(
  outcome_clump, proxy_outcome
) %>% 
  arrange(chr.outcome, pos.outcome)

```

# 2nd: Exposure 1 and 2 harmonization. 

``` {r mtDNAcn_clump_chong_harmony}

#Temporary name change of exposures for harmonzation.

exposure1_dat <- chong_plate_dat_proxy # Clumped and LD proxy dataset.

exposure1 <- Format_mtDNAcn_chong # Exposure1  not yet clumped



mvmr_snps <- bind_rows(
  select(exposure1_dat, SNP), 
  select(exposure2_dat, SNP), 
) %>%
  distinct(SNP) %>%
  pull(SNP)
comb_exp <- bind_rows(
  exposure1 %>% filter(SNP %in% mvmr_snps),
  exposure2 %>% filter(SNP %in% mvmr_snps),
)
comb_exp_out <- bind_rows(
  exposure1 %>% filter(SNP %in% mvmr_snps) %>% 
    rename(pval = pval.exposure, trait = exposure, chr = chr.exposure, pos = pos.exposure), 
  exposure2 %>% filter(SNP %in% mvmr_snps) %>% 
    rename(pval = pval.exposure, trait = exposure, chr = chr.exposure, pos = pos.exposure), 
  outcome_AD_Format_kunkle %>% filter(SNP %in% mvmr_snps) %>% 
    rename(pval = pval.outcome, trait = outcome, chr = chr.outcome, pos = pos.outcome)
)
## How many SNPs are GWS across the exposures 
comb_exp_wide <- comb_exp_out %>%
   arrange(chr, pos) %>%
   select(SNP, trait, pval) %>%
   pivot_wider(names_from = trait, values_from = pval) %>%
   mutate(pval.exposure = pmin(mtDNA_chong, Platlet_cn, na.rm = T))
comb_exp_wide %>% count(mtDNA_chong < 5e-8, Platlet_cn < 5e-8)
comb_exp_wide %>% count(mtDNA_chong < 5e-8, Platlet_cn < 5e-8, kunkle < 5e-8)
## Do another round of LD clumping 
comb_exp_clump <- comb_exp %>% 
  clump_data(.,
  clump_kb = 10000,
  clump_r2 = 0.001,
  clump_p1 = 1,
  clump_p2 = 1,
  pop = "EUR"
)
comb_exp_clump %>% count(mtDNA_chong < 5e-8, Platlet_cn < 5e-8)
comb_exp <- comb_exp %>% filter(SNP %in% comb_exp_clump$SNP) 
```

# Step 3: LD Proxies with 2 exposures combined 
``` {r chong_kunkle_LDproxy}

# extract exposure SNPs present in outcome (outcome datasets,then exposure dataset)
outcome_clump <- semi_join(
  outcome_AD_Format_kunkle, comb_exp, by = "SNP"
)

# Exposure SNPs not present in outcome (exposure dataset, then outcome dataset)
exp_snps_wo <- anti_join(
  comb_exp, outcome_AD_Format_kunkle, by = "SNP"
)


# Use LDLinkR to identify proxy snps based on SNP dataset above
# that presents SNPs with no corresponding match in the other.
LDproxy_batch(exp_snps_wo$SNP, 
        pop = "CEU", 
        r2d = "r2", 
        token = 'a6deee62cc4a', 
        append = TRUE,
        genome_build = "grch37")
system("mv combined_query_snp_list_grch37.txt data/exposure_outcome_proxy_snps_chong_plate_kunkle.txt")


# Munge proxy snp file
proxy_snps <- read_tsv("data/exposure_outcome_proxy_snps_chong_plate_kunkle.txt", skip = 1, col_names = F) %>%
  rename(id = X1, func = X2, proxy_snp = X3, coord = X4, alleles = X5, maf = X6, 
         distance = X7, dprime = X8, rsq = X9, correlated_alleles = X10, FORGEdb = X11, RegulomeDB = X12) %>%
  separate(coord, c('chr', 'pos'), sep = ":") %>%
  mutate(snp = ifelse(id == 1, proxy_snp, NA), 
         chr = str_replace(chr, 'chr', ""), 
         chr = as.numeric(chr), 
         pos = as.numeric(pos)) %>%
  fill(snp, .direction = 'down') %>%
  relocate(snp, .before = proxy_snp) %>%
  dplyr::select(-id, -func, -FORGEdb, -RegulomeDB) %>%
  filter(rsq >= 0.8)

# Munge proxy snp and outcome data 
proxy_outcome <- left_join(
  proxy_snps, outcome_AD_Format_kunkle, by = c("proxy_snp" = "SNP")
) %>%
  separate(correlated_alleles, c("target_a1.outcome", "proxy_a1.outcome", 
                                 "target_a2.outcome", "proxy_a2.outcome"), sep = ",|=") %>%
  filter(!is.na(chr.outcome)) %>%
  arrange(snp, -rsq, abs(distance)) %>%
  group_by(snp) %>%
  slice(1) %>%
  ungroup() %>%
  mutate(
     proxy.outcome = TRUE,
     target_snp.outcome = snp,
     proxy_snp.outcome = proxy_snp, 
  ) %>% 
  mutate(
       new_effect_allele.outcome = case_when(
        proxy_a1.outcome == effect_allele.outcome & proxy_a2.outcome == other_allele.outcome ~ target_a1.outcome,
        proxy_a2.outcome == effect_allele.outcome & proxy_a1.outcome == other_allele.outcome ~ target_a2.outcome,
        TRUE ~ NA_character_
     ), 
      new_other_allele.outcome = case_when(
        proxy_a1.outcome == effect_allele.outcome & proxy_a2.outcome == other_allele.outcome ~ target_a2.outcome,
        proxy_a2.outcome == effect_allele.outcome & proxy_a1.outcome == other_allele.outcome ~ target_a1.outcome,
        TRUE ~ NA_character_
     ), 
     effect_allele.outcome = new_effect_allele.outcome, 
     other_allele.outcome = new_other_allele.outcome
  ) %>%
  dplyr::select(-proxy_snp, -chr, -pos, -alleles, -maf, -distance, -rsq, -dprime,  
         -new_effect_allele.outcome, -new_other_allele.outcome) %>%
  relocate(target_a1.outcome, proxy_a1.outcome, target_a2.outcome, proxy_a2.outcome, .after = proxy_snp.outcome) %>%
  rename(SNP = snp) %>%
  relocate(SNP, .after = samplesize.outcome)


# Merge outcome and proxy outcomes
chong_kunkle_dat_proxy <- bind_rows(
  outcome_clump, proxy_outcome
) %>% 
  arrange(chr.outcome, pos.outcome)

outcome_dat <- chong_kunkle_dat_proxy

```


## Venn diagram of GWAS SNPS (Brian, you must download package first.)
```{r VennDiagram}
library(nVennR)
## GWS SNPs
set1 <- list(pull(exposure1_dat, SNP))
set2 <- list(pull(exposure2_dat, SNP))
myNV <- plotVenn(list(set1, set2), sNames=c("mtDNA_chong", "Platlet_cn"), 
                 showPlot = F)
showSVG(myNV, opacity=0.2, outFile = "results/plots/mvmr_gws_venn.svg")
## All SNPs 
set1 <- list(exposure1 %>% filter(SNP %in% mvmr_snps, pval.exposure < 5e-8) %>% pull(SNP))
set2 <- list(exposure2 %>% filter(SNP %in% mvmr_snps, pval.exposure < 5e-8) %>% pull(SNP))
set3 <- list(outcome %>% filter(SNP %in% mvmr_snps) %>% pull(SNP))
myNV <- plotVenn(list(set1, set2, set3), sNames=c("mtDNA_chong", "Platlet_cn", "kunkle"), 
                 showPlot = F)
showSVG(myNV, opacity=0.2, outFile = "results/plots/mvmr_venn_chong_kunkle.svg")
```

## Harmonize Exposure1_Exposure2_Outcome

``` {r harm_exposures_outcome}

mvdat <- mv_harmonise_data(comb_exp, outcome_dat)

```

# Export Code
```{r export_harmonized_dat}
write_rds(mvdat, 'data/harmonized_mvmr_chong_platelet_kunkle.rds')
```

## Multivariate MR

```{r TwoSampleMR_MVMR}
## Multivariable MR
mvmr_res <- mv_multiple(mvdat, plots = T)
mvmr_res$result
```


``` {r export}
write_csv(mvmr_res$result, 'results/mvmr_res_chong_platelet_kunkle.csv')

```





## MVMR longchamps_exposure2_kunkle

# Harmonization 

Involves two rounds of LD proxies, with 1 round of exposure 1 and exposure 2 harmonization
in between.
1st. LD Proxies exposure 1 with exposure 2 as "outcome"
2nd. Exposure 1 and 2 harmonization.
3rd. LD proxies exposure1_2 combined with outcome AD Kunkle

# Step 1: LD Proxies exposure 1 with exposure 2 as "outcome"
``` {r longchamps_exposure2_LDproxy}

# extract exposure SNPs present in outcome (outcome datasets,then exposure dataset)
outcome_clump <- semi_join(
  exposure2_dat, mtDNAcn_clump_longchamps, by = "SNP"
)

# Exposure SNPs not present in outcome (exposure dataset, then outcome dataset)
exp_snps_wo <- anti_join(
  mtDNAcn_clump_longchamps, exposure2_dat, by = "SNP"
)

# My LD_Proxy code 7dbc1afe7454



# Use LDLinkR to identify proxy snps based on SNP dataset above
# that presents SNPs with no corresponding match in the other.
LDproxy_batch(exp_snps_wo$SNP, 
        pop = "CEU", 
        r2d = "r2", 
        token = '7dbc1afe7454', 
        append = TRUE,
        genome_build = "grch37")
system("mv combined_query_snp_list_grch37.txt data/exposure_outcome_proxy_snps_longchamps_plate.txt")


# Munge proxy snp file
proxy_snps <- read_tsv("data/exposure_outcome_proxy_snps_longchamps_plate.txt", skip = 1, col_names = F) %>%
  rename(id = X1, func = X2, proxy_snp = X3, coord = X4, alleles = X5, maf = X6, 
         distance = X7, dprime = X8, rsq = X9, correlated_alleles = X10, FORGEdb = X11, RegulomeDB = X12) %>%
  separate(coord, c('chr', 'pos'), sep = ":") %>%
  mutate(snp = ifelse(id == 1, proxy_snp, NA), 
         chr = str_replace(chr, 'chr', ""), 
         chr = as.numeric(chr), 
         pos = as.numeric(pos)) %>%
  fill(snp, .direction = 'down') %>%
  relocate(snp, .before = proxy_snp) %>%
  dplyr::select(-id, -func, -FORGEdb, -RegulomeDB) %>%
  filter(rsq >= 0.8)

# Munge proxy snp and outcome data 
proxy_outcome <- left_join(
  proxy_snps, exposure2_dat, by = c("proxy_snp" = "SNP")
) %>%
  separate(correlated_alleles, c("target_a1.outcome", "proxy_a1.outcome", 
                                 "target_a2.outcome", "proxy_a2.outcome"), sep = ",|=") %>%
  filter(!is.na(chr.outcome)) %>%
  arrange(snp, -rsq, abs(distance)) %>%
  group_by(snp) %>%
  slice(1) %>%
  ungroup() %>%
  mutate(
     proxy.outcome = TRUE,
     target_snp.outcome = snp,
     proxy_snp.outcome = proxy_snp, 
  ) %>% 
  mutate(
       new_effect_allele.outcome = case_when(
        proxy_a1.outcome == effect_allele.outcome & proxy_a2.outcome == other_allele.outcome ~ target_a1.outcome,
        proxy_a2.outcome == effect_allele.outcome & proxy_a1.outcome == other_allele.outcome ~ target_a2.outcome,
        TRUE ~ NA_character_
     ), 
      new_other_allele.outcome = case_when(
        proxy_a1.outcome == effect_allele.outcome & proxy_a2.outcome == other_allele.outcome ~ target_a2.outcome,
        proxy_a2.outcome == effect_allele.outcome & proxy_a1.outcome == other_allele.outcome ~ target_a1.outcome,
        TRUE ~ NA_character_
     ), 
     effect_allele.outcome = new_effect_allele.outcome, 
     other_allele.outcome = new_other_allele.outcome
  ) %>%
  dplyr::select(-proxy_snp, -chr, -pos, -alleles, -maf, -distance, -rsq, -dprime,  
         -new_effect_allele.outcome, -new_other_allele.outcome) %>%
  relocate(target_a1.outcome, proxy_a1.outcome, target_a2.outcome, proxy_a2.outcome, .after = proxy_snp.outcome) %>%
  rename(SNP = snp) %>%
  relocate(SNP, .after = samplesize.outcome)


# Merge outcome and proxy outcomes
longchamps_plate_dat_proxy <- bind_rows(
  outcome_clump, proxy_outcome
) %>% 
  arrange(chr.outcome, pos.outcome)

```

# 2nd: Exposure 1 and 2 harmonization. 

``` {r mtDNAcn_clump_longchamps_harmony}

#Temporary name change of exposures for harmonzation.

exposure1_dat <- longchamps_plate_dat_proxy # Clumped and LD proxy dataset.

exposure1 <- Format_mtDNAcn_longchamps # Exposure1  not yet clumped



mvmr_snps <- bind_rows(
  select(exposure1_dat, SNP), 
  select(exposure2_dat, SNP), 
) %>%
  distinct(SNP) %>%
  pull(SNP)
comb_exp <- bind_rows(
  exposure1 %>% filter(SNP %in% mvmr_snps),
  exposure2 %>% filter(SNP %in% mvmr_snps),
)
comb_exp_out <- bind_rows(
  exposure1 %>% filter(SNP %in% mvmr_snps) %>% 
    rename(pval = pval.exposure, trait = exposure, chr = chr.exposure, pos = pos.exposure), 
  exposure2 %>% filter(SNP %in% mvmr_snps) %>% 
    rename(pval = pval.exposure, trait = exposure, chr = chr.exposure, pos = pos.exposure), 
  outcome_AD_Format_kunkle %>% filter(SNP %in% mvmr_snps) %>% 
    rename(pval = pval.outcome, trait = outcome, chr = chr.outcome, pos = pos.outcome)
)
## How many SNPs are GWS across the exposures 
comb_exp_wide <- comb_exp_out %>%
   arrange(chr, pos) %>%
   select(SNP, trait, pval) %>%
   pivot_wider(names_from = trait, values_from = pval) %>%
   mutate(pval.exposure = pmin(mtDNA_longchamps, Platlet_cn, na.rm = T))
comb_exp_wide %>% count(mtDNA_longchamps < 5e-8, Platlet_cn < 5e-8)
comb_exp_wide %>% count(mtDNA_longchamps < 5e-8, Platlet_cn < 5e-8, kunkle < 5e-8)
## Do another round of LD clumping 
comb_exp_clump <- comb_exp %>% 
  clump_data(.,
  clump_kb = 10000,
  clump_r2 = 0.001,
  clump_p1 = 1,
  clump_p2 = 1,
  pop = "EUR"
)
comb_exp_clump %>% count(mtDNA_longchamps < 5e-8, Platlet_cn < 5e-8)
comb_exp <- comb_exp %>% filter(SNP %in% comb_exp_clump$SNP) 
```

# Step 3: LD Proxies with 2 exposures combined 
``` {r longchamps_kunkle_LDproxy}

# extract exposure SNPs present in outcome (outcome datasets,then exposure dataset)
outcome_clump <- semi_join(
  outcome_AD_Format_kunkle, comb_exp, by = "SNP"
)

# Exposure SNPs not present in outcome (exposure dataset, then outcome dataset)
exp_snps_wo <- anti_join(
  comb_exp, outcome_AD_Format_kunkle, by = "SNP"
)


# Use LDLinkR to identify proxy snps based on SNP dataset above
# that presents SNPs with no corresponding match in the other.
LDproxy_batch(exp_snps_wo$SNP, 
        pop = "CEU", 
        r2d = "r2", 
        token = 'a6deee62cc4a', 
        append = TRUE,
        genome_build = "grch37")
system("mv combined_query_snp_list_grch37.txt data/exposure_outcome_proxy_snps_longchamps_plate_kunkle.txt")


# Munge proxy snp file
proxy_snps <- read_tsv("data/exposure_outcome_proxy_snps_longchamps_plate_kunkle.txt", skip = 1, col_names = F) %>%
  rename(id = X1, func = X2, proxy_snp = X3, coord = X4, alleles = X5, maf = X6, 
         distance = X7, dprime = X8, rsq = X9, correlated_alleles = X10, FORGEdb = X11, RegulomeDB = X12) %>%
  separate(coord, c('chr', 'pos'), sep = ":") %>%
  mutate(snp = ifelse(id == 1, proxy_snp, NA), 
         chr = str_replace(chr, 'chr', ""), 
         chr = as.numeric(chr), 
         pos = as.numeric(pos)) %>%
  fill(snp, .direction = 'down') %>%
  relocate(snp, .before = proxy_snp) %>%
  dplyr::select(-id, -func, -FORGEdb, -RegulomeDB) %>%
  filter(rsq >= 0.8)

# Munge proxy snp and outcome data 
proxy_outcome <- left_join(
  proxy_snps, outcome_AD_Format_kunkle, by = c("proxy_snp" = "SNP")
) %>%
  separate(correlated_alleles, c("target_a1.outcome", "proxy_a1.outcome", 
                                 "target_a2.outcome", "proxy_a2.outcome"), sep = ",|=") %>%
  filter(!is.na(chr.outcome)) %>%
  arrange(snp, -rsq, abs(distance)) %>%
  group_by(snp) %>%
  slice(1) %>%
  ungroup() %>%
  mutate(
     proxy.outcome = TRUE,
     target_snp.outcome = snp,
     proxy_snp.outcome = proxy_snp, 
  ) %>% 
  mutate(
       new_effect_allele.outcome = case_when(
        proxy_a1.outcome == effect_allele.outcome & proxy_a2.outcome == other_allele.outcome ~ target_a1.outcome,
        proxy_a2.outcome == effect_allele.outcome & proxy_a1.outcome == other_allele.outcome ~ target_a2.outcome,
        TRUE ~ NA_character_
     ), 
      new_other_allele.outcome = case_when(
        proxy_a1.outcome == effect_allele.outcome & proxy_a2.outcome == other_allele.outcome ~ target_a2.outcome,
        proxy_a2.outcome == effect_allele.outcome & proxy_a1.outcome == other_allele.outcome ~ target_a1.outcome,
        TRUE ~ NA_character_
     ), 
     effect_allele.outcome = new_effect_allele.outcome, 
     other_allele.outcome = new_other_allele.outcome
  ) %>%
  dplyr::select(-proxy_snp, -chr, -pos, -alleles, -maf, -distance, -rsq, -dprime,  
         -new_effect_allele.outcome, -new_other_allele.outcome) %>%
  relocate(target_a1.outcome, proxy_a1.outcome, target_a2.outcome, proxy_a2.outcome, .after = proxy_snp.outcome) %>%
  rename(SNP = snp) %>%
  relocate(SNP, .after = samplesize.outcome)


# Merge outcome and proxy outcomes
longchamps_kunkle_dat_proxy <- bind_rows(
  outcome_clump, proxy_outcome
) %>% 
  arrange(chr.outcome, pos.outcome)

outcome_dat <- longchamps_kunkle_dat_proxy

```


## Venn diagram of GWAS SNPS 
```{r VennDiagram}
library(nVennR)
## GWS SNPs
set1 <- list(pull(exposure1_dat, SNP))
set2 <- list(pull(exposure2_dat, SNP))
myNV <- plotVenn(list(set1, set2), sNames=c("mtDNA_longchamps", "Platlet_cn"), 
                 showPlot = F)
showSVG(myNV, opacity=0.2, outFile = "results/plots/mvmr_gws_venn.svg")
## All SNPs 
set1 <- list(exposure1 %>% filter(SNP %in% mvmr_snps, pval.exposure < 5e-8) %>% pull(SNP))
set2 <- list(exposure2 %>% filter(SNP %in% mvmr_snps, pval.exposure < 5e-8) %>% pull(SNP))
set3 <- list(outcome %>% filter(SNP %in% mvmr_snps) %>% pull(SNP))
myNV <- plotVenn(list(set1, set2, set3), sNames=c("mtDNA_longchamps", "Platlet_cn", "kunkle"), 
                 showPlot = F)
showSVG(myNV, opacity=0.2, outFile = "results/plots/mvmr_venn_longchamps_kunkle.svg")
```

## Harmonize Exposure1_Exposure2_Outcome

``` {r harm_exposures_outcome}

mvdat <- mv_harmonise_data(comb_exp, outcome_dat)

```

# Export Code
```{r export_harmonized_dat}
write_rds(mvdat, 'data/harmonized_mvmr_longchamps_platelet_kunkle.rds')
```

## Multivariate MR

```{r TwoSampleMR_MVMR}
## Multivariable MR
mvmr_res <- mv_multiple(mvdat, plots = T)
mvmr_res$result
```


``` {r export}
write_csv(mvmr_res$result, 'results/mvmr_res_longchamps_platelet_kunkle.csv')

```








## MVMR hagg_exposure2_kunkle

# Harmonization 

Involves two rounds of LD proxies, with 1 round of exposure 1 and exposure 2 harmonization
in between.
1st. LD Proxies exposure 1 with exposure 2 as "outcome"
2nd. Exposure 1 and 2 harmonization.
3rd. LD proxies exposure1_2 combined with outcome AD Kunkle

# Step 1: LD Proxies exposure 1 with exposure 2 as "outcome"
``` {r hagg_exposure2_LDproxy}

# extract exposure SNPs present in outcome (outcome datasets,then exposure dataset)
outcome_clump <- semi_join(
  exposure2_dat, mtDNAcn_clump_hagg, by = "SNP"
)

# Exposure SNPs not present in outcome (exposure dataset, then outcome dataset)
exp_snps_wo <- anti_join(
  mtDNAcn_clump_hagg, exposure2_dat, by = "SNP"
)

# My LD_Proxy code 7dbc1afe7454



# Use LDLinkR to identify proxy snps based on SNP dataset above
# that presents SNPs with no corresponding match in the other.
LDproxy_batch(exp_snps_wo$SNP, 
        pop = "CEU", 
        r2d = "r2", 
        token = '7dbc1afe7454', 
        append = TRUE,
        genome_build = "grch37")
system("mv combined_query_snp_list_grch37.txt data/exposure_outcome_proxy_snps_hagg_plate.txt")


# Munge proxy snp file
proxy_snps <- read_tsv("data/exposure_outcome_proxy_snps_hagg_plate.txt", skip = 1, col_names = F) %>%
  rename(id = X1, func = X2, proxy_snp = X3, coord = X4, alleles = X5, maf = X6, 
         distance = X7, dprime = X8, rsq = X9, correlated_alleles = X10, FORGEdb = X11, RegulomeDB = X12) %>%
  separate(coord, c('chr', 'pos'), sep = ":") %>%
  mutate(snp = ifelse(id == 1, proxy_snp, NA), 
         chr = str_replace(chr, 'chr', ""), 
         chr = as.numeric(chr), 
         pos = as.numeric(pos)) %>%
  fill(snp, .direction = 'down') %>%
  relocate(snp, .before = proxy_snp) %>%
  dplyr::select(-id, -func, -FORGEdb, -RegulomeDB) %>%
  filter(rsq >= 0.8)

# Munge proxy snp and outcome data 
proxy_outcome <- left_join(
  proxy_snps, exposure2_dat, by = c("proxy_snp" = "SNP")
) %>%
  separate(correlated_alleles, c("target_a1.outcome", "proxy_a1.outcome", 
                                 "target_a2.outcome", "proxy_a2.outcome"), sep = ",|=") %>%
  filter(!is.na(chr.outcome)) %>%
  arrange(snp, -rsq, abs(distance)) %>%
  group_by(snp) %>%
  slice(1) %>%
  ungroup() %>%
  mutate(
     proxy.outcome = TRUE,
     target_snp.outcome = snp,
     proxy_snp.outcome = proxy_snp, 
  ) %>% 
  mutate(
       new_effect_allele.outcome = case_when(
        proxy_a1.outcome == effect_allele.outcome & proxy_a2.outcome == other_allele.outcome ~ target_a1.outcome,
        proxy_a2.outcome == effect_allele.outcome & proxy_a1.outcome == other_allele.outcome ~ target_a2.outcome,
        TRUE ~ NA_character_
     ), 
      new_other_allele.outcome = case_when(
        proxy_a1.outcome == effect_allele.outcome & proxy_a2.outcome == other_allele.outcome ~ target_a2.outcome,
        proxy_a2.outcome == effect_allele.outcome & proxy_a1.outcome == other_allele.outcome ~ target_a1.outcome,
        TRUE ~ NA_character_
     ), 
     effect_allele.outcome = new_effect_allele.outcome, 
     other_allele.outcome = new_other_allele.outcome
  ) %>%
  dplyr::select(-proxy_snp, -chr, -pos, -alleles, -maf, -distance, -rsq, -dprime,  
         -new_effect_allele.outcome, -new_other_allele.outcome) %>%
  relocate(target_a1.outcome, proxy_a1.outcome, target_a2.outcome, proxy_a2.outcome, .after = proxy_snp.outcome) %>%
  rename(SNP = snp) %>%
  relocate(SNP, .after = samplesize.outcome)


# Merge outcome and proxy outcomes
hagg_plate_dat_proxy <- bind_rows(
  outcome_clump, proxy_outcome
) %>% 
  arrange(chr.outcome, pos.outcome)

```

# 2nd: Exposure 1 and 2 harmonization. 

``` {r mtDNAcn_clump_hagg_harmony}

#Temporary name change of exposures for harmonzation.

exposure1_dat <- hagg_plate_dat_proxy # Clumped and LD proxy dataset.

exposure1 <- Format_mtDNAcn_hagg # Exposure1  not yet clumped



mvmr_snps <- bind_rows(
  select(exposure1_dat, SNP), 
  select(exposure2_dat, SNP), 
) %>%
  distinct(SNP) %>%
  pull(SNP)
comb_exp <- bind_rows(
  exposure1 %>% filter(SNP %in% mvmr_snps),
  exposure2 %>% filter(SNP %in% mvmr_snps),
)
comb_exp_out <- bind_rows(
  exposure1 %>% filter(SNP %in% mvmr_snps) %>% 
    rename(pval = pval.exposure, trait = exposure, chr = chr.exposure, pos = pos.exposure), 
  exposure2 %>% filter(SNP %in% mvmr_snps) %>% 
    rename(pval = pval.exposure, trait = exposure, chr = chr.exposure, pos = pos.exposure), 
  outcome_AD_Format_kunkle %>% filter(SNP %in% mvmr_snps) %>% 
    rename(pval = pval.outcome, trait = outcome, chr = chr.outcome, pos = pos.outcome)
)
## How many SNPs are GWS across the exposures 
comb_exp_wide <- comb_exp_out %>%
   arrange(chr, pos) %>%
   select(SNP, trait, pval) %>%
   pivot_wider(names_from = trait, values_from = pval) %>%
   mutate(pval.exposure = pmin(mtDNA_hagg, Platlet_cn, na.rm = T))
comb_exp_wide %>% count(mtDNA_hagg < 5e-8, Platlet_cn < 5e-8)
comb_exp_wide %>% count(mtDNA_hagg < 5e-8, Platlet_cn < 5e-8, kunkle < 5e-8)
## Do another round of LD clumping 
comb_exp_clump <- comb_exp %>% 
  clump_data(.,
  clump_kb = 10000,
  clump_r2 = 0.001,
  clump_p1 = 1,
  clump_p2 = 1,
  pop = "EUR"
)
comb_exp_clump %>% count(mtDNA_hagg < 5e-8, Platlet_cn < 5e-8)
comb_exp <- comb_exp %>% filter(SNP %in% comb_exp_clump$SNP) 
```

# Step 3: LD Proxies with 2 exposures combined 
``` {r hagg_kunkle_LDproxy}

# extract exposure SNPs present in outcome (outcome datasets,then exposure dataset)
outcome_clump <- semi_join(
  outcome_AD_Format_kunkle, comb_exp, by = "SNP"
)

# Exposure SNPs not present in outcome (exposure dataset, then outcome dataset)
exp_snps_wo <- anti_join(
  comb_exp, outcome_AD_Format_kunkle, by = "SNP"
)


# Use LDLinkR to identify proxy snps based on SNP dataset above
# that presents SNPs with no corresponding match in the other.
LDproxy_batch(exp_snps_wo$SNP, 
        pop = "CEU", 
        r2d = "r2", 
        token = 'a6deee62cc4a', 
        append = TRUE,
        genome_build = "grch37")
system("mv combined_query_snp_list_grch37.txt data/exposure_outcome_proxy_snps_hagg_plate_kunkle.txt")


# Munge proxy snp file
proxy_snps <- read_tsv("data/exposure_outcome_proxy_snps_hagg_plate_kunkle.txt", skip = 1, col_names = F) %>%
  rename(id = X1, func = X2, proxy_snp = X3, coord = X4, alleles = X5, maf = X6, 
         distance = X7, dprime = X8, rsq = X9, correlated_alleles = X10, FORGEdb = X11, RegulomeDB = X12) %>%
  separate(coord, c('chr', 'pos'), sep = ":") %>%
  mutate(snp = ifelse(id == 1, proxy_snp, NA), 
         chr = str_replace(chr, 'chr', ""), 
         chr = as.numeric(chr), 
         pos = as.numeric(pos)) %>%
  fill(snp, .direction = 'down') %>%
  relocate(snp, .before = proxy_snp) %>%
  dplyr::select(-id, -func, -FORGEdb, -RegulomeDB) %>%
  filter(rsq >= 0.8)

# Munge proxy snp and outcome data 
proxy_outcome <- left_join(
  proxy_snps, outcome_AD_Format_kunkle, by = c("proxy_snp" = "SNP")
) %>%
  separate(correlated_alleles, c("target_a1.outcome", "proxy_a1.outcome", 
                                 "target_a2.outcome", "proxy_a2.outcome"), sep = ",|=") %>%
  filter(!is.na(chr.outcome)) %>%
  arrange(snp, -rsq, abs(distance)) %>%
  group_by(snp) %>%
  slice(1) %>%
  ungroup() %>%
  mutate(
     proxy.outcome = TRUE,
     target_snp.outcome = snp,
     proxy_snp.outcome = proxy_snp, 
  ) %>% 
  mutate(
       new_effect_allele.outcome = case_when(
        proxy_a1.outcome == effect_allele.outcome & proxy_a2.outcome == other_allele.outcome ~ target_a1.outcome,
        proxy_a2.outcome == effect_allele.outcome & proxy_a1.outcome == other_allele.outcome ~ target_a2.outcome,
        TRUE ~ NA_character_
     ), 
      new_other_allele.outcome = case_when(
        proxy_a1.outcome == effect_allele.outcome & proxy_a2.outcome == other_allele.outcome ~ target_a2.outcome,
        proxy_a2.outcome == effect_allele.outcome & proxy_a1.outcome == other_allele.outcome ~ target_a1.outcome,
        TRUE ~ NA_character_
     ), 
     effect_allele.outcome = new_effect_allele.outcome, 
     other_allele.outcome = new_other_allele.outcome
  ) %>%
  dplyr::select(-proxy_snp, -chr, -pos, -alleles, -maf, -distance, -rsq, -dprime,  
         -new_effect_allele.outcome, -new_other_allele.outcome) %>%
  relocate(target_a1.outcome, proxy_a1.outcome, target_a2.outcome, proxy_a2.outcome, .after = proxy_snp.outcome) %>%
  rename(SNP = snp) %>%
  relocate(SNP, .after = samplesize.outcome)


# Merge outcome and proxy outcomes
hagg_kunkle_dat_proxy <- bind_rows(
  outcome_clump, proxy_outcome
) %>% 
  arrange(chr.outcome, pos.outcome)

outcome_dat <- hagg_kunkle_dat_proxy

```


## Venn diagram of GWAS SNPS (Brian, you must download package first.)
```{r VennDiagram}
library(nVennR)
## GWS SNPs
set1 <- list(pull(exposure1_dat, SNP))
set2 <- list(pull(exposure2_dat, SNP))
myNV <- plotVenn(list(set1, set2), sNames=c("mtDNA_hagg", "Platlet_cn"), 
                 showPlot = F)
showSVG(myNV, opacity=0.2, outFile = "results/plots/mvmr_gws_venn_hagg_kunkle.svg")
## All SNPs 
set1 <- list(exposure1 %>% filter(SNP %in% mvmr_snps, pval.exposure < 5e-8) %>% pull(SNP))
set2 <- list(exposure2 %>% filter(SNP %in% mvmr_snps, pval.exposure < 5e-8) %>% pull(SNP))
set3 <- list(outcome %>% filter(SNP %in% mvmr_snps) %>% pull(SNP))
myNV <- plotVenn(list(set1, set2, set3), sNames=c("mtDNA_hagg", "Platlet_cn", "kunkle"), 
                 showPlot = F)
showSVG(myNV, opacity=0.2, outFile = "results/plots/mvmr_venn_hagg_kunkle.svg")
```

## Harmonize Exposure1_Exposure2_Outcome

``` {r harm_exposures_outcome}

mvdat <- mv_harmonise_data(comb_exp, outcome_dat)

```

# Export Code
```{r export_harmonized_dat}
write_rds(mvdat, 'data/harmonized_mvmr_hagg_platelet_kunkle.rds')
```

## Multivariate MR

```{r TwoSampleMR_MVMR}
## Multivariable MR
mvmr_res <- mv_multiple(mvdat, plots = T)
mvmr_res$result
```


``` {r export}
write_csv(mvmr_res$result, 'results/mvmr_res_hagg_platelet_kunkle.csv')

```








## MVMR chong_exposure2_bellenguez

# Harmonization 

Involves two rounds of LD proxies, with 1 round of exposure 1 and exposure 2 harmonization
in between.
1st. LD Proxies exposure 1 with exposure 2 as "outcome"
2nd. Exposure 1 and 2 harmonization.
3rd. LD proxies exposure1_2 combined with outcome AD bellenguez



# 2nd: Exposure 1 and 2 harmonization. 

``` {r mtDNAcn_clump_chong_harmony}
# Reusing LD proxies of exposure1 on exposure 2 from 1st MVMR pair
#Temporary name change of exposures for harmonzation.

exposure1_dat <- chong_plate_dat_proxy # Clumped and LD proxy dataset.

exposure1 <- Format_mtDNAcn_chong # Exposure1  not yet clumped



mvmr_snps <- bind_rows(
  select(exposure1_dat, SNP), 
  select(exposure2_dat, SNP), 
) %>%
  distinct(SNP) %>%
  pull(SNP)
comb_exp <- bind_rows(
  exposure1 %>% filter(SNP %in% mvmr_snps),
  exposure2 %>% filter(SNP %in% mvmr_snps),
)
comb_exp_out <- bind_rows(
  exposure1 %>% filter(SNP %in% mvmr_snps) %>% 
    rename(pval = pval.exposure, trait = exposure, chr = chr.exposure, pos = pos.exposure), 
  exposure2 %>% filter(SNP %in% mvmr_snps) %>% 
    rename(pval = pval.exposure, trait = exposure, chr = chr.exposure, pos = pos.exposure), 
  outcome_AD_Format_bellenguez %>% filter(SNP %in% mvmr_snps) %>% 
    rename(pval = pval.outcome, trait = outcome, chr = chr.outcome, pos = pos.outcome)
)
## How many SNPs are GWS across the exposures 
comb_exp_wide <- comb_exp_out %>%
   arrange(chr, pos) %>%
   select(SNP, trait, pval) %>%
   pivot_wider(names_from = trait, values_from = pval) %>%
   mutate(pval.exposure = pmin(mtDNA_chong, Platlet_cn, na.rm = T))
comb_exp_wide %>% count(mtDNA_chong < 5e-8, Platlet_cn < 5e-8)
comb_exp_wide %>% count(mtDNA_chong < 5e-8, Platlet_cn < 5e-8, bellenguez < 5e-8)
## Do another round of LD clumping 
comb_exp_clump <- comb_exp %>% 
  clump_data(.,
  clump_kb = 10000,
  clump_r2 = 0.001,
  clump_p1 = 1,
  clump_p2 = 1,
  pop = "EUR"
)
comb_exp_clump %>% count(mtDNA_chong < 5e-8, Platlet_cn < 5e-8)
comb_exp <- comb_exp %>% filter(SNP %in% comb_exp_clump$SNP) 
```

# Step 3: LD Proxies with 2 exposures combined 
``` {r longchamps_kunkle_LDproxy}

# extract exposure SNPs present in outcome (outcome datasets,then exposure dataset)
outcome_clump <- semi_join(
  outcome_AD_Format_bellenguz, comb_exp, by = "SNP"
)

# Exposure SNPs not present in outcome (exposure dataset, then outcome dataset)
exp_snps_wo <- anti_join(
  comb_exp, outcome_AD_Format_bellenguz, by = "SNP"
)


# Use LDLinkR to identify proxy snps based on SNP dataset above
# that presents SNPs with no corresponding match in the other.
LDproxy_batch(exp_snps_wo$SNP, 
        pop = "CEU", 
        r2d = "r2", 
        token = 'a6deee62cc4a', 
        append = TRUE,
        genome_build = "grch37")
system("mv combined_query_snp_list_grch37.txt data/exposure_outcome_proxy_snps_chong_plate_bellenguez.txt")


# Munge proxy snp file
proxy_snps <- read_tsv("data/exposure_outcome_proxy_snps_chong_plate_bellenguez.txt", skip = 1, col_names = F) %>%
  rename(id = X1, func = X2, proxy_snp = X3, coord = X4, alleles = X5, maf = X6, 
         distance = X7, dprime = X8, rsq = X9, correlated_alleles = X10, FORGEdb = X11, RegulomeDB = X12) %>%
  separate(coord, c('chr', 'pos'), sep = ":") %>%
  mutate(snp = ifelse(id == 1, proxy_snp, NA), 
         chr = str_replace(chr, 'chr', ""), 
         chr = as.numeric(chr), 
         pos = as.numeric(pos)) %>%
  fill(snp, .direction = 'down') %>%
  relocate(snp, .before = proxy_snp) %>%
  dplyr::select(-id, -func, -FORGEdb, -RegulomeDB) %>%
  filter(rsq >= 0.8)

# Munge proxy snp and outcome data 
proxy_outcome <- left_join(
  proxy_snps, outcome_AD_Format_bellenguez, by = c("proxy_snp" = "SNP")
) %>%
  separate(correlated_alleles, c("target_a1.outcome", "proxy_a1.outcome", 
                                 "target_a2.outcome", "proxy_a2.outcome"), sep = ",|=") %>%
  filter(!is.na(chr.outcome)) %>%
  arrange(snp, -rsq, abs(distance)) %>%
  group_by(snp) %>%
  slice(1) %>%
  ungroup() %>%
  mutate(
     proxy.outcome = TRUE,
     target_snp.outcome = snp,
     proxy_snp.outcome = proxy_snp, 
  ) %>% 
  mutate(
       new_effect_allele.outcome = case_when(
        proxy_a1.outcome == effect_allele.outcome & proxy_a2.outcome == other_allele.outcome ~ target_a1.outcome,
        proxy_a2.outcome == effect_allele.outcome & proxy_a1.outcome == other_allele.outcome ~ target_a2.outcome,
        TRUE ~ NA_character_
     ), 
      new_other_allele.outcome = case_when(
        proxy_a1.outcome == effect_allele.outcome & proxy_a2.outcome == other_allele.outcome ~ target_a2.outcome,
        proxy_a2.outcome == effect_allele.outcome & proxy_a1.outcome == other_allele.outcome ~ target_a1.outcome,
        TRUE ~ NA_character_
     ), 
     effect_allele.outcome = new_effect_allele.outcome, 
     other_allele.outcome = new_other_allele.outcome
  ) %>%
  dplyr::select(-proxy_snp, -chr, -pos, -alleles, -maf, -distance, -rsq, -dprime,  
         -new_effect_allele.outcome, -new_other_allele.outcome) %>%
  relocate(target_a1.outcome, proxy_a1.outcome, target_a2.outcome, proxy_a2.outcome, .after = proxy_snp.outcome) %>%
  rename(SNP = snp) %>%
  relocate(SNP, .after = samplesize.outcome)


# Merge outcome and proxy outcomes
chong_bellenguez_dat_proxy <- bind_rows(
  outcome_clump, proxy_outcome
) %>% 
  arrange(chr.outcome, pos.outcome)

outcome_dat <- chong_bellenguez_dat_proxy

```


## Venn diagram of GWAS SNPS (Brian, you must download package first.)
```{r VennDiagram}
library(nVennR)
## GWS SNPs
set1 <- list(pull(exposure1_dat, SNP))
set2 <- list(pull(exposure2_dat, SNP))
myNV <- plotVenn(list(set1, set2), sNames=c("mtDNA_chong", "Platlet_cn"), 
                 showPlot = F)
showSVG(myNV, opacity=0.2, outFile = "results/plots/mvmr_gws_venn_chong_bellenguez.svg")
## All SNPs 
set1 <- list(exposure1 %>% filter(SNP %in% mvmr_snps, pval.exposure < 5e-8) %>% pull(SNP))
set2 <- list(exposure2 %>% filter(SNP %in% mvmr_snps, pval.exposure < 5e-8) %>% pull(SNP))
set3 <- list(outcome %>% filter(SNP %in% mvmr_snps) %>% pull(SNP))
myNV <- plotVenn(list(set1, set2, set3), sNames=c("mtDNA_hagg", "Platlet_cn", "bellenguez"), 
                 showPlot = F)
showSVG(myNV, opacity=0.2, outFile = "results/plots/mvmr_venn_chong_bellenguez.svg")
```

## Harmonize Exposure1_Exposure2_Outcome

``` {r harm_exposures_outcome}

mvdat <- mv_harmonise_data(comb_exp, outcome_dat)

```

# Export Code
```{r export_harmonized_dat}
write_rds(mvdat, 'data/harmonized_mvmr_chong_platelet_bellenguez.rds')
```

## Multivariate MR

```{r TwoSampleMR_MVMR}
## Multivariable MR
mvmr_res <- mv_multiple(mvdat, plots = T)
mvmr_res$result
```


``` {r export}
write_csv(mvmr_res$result, 'results/mvmr_res_chong_platelet_bellenguez.csv')

```









## MVMR longchamps_exposure2_bellenguez

# Harmonization 

Involves two rounds of LD proxies, with 1 round of exposure 1 and exposure 2 harmonization
in between.
1st. LD Proxies exposure 1 with exposure 2 as "outcome"
2nd. Exposure 1 and 2 harmonization.
3rd. LD proxies exposure1_2 combined with outcome AD bellenguez



# 2nd: Exposure 1 and 2 harmonization. 

``` {r mtDNAcn_clump_longchamps_harmony}
# Reusing LD proxies of exposure1 on exposure 2 from 1st MVMR pair
#Temporary name change of exposures for harmonzation.

exposure1_dat <- longchamps_plate_dat_proxy # Clumped and LD proxy dataset.

exposure1 <- Format_mtDNAcn_longchamps # Exposure1  not yet clumped



mvmr_snps <- bind_rows(
  select(exposure1_dat, SNP), 
  select(exposure2_dat, SNP), 
) %>%
  distinct(SNP) %>%
  pull(SNP)
comb_exp <- bind_rows(
  exposure1 %>% filter(SNP %in% mvmr_snps),
  exposure2 %>% filter(SNP %in% mvmr_snps),
)
comb_exp_out <- bind_rows(
  exposure1 %>% filter(SNP %in% mvmr_snps) %>% 
    rename(pval = pval.exposure, trait = exposure, chr = chr.exposure, pos = pos.exposure), 
  exposure2 %>% filter(SNP %in% mvmr_snps) %>% 
    rename(pval = pval.exposure, trait = exposure, chr = chr.exposure, pos = pos.exposure), 
  outcome_AD_Format_bellenguez %>% filter(SNP %in% mvmr_snps) %>% 
    rename(pval = pval.outcome, trait = outcome, chr = chr.outcome, pos = pos.outcome)
)
## How many SNPs are GWS across the exposures 
comb_exp_wide <- comb_exp_out %>%
   arrange(chr, pos) %>%
   select(SNP, trait, pval) %>%
   pivot_wider(names_from = trait, values_from = pval) %>%
   mutate(pval.exposure = pmin(mtDNA_longchamps, Platlet_cn, na.rm = T))
comb_exp_wide %>% count(mtDNA_longchamps < 5e-8, Platlet_cn < 5e-8)
comb_exp_wide %>% count(mtDNA_longchamps < 5e-8, Platlet_cn < 5e-8, bellenguez < 5e-8)
## Do another round of LD clumping 
comb_exp_clump <- comb_exp %>% 
  clump_data(.,
  clump_kb = 10000,
  clump_r2 = 0.001,
  clump_p1 = 1,
  clump_p2 = 1,
  pop = "EUR"
)
comb_exp_clump %>% count(mtDNA_longchamps < 5e-8, Platlet_cn < 5e-8)
comb_exp <- comb_exp %>% filter(SNP %in% comb_exp_clump$SNP) 
```

# Step 3: LD Proxies with 2 exposures combined 
``` {r longchamps_bellenguez_LDproxy}

# extract exposure SNPs present in outcome (outcome datasets,then exposure dataset)
outcome_clump <- semi_join(
  outcome_AD_Format_bellenguez, comb_exp, by = "SNP"
)

# Exposure SNPs not present in outcome (exposure dataset, then outcome dataset)
exp_snps_wo <- anti_join(
  comb_exp, outcome_AD_Format_bellenguez, by = "SNP"
)


# Use LDLinkR to identify proxy snps based on SNP dataset above
# that presents SNPs with no corresponding match in the other.
LDproxy_batch(exp_snps_wo$SNP, 
        pop = "CEU", 
        r2d = "r2", 
        token = 'a6deee62cc4a', 
        append = TRUE,
        genome_build = "grch37")
system("mv combined_query_snp_list_grch37.txt data/exposure_outcome_proxy_snps_longchamps_plate_bellenguez.txt")


# Munge proxy snp file
proxy_snps <- read_tsv("data/exposure_outcome_proxy_snps_longchamps_plate_bellenguez.txt", skip = 1, col_names = F) %>%
  rename(id = X1, func = X2, proxy_snp = X3, coord = X4, alleles = X5, maf = X6, 
         distance = X7, dprime = X8, rsq = X9, correlated_alleles = X10, FORGEdb = X11, RegulomeDB = X12) %>%
  separate(coord, c('chr', 'pos'), sep = ":") %>%
  mutate(snp = ifelse(id == 1, proxy_snp, NA), 
         chr = str_replace(chr, 'chr', ""), 
         chr = as.numeric(chr), 
         pos = as.numeric(pos)) %>%
  fill(snp, .direction = 'down') %>%
  relocate(snp, .before = proxy_snp) %>%
  dplyr::select(-id, -func, -FORGEdb, -RegulomeDB) %>%
  filter(rsq >= 0.8)

# Munge proxy snp and outcome data 
proxy_outcome <- left_join(
  proxy_snps, outcome_AD_Format_bellenguez, by = c("proxy_snp" = "SNP")
) %>%
  separate(correlated_alleles, c("target_a1.outcome", "proxy_a1.outcome", 
                                 "target_a2.outcome", "proxy_a2.outcome"), sep = ",|=") %>%
  filter(!is.na(chr.outcome)) %>%
  arrange(snp, -rsq, abs(distance)) %>%
  group_by(snp) %>%
  slice(1) %>%
  ungroup() %>%
  mutate(
     proxy.outcome = TRUE,
     target_snp.outcome = snp,
     proxy_snp.outcome = proxy_snp, 
  ) %>% 
  mutate(
       new_effect_allele.outcome = case_when(
        proxy_a1.outcome == effect_allele.outcome & proxy_a2.outcome == other_allele.outcome ~ target_a1.outcome,
        proxy_a2.outcome == effect_allele.outcome & proxy_a1.outcome == other_allele.outcome ~ target_a2.outcome,
        TRUE ~ NA_character_
     ), 
      new_other_allele.outcome = case_when(
        proxy_a1.outcome == effect_allele.outcome & proxy_a2.outcome == other_allele.outcome ~ target_a2.outcome,
        proxy_a2.outcome == effect_allele.outcome & proxy_a1.outcome == other_allele.outcome ~ target_a1.outcome,
        TRUE ~ NA_character_
     ), 
     effect_allele.outcome = new_effect_allele.outcome, 
     other_allele.outcome = new_other_allele.outcome
  ) %>%
  dplyr::select(-proxy_snp, -chr, -pos, -alleles, -maf, -distance, -rsq, -dprime,  
         -new_effect_allele.outcome, -new_other_allele.outcome) %>%
  relocate(target_a1.outcome, proxy_a1.outcome, target_a2.outcome, proxy_a2.outcome, .after = proxy_snp.outcome) %>%
  rename(SNP = snp) %>%
  relocate(SNP, .after = samplesize.outcome)


# Merge outcome and proxy outcomes
longchamps_bellenguez_dat_proxy <- bind_rows(
  outcome_clump, proxy_outcome
) %>% 
  arrange(chr.outcome, pos.outcome)

outcome_dat <- longchamps_bellenguez_dat_proxy

```


## Venn diagram of GWAS SNPS (Brian, you must download package first.)
```{r VennDiagram}
library(nVennR)
## GWS SNPs
set1 <- list(pull(exposure1_dat, SNP))
set2 <- list(pull(exposure2_dat, SNP))
myNV <- plotVenn(list(set1, set2), sNames=c("mtDNA_longchamps", "Platlet_cn"), 
                 showPlot = F)
showSVG(myNV, opacity=0.2, outFile = "results/plots/mvmr_gws_venn_longchamps_bellenguez.svg")
## All SNPs 
set1 <- list(exposure1 %>% filter(SNP %in% mvmr_snps, pval.exposure < 5e-8) %>% pull(SNP))
set2 <- list(exposure2 %>% filter(SNP %in% mvmr_snps, pval.exposure < 5e-8) %>% pull(SNP))
set3 <- list(outcome %>% filter(SNP %in% mvmr_snps) %>% pull(SNP))
myNV <- plotVenn(list(set1, set2, set3), sNames=c("mtDNA_longchamps", "Platlet_cn", "bellenguez"), 
                 showPlot = F)
showSVG(myNV, opacity=0.2, outFile = "results/plots/mvmr_venn_longchamps_bellenguez.svg")
```

## Harmonize Exposure1_Exposure2_Outcome

``` {r harm_exposures_outcome}

mvdat <- mv_harmonise_data(comb_exp, outcome_dat)

```

# Export Code
```{r export_harmonized_dat}
write_rds(mvdat, 'data/harmonized_mvmr_longchamps_platelet_bellenguez.rds')
```

## Multivariate MR

```{r TwoSampleMR_MVMR}
## Multivariable MR
mvmr_res <- mv_multiple(mvdat, plots = T)
mvmr_res$result
```


``` {r export}
write_csv(mvmr_res$result, 'results/mvmr_res_longchamps_platelet_bellenguez.csv')

```











## MVMR hagg_exposure2_bellenguez

# Harmonization 

Involves two rounds of LD proxies, with 1 round of exposure 1 and exposure 2 harmonization
in between.
1st. LD Proxies exposure 1 with exposure 2 as "outcome"
2nd. Exposure 1 and 2 harmonization.
3rd. LD proxies exposure1_2 combined with outcome AD bellenguez



# 2nd: Exposure 1 and 2 harmonization. 

``` {r mtDNAcn_clump_hagg_harmony}
# Reusing LD proxies of exposure1 on exposure 2 from 1st MVMR pair
#Temporary name change of exposures for harmonzation.

exposure1_dat <- hagg_plate_dat_proxy # Clumped and LD proxy dataset.

exposure1 <- Format_mtDNAcn_hagg # Exposure1  not yet clumped



mvmr_snps <- bind_rows(
  select(exposure1_dat, SNP), 
  select(exposure2_dat, SNP), 
) %>%
  distinct(SNP) %>%
  pull(SNP)
comb_exp <- bind_rows(
  exposure1 %>% filter(SNP %in% mvmr_snps),
  exposure2 %>% filter(SNP %in% mvmr_snps),
)
comb_exp_out <- bind_rows(
  exposure1 %>% filter(SNP %in% mvmr_snps) %>% 
    rename(pval = pval.exposure, trait = exposure, chr = chr.exposure, pos = pos.exposure), 
  exposure2 %>% filter(SNP %in% mvmr_snps) %>% 
    rename(pval = pval.exposure, trait = exposure, chr = chr.exposure, pos = pos.exposure), 
  outcome_AD_Format_bellenguez %>% filter(SNP %in% mvmr_snps) %>% 
    rename(pval = pval.outcome, trait = outcome, chr = chr.outcome, pos = pos.outcome)
)
## How many SNPs are GWS across the exposures 
comb_exp_wide <- comb_exp_out %>%
   arrange(chr, pos) %>%
   select(SNP, trait, pval) %>%
   pivot_wider(names_from = trait, values_from = pval) %>%
   mutate(pval.exposure = pmin(mtDNA_hagg, Platlet_cn, na.rm = T))
comb_exp_wide %>% count(mtDNA_hagg < 5e-8, Platlet_cn < 5e-8)
comb_exp_wide %>% count(mtDNA_hagg < 5e-8, Platlet_cn < 5e-8, bellenguez < 5e-8)
## Do another round of LD clumping 
comb_exp_clump <- comb_exp %>% 
  clump_data(.,
  clump_kb = 10000,
  clump_r2 = 0.001,
  clump_p1 = 1,
  clump_p2 = 1,
  pop = "EUR"
)
comb_exp_clump %>% count(mtDNA_hagg < 5e-8, Platlet_cn < 5e-8)
comb_exp <- comb_exp %>% filter(SNP %in% comb_exp_clump$SNP) 
```

# Step 3: LD Proxies with 2 exposures combined 
``` {r longchamps_bellenguez_LDproxy}

# extract exposure SNPs present in outcome (outcome datasets,then exposure dataset)
outcome_clump <- semi_join(
  outcome_AD_Format_bellenguez, comb_exp, by = "SNP"
)

# Exposure SNPs not present in outcome (exposure dataset, then outcome dataset)
exp_snps_wo <- anti_join(
  comb_exp, outcome_AD_Format_bellenguez, by = "SNP"
)


# Use LDLinkR to identify proxy snps based on SNP dataset above
# that presents SNPs with no corresponding match in the other.
LDproxy_batch(exp_snps_wo$SNP, 
        pop = "CEU", 
        r2d = "r2", 
        token = 'a6deee62cc4a', 
        append = TRUE,
        genome_build = "grch37")
system("mv combined_query_snp_list_grch37.txt data/exposure_outcome_proxy_snps_hagg_plate_bellenguez.txt")


# Munge proxy snp file
proxy_snps <- read_tsv("data/exposure_outcome_proxy_snps_hagg_plate_bellenguez.txt", skip = 1, col_names = F) %>%
  rename(id = X1, func = X2, proxy_snp = X3, coord = X4, alleles = X5, maf = X6, 
         distance = X7, dprime = X8, rsq = X9, correlated_alleles = X10, FORGEdb = X11, RegulomeDB = X12) %>%
  separate(coord, c('chr', 'pos'), sep = ":") %>%
  mutate(snp = ifelse(id == 1, proxy_snp, NA), 
         chr = str_replace(chr, 'chr', ""), 
         chr = as.numeric(chr), 
         pos = as.numeric(pos)) %>%
  fill(snp, .direction = 'down') %>%
  relocate(snp, .before = proxy_snp) %>%
  dplyr::select(-id, -func, -FORGEdb, -RegulomeDB) %>%
  filter(rsq >= 0.8)

# Munge proxy snp and outcome data 
proxy_outcome <- left_join(
  proxy_snps, outcome_AD_Format_bellenguez, by = c("proxy_snp" = "SNP")
) %>%
  separate(correlated_alleles, c("target_a1.outcome", "proxy_a1.outcome", 
                                 "target_a2.outcome", "proxy_a2.outcome"), sep = ",|=") %>%
  filter(!is.na(chr.outcome)) %>%
  arrange(snp, -rsq, abs(distance)) %>%
  group_by(snp) %>%
  slice(1) %>%
  ungroup() %>%
  mutate(
     proxy.outcome = TRUE,
     target_snp.outcome = snp,
     proxy_snp.outcome = proxy_snp, 
  ) %>% 
  mutate(
       new_effect_allele.outcome = case_when(
        proxy_a1.outcome == effect_allele.outcome & proxy_a2.outcome == other_allele.outcome ~ target_a1.outcome,
        proxy_a2.outcome == effect_allele.outcome & proxy_a1.outcome == other_allele.outcome ~ target_a2.outcome,
        TRUE ~ NA_character_
     ), 
      new_other_allele.outcome = case_when(
        proxy_a1.outcome == effect_allele.outcome & proxy_a2.outcome == other_allele.outcome ~ target_a2.outcome,
        proxy_a2.outcome == effect_allele.outcome & proxy_a1.outcome == other_allele.outcome ~ target_a1.outcome,
        TRUE ~ NA_character_
     ), 
     effect_allele.outcome = new_effect_allele.outcome, 
     other_allele.outcome = new_other_allele.outcome
  ) %>%
  dplyr::select(-proxy_snp, -chr, -pos, -alleles, -maf, -distance, -rsq, -dprime,  
         -new_effect_allele.outcome, -new_other_allele.outcome) %>%
  relocate(target_a1.outcome, proxy_a1.outcome, target_a2.outcome, proxy_a2.outcome, .after = proxy_snp.outcome) %>%
  rename(SNP = snp) %>%
  relocate(SNP, .after = samplesize.outcome)


# Merge outcome and proxy outcomes
hagg_bellenguez_dat_proxy <- bind_rows(
  outcome_clump, proxy_outcome
) %>% 
  arrange(chr.outcome, pos.outcome)

outcome_dat <- hagg_bellenguez_dat_proxy

```


## Venn diagram of GWAS SNPS (Brian, you must download package first.)
```{r VennDiagram}
library(nVennR)
## GWS SNPs
set1 <- list(pull(exposure1_dat, SNP))
set2 <- list(pull(exposure2_dat, SNP))
myNV <- plotVenn(list(set1, set2), sNames=c("mtDNA_hagg", "Platlet_cn"), 
                 showPlot = F)
showSVG(myNV, opacity=0.2, outFile = "results/plots/mvmr_gws_venn_hagg_bellenguez.svg")
## All SNPs 
set1 <- list(exposure1 %>% filter(SNP %in% mvmr_snps, pval.exposure < 5e-8) %>% pull(SNP))
set2 <- list(exposure2 %>% filter(SNP %in% mvmr_snps, pval.exposure < 5e-8) %>% pull(SNP))
set3 <- list(outcome %>% filter(SNP %in% mvmr_snps) %>% pull(SNP))
myNV <- plotVenn(list(set1, set2, set3), sNames=c("mtDNA_hagg", "Platlet_cn", "bellenguez"), 
                 showPlot = F)
showSVG(myNV, opacity=0.2, outFile = "results/plots/mvmr_venn_hagg_bellenguez.svg")
```

## Harmonize Exposure1_Exposure2_Outcome

``` {r harm_exposures_outcome}

mvdat <- mv_harmonise_data(comb_exp, outcome_dat)

```

# Export Code
```{r export_harmonized_dat}
write_rds(mvdat, 'data/harmonized_mvmr_hagg_platelet_bellenguez.rds')
```

## Multivariate MR

```{r TwoSampleMR_MVMR}
## Multivariable MR
mvmr_res <- mv_multiple(mvdat, plots = T)
mvmr_res$result
```


``` {r export}
write_csv(mvmr_res$result, 'results/mvmr_res_hagg_platelet_bellenguez.csv')

```

