---
title: "Chong LHC MR"
author: "Brian Alvarez"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

knitr::opts_knit$set(root.dir = 'C:/Users/brian/OneDrive/Documents/GitHub/Casual_Estimate_mtDNAcn_on_Alzheimer')

 
```


## Libraries

Loads the appropriate packages for MR analysis

``` {r eval=FALSE}
library(tidyverse) # Data wrangling 

library (usethis)

library(devtools)

library(ggman)

library(TwoSampleMR) # MR 

library(R.utils) #To read in Gz data? 
#For the read_outcome_data step

library(RadialMR) # Radial MR sensitivity analysis 

library(phenoscanner)

library(LDlinkR) # LD and proxy snps

library(lhcMR)

library(janitor)
```


## Importing and formatting Exposure

This step imports our GWAS Exposure data sets and formats them up for MR analysis


```{r eval=FALSE}
# Define column types for summary statistics GWAS
coltypes = cols(
  ID = col_character(),
  CHROM = col_double(),
  POS = col_double(),
  REF = col_character(),
  ALT = col_character(),
  AF = col_double(),
  TRAIT = col_character(),
  BETA = col_double(),
  SE = col_double(),
  Z = col_double(),
  P = col_double(),
  N = col_double(),
  OR = col_double(),
  OR_L95 = col_double(),
  OR_U95 = col_double(),
  DIR = col_character(),
  G1000_ID = col_character(),
  G1000_VARIANT = col_character(),
  DBSNP_ID = col_character(),
  DBSNP_VARIANT = col_character(),
  OLD_ID = col_character(),
  OLD_VARIANT = col_character()
)



```

## Step 1: Importing Exposure and Outcome Data
```{r ld}
## File paths needed for the analysis
LD.filepath = "Resources/LDscores_filtered.csv" # LD scores
rho.filepath = "Resources/LD_GM2_2prm.csv" # local/SNP-specific LD scores
ld = "Resources/eur_w_ld_chr/"
hm3 = "Resources/w_hm3.snplist"
```

## Exposure Import
Mitochondrial DNA count Number (X variable)

```{r read_exposure_Chong}

exposure_path = "Resources/Chong2022mtdnacn_eur.chrall.CPRA_b37.tsv.gz"
exposure_ss_Chong <- read_tsv(exposure_path, comment = "##", col_types = coltypes, 
                        col_select = c(DBSNP_ID, CHROM, POS, REF, ALT, AF, BETA, SE, Z, P, N)) %>%
  rename(SNP = DBSNP_ID)



```

```{r read_exposure_Longchamps}

Longchamps_path = "Resources/Longchamps2021mtdnacn.chrall.CPRA_b37.tsv.gz"
exposure_ss_Longchamps <- read_tsv(Longchamps_path, comment = "##", col_types = coltypes, 
                        col_select = c(DBSNP_ID, CHROM, POS, REF, ALT, AF, BETA, SE, Z, P, N)) %>%
  rename(SNP = DBSNP_ID)
```

```{r read_exposure_Hagg}

Hagg_path = "Resources/Hagg2020mtdnacn.chrall.CPRA_b37.tsv.gz"
exposure_ss_Hagg <- read_tsv(Hagg_path, comment = "##", col_types = coltypes, 
                        col_select = c(DBSNP_ID, CHROM, POS, REF, ALT, AF, BETA, SE, Z, P, N)) %>%
  rename(SNP = DBSNP_ID)
```

## Outcome Import 
Alzheimer's Disease (Y variable)

```{r read_outcome_Kunkle}
outcome_path = "Resources/Kunkle2019load_stage123.chrall.CPRA_b37.tsv.gz"
outcome_ss_Kunkle <- read_tsv(outcome_path, comment = "##",  col_types = coltypes, 
                        col_select = c(DBSNP_ID, CHROM, POS, REF, ALT, AF, BETA, SE, Z, P, N))  %>%
  rename(SNP = DBSNP_ID)
```

```{r read_outcome_Bellenguez}

Bellenguez_path = "Resources/Bellenguez2022load.chrall.CPRA_b37.tsv.gz"
outcome_ss_Bellenguez <- read_tsv(Bellenguez_path, comment = "##",  col_types = coltypes, 
                        col_select = c(DBSNP_ID, CHROM, POS, REF, ALT, AF, BETA, SE, Z, P, N))  %>%
  rename(SNP = DBSNP_ID)

```




## Merging Exposure and outcome datasets (X and Y) and perform LHC-MR

Will perform Multiple merging steps, as well as mulitple LHC-MR Analysis.
Warning! : be sure to move all LHC results to a different folder before performing
another LHC-MR paired analysis, as it will overwrite the previous ones results.
Please check your directory to see where the files were placed and place them 
in a new folder for safe keeping. 
Perform one LHC-MR at a time.
Running each LHC-MR pair will take approximately 6-12 hours depending on your machine
and the amount of cores specified in "ncores" (more cores speeds up the process).



## Chong-Kunkle-LHCMR

```{r merge_Chong_Kunkle}
## Step 1
trait.names=c("mtDNAcn_Apoe_retained","AD")
input.files = list(exposure_ss_Chong, outcome_ss_Kunkle)
df = merge_sumstats(input.files, trait.names, LD.filepath, rho.filepath)
```


## Step 2: Calculating smart starting points for the likelihood optimisation

```{r read_merge}
SP_list = calculate_SP(df,trait.names,run_ldsc=TRUE,run_MR=TRUE,hm3=hm3,ld="Resources/eur_w_ld_chr",nStep = 2,
                       SP_single=3,SP_pair=50,SNP_filter=10)
```

## Step 3: Running the likelihood optimisation to estimate the parameters, followed by a block-jackknife procedure to calculate parameter-SE

```{r read_merge_LHCMR}
## Step 3
res_Chong_Kunkle_apoe_keep  = lhc_mr(SP_list, trait.names, paral_method="lapply", nCores = 1, nBlock=200)
res_Chong_Kunkle_apoe_keep  %>% as_tibble() %>% write_csv(., "results/LHC_MR_Results_Chong_Kunkle/lhcmr_Chong_Kunkle_apoe_retained.csv")


view(res_Chong_Kunkle_apoe_keep)



```

## Chong-Bellenguez-LHCMR

```{r merge_Chong_Bellenguez}
## Step 1
trait.names=c("mtDNAcn_Apoe_retained_chong","AD_bellenguez")
input.files = list(exposure_ss_Chong, outcome_ss_Bellenguez)
df = merge_sumstats(input.files, trait.names, LD.filepath, rho.filepath)
```


## Step 2: Calculating smart starting points for the likelihood optimisation

```{r read_merge}
SP_list = calculate_SP(df,trait.names,run_ldsc=TRUE,run_MR=TRUE,hm3=hm3,ld="Resources/eur_w_ld_chr",nStep = 2,
                       SP_single=3,SP_pair=50,SNP_filter=10)
```

## Step 3: Running the likelihood optimisation to estimate the parameters, followed by a block-jackknife procedure to calculate parameter-SE

```{r read_merge_LHCMR}
## Step 3
res_Chong_Bellenguez_apoe_keep = lhc_mr(SP_list, trait.names, paral_method="lapply", nCores = 1, nBlock=200)
res_Chong_Bellenguez_apoe_keep %>% as_tibble() %>% write_csv(., "results/LHC_MR_Results_Chong_Bellenguez/lhcmr_Chong_bellenguez_apoe_retained.csv")


view(res_Chong_Bellenguez_apoe_keep)


```

## Longchamps-Bellenguez-LHCMR

```{r merge_longchamps_Bellenguez}
## Step 1
trait.names=c("mtDNAcn_Apoe_retained_longchamps","AD_bellenguez")
input.files = list(exposure_ss_Longchamps, outcome_ss_Bellenguez)
df = merge_sumstats(input.files, trait.names, LD.filepath, rho.filepath)
```


## Step 2: Calculating smart starting points for the likelihood optimisation

```{r read_merge}
SP_list = calculate_SP(df,trait.names,run_ldsc=TRUE,run_MR=TRUE,hm3=hm3,ld="Resources/eur_w_ld_chr",nStep = 2,
                       SP_single=3,SP_pair=50,SNP_filter=10)
```

## Step 3: Running the likelihood optimisation to estimate the parameters, followed by a block-jackknife procedure to calculate parameter-SE

```{r read_merge_LHCMR}
## Step 3
res_longchamps_Bellenguez_apoe_keep = lhc_mr(SP_list, trait.names, paral_method="lapply", nCores = 1, nBlock=200)
res_longchamps_Bellenguez_apoe_keep %>% as_tibble() %>% write_csv(., "results/LHC_MR_Results_Longchamps_Bellenguez/lhcmr_Longchamps_bellenguez_apoe_retained.csv")


view(res_longchamps_Bellenguez_apoe_keep)


```


## Hagg-Bellenguez-LHCMR


```{r merge_Hagg_Bellenguez}
## Step 1
trait.names=c("mtDNAcn_Apoe_retained_hagg","AD_bellenguez")
input.files = list(exposure_ss_Hagg, outcome_ss_Bellenguez)
df = merge_sumstats(input.files, trait.names, LD.filepath, rho.filepath)
```


## Step 2: Calculating smart starting points for the likelihood optimisation

```{r read_merge}
SP_list = calculate_SP(df,trait.names,run_ldsc=TRUE,run_MR=TRUE,hm3=hm3,ld="Resources/eur_w_ld_chr",nStep = 2,
                       SP_single=3,SP_pair=50,SNP_filter=10)
```

## Step 3: Running the likelihood optimisation to estimate the parameters, followed by a block-jackknife procedure to calculate parameter-SE

```{r read_merge_LHCMR}
## Step 3
res_Hagg_Bellenguez_apoe_keep = lhc_mr(SP_list, trait.names, paral_method="lapply", nCores = 1, nBlock=200)
res_Hagg_Bellenguez_apoe_keep %>% as_tibble() %>% write_csv(., "results/LHC_MR_Results_Hagg_Bellenguez/lhcmr_Hagg_bellenguez_apoe_retained.csv")


view(res_Hagg_Bellenguez_apoe_keep)


```