---
title: "Mendelian Randomization mtDNA-CN on Alzheimer's Disease "
author: "Brian Alvarez"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

knitr::opts_knit$set(root.dir = 'C:/Users/brian/OneDrive/Documents/GitHub/Casual_Estimate_mtDNAcn_on_Alzheimer')
```


## Libraries

Loads the appropriate packages for MR analysis

``` {r eval=FALSE}
library(tidyverse) # Data wrangling 

library (usethis)

library(devtools)

library(ggman)

library(TwoSampleMR) # MR 

library(R.utils) #For the read_outcome_data step

library(RadialMR) # Radial MR sensitivity analysis 

library(phenoscanner)

library(LDlinkR) # LD and proxy snps

library(lhcMR)

library(janitor)
```


## Importing and formatting Exposure

This step imports our GWAS Exposure data sets(mtDNAcn) and formats them up for MR analysis


```{r eval=FALSE}
# Define column types for summary statistics GWAS
coltypes = cols(
  ID = col_character(),
  CHROM = col_double(),
  POS = col_double(),
  REF = col_character(),
  ALT = col_character(),
  AF = col_double(),
  TRAIT = col_character(),
  BETA = col_double(),
  SE = col_double(),
  Z = col_double(),
  P = col_double(),
  N = col_double(),
  OR = col_double(),
  OR_L95 = col_double(),
  OR_U95 = col_double(),
  DIR = col_character(),
  G1000_ID = col_character(),
  G1000_VARIANT = col_character(),
  DBSNP_ID = col_character(),
  DBSNP_VARIANT = col_character(),
  OLD_ID = col_character(),
  OLD_VARIANT = col_character()
)

```


``` {r read_in_chong}

exposure_path_chong = "Resources/Chong2022mtdnacn_eur.chrall.CPRA_b37.tsv.gz"

Exposure_mtDNAcn_chong <- read_tsv(exposure_path_chong,comment="#")


# To format the GWAS results for MR analysis Exposure. 

# Rename right-side variables to match column names of your GWAS dataset 
Format_mtDNAcn_chong <- Exposure_mtDNAcn_chong %>% format_data(.,
  type = "exposure",
  snps = NULL,
  header = TRUE,
  phenotype_col = "TRAIT",
  snp_col = "DBSNP_ID",
  beta_col = "BETA",
  se_col = "SE",
  eaf_col = "AF",
  effect_allele_col = "REF",
  other_allele_col = "ALT",
  pval_col = "P",
  units_col = "units",
  ncase_col = "N_CASES",
  ncontrol_col = "N_CTRLS",
  samplesize_col = "N",
  gene_col = "gene",
  min_pval = 1e-200,
  z_col = "Z",
  info_col = "info",
  chr_col = "CHROM",
  pos_col = "POS",
  log_pval = FALSE
)  %>%
  as_tibble()


```


``` {r read_in_longchamps}

exposure_path_longchamps = "Resources/Longchamps2021mtdnacn.chrall.CPRA_b37.tsv.gz"

Exposure_mtDNAcn_longchamps <- read_tsv(exposure_path_longchamps,comment="#")


# To format the GWAS results for MR analysis Exposure. 
# Rename right-side variables to match column names of your GWAS dataset.
Format_mtDNAcn_longchamps <- Exposure_mtDNAcn_longchamps %>% format_data(.,
  type = "exposure",
  snps = NULL,
  header = TRUE,
  phenotype_col = "TRAIT",
  snp_col = "DBSNP_ID",
  beta_col = "BETA",
  se_col = "SE",
  eaf_col = "AF",
  effect_allele_col = "REF",
  other_allele_col = "ALT",
  pval_col = "P",
  units_col = "units",
  ncase_col = "N_CASES",
  ncontrol_col = "N_CTRLS",
  samplesize_col = "N",
  gene_col = "gene",
  min_pval = 1e-200,
  z_col = "Z",
  info_col = "info",
  chr_col = "CHROM",
  pos_col = "POS",
  log_pval = FALSE
)  %>%
  as_tibble()



```



``` {r read_in_hagg}

exposure_path_hagg = "Resources/Hagg2020mtdnacn.chrall.CPRA_b37.tsv.gz"

Exposure_mtDNAcn_hagg <- read_tsv(exposure_path_hagg,comment="#")


# Format the GWAS results for MR analysis Exposure. 

# Rename right-side variables to match column names of your GWAS dataset.
Format_mtDNAcn_hagg <- Exposure_mtDNAcn_hagg %>% format_data(.,
  type = "exposure",
  snps = NULL,
  header = TRUE,
  phenotype_col = "TRAIT",
  snp_col = "DBSNP_ID",
  beta_col = "BETA",
  se_col = "SE",
  eaf_col = "AF",
  effect_allele_col = "REF",
  other_allele_col = "ALT",
  pval_col = "P",
  units_col = "units",
  ncase_col = "N_CASES",
  ncontrol_col = "N_CTRLS",
  samplesize_col = "N",
  gene_col = "gene",
  min_pval = 1e-200,
  z_col = "Z",
  info_col = "info",
  chr_col = "CHROM",
  pos_col = "POS",
  log_pval = FALSE
)  %>%
  as_tibble()

```


## Importing and formatting Outcome

This step imports our GWAS Outcome data sets and formats them up for MR analysis

``` {r read_in_Bellenguez}

outcome_path_bellenguez = "Resources/Bellenguez2022load.chrall.CPRA_b37.tsv.gz"

# Import Code for outcome GWAS 
OutcomeAD_bellenguez <- read_tsv(outcome_path_bellenguez,comment="#")

# Rename right-side variables to match column names of your GWAS dataset 

outcome_AD_Format_bellenguez <- OutcomeAD_bellenguez %>%
  format_data(.,
    type = "outcome",
    snps = NULL,
    header = TRUE,
    phenotype_col = "Phenotype",
    snp_col = "DBSNP_ID",
    beta_col = "BETA",
    se_col = "SE",
    eaf_col = "AF",
    effect_allele_col = "ALT",
    other_allele_col = "REF",
    pval_col = "P",
    samplesize_col = "N",
    z_col = "Z",
    chr_col = "CHROM",
    pos_col = "POS",
    log_pval = FALSE
) %>%
  as_tibble()

```



``` {r read_in_Kunkle}


outcome_path_kunkle = "Resources/Kunkle2019load_stage123.chrall.CPRA_b37.tsv.gz"

# Import Code for outcome GWAS 
OutcomeAD_kunkle <- read_tsv(outcome_path_kunkle,comment="#")

# Rename right-side variables to match column names of your GWAS dataset 

outcome_AD_Format_kunkle <- OutcomeAD_kunkle %>%
  format_data(.,
    type = "outcome",
    snps = NULL,
    header = TRUE,
    phenotype_col = "Phenotype",
    snp_col = "DBSNP_ID",
    beta_col = "BETA",
    se_col = "SE",
    eaf_col = "AF",
    effect_allele_col = "ALT",
    other_allele_col = "REF",
    pval_col = "P",
    samplesize_col = "N",
    z_col = "Z",
    chr_col = "CHROM",
    pos_col = "POS",
    log_pval = FALSE
) %>%
  as_tibble()



```


## Clumping Step
This step involving clumping and pruning to eliminate High linkage 
disequilibrium SNP (SNPS correlated with other SNPs). Removing High linkage 
disequilibrium SNP's are crucial to avoid "double counting" SNP's and
biasing our MR estimate.
The dataset will also be filter to remove a portion of insignficant SNP's for 
computational speed improvements.

``` {r eval=FALSE}

#Extract Instrumental variables from the exposure.
# Note that the lower Clump
#Clump R^2 is, the less linkage disequilibrium snps will
#be included. 
# Filter allows for more time efficient clumping

# Filter Clumped SNPS to include only those SNPS whose P value is 
# Below 5e-8, Colname for P value is Pval.exposure in the filtered dataset.

mtDNAcn_filter_chong <- filter(Format_mtDNAcn_chong , pval.exposure < 5e-8)


mtDNAcn_clump_chong <- clump_data(mtDNAcn_filter_chong, clump_kb = 10000,
                            clump_r2 = 0.001,
                            clump_p1 = 1,
                            clump_p2 = 1,
                            pop = "EUR"
)%>% 
  as_tibble()


mtDNAcn_filter_longchamps <- filter(Exposure_mtDNAcn_longchamps , pval.exposure < 5e-8)


mtDNAcn_Clump_chong <- clump_data(mtDNAcn_Format_chong, clump_kb = 10000,
                            clump_r2 = 0.001,
                            clump_p1 = 1,
                            clump_p2 = 1,
                            pop = "EUR"
)%>% 
  as_tibble()

# The filter statements makes our dataset small in 
#reference to a pval.
#We then use the clump_data command.
#Extract Instrumental variables from the exposure.
# Note that the lower R^2 is, the less linkage disequilibrium snps will
#be included. Use clump_data. Filter will take all
#Observations below p value of 1e-05, change to
#8e-08
# This filters all of the significant SNPS and leaves all non significant 
# SNPs from the analysis.






#For outcome data I believe you use this code. 
#It places the SNPS
#From our clumped and pruned dataset, and 
#references the outcome
#dataset to pull out those same SNPS.
#Makes sure to write in quotations the columns names for
#,"RDatasetsb/Kunkle2019load_stage123.chrall.CPRA_b37.tsv.gz"
#Put above name in file name statement.
#Put \t in "sep" statement.
#Change name from Cholesadcombine to Outcome_AD_2

# Added code for Phenotype name="TRAIT" 
# phenotype_col = "TRAIT"
# Did not work :( Must be missing something!

# Data set to be used in LD Proxy step
mtDNAcn_Clump_chong_data <- mtDNAcn_Clump_chong 
```


## LD Proxies
This attempts to retain any missing SNP's that are not both present in the 
exposure and outcome dataset by utilizing LD proxies in their place. This step will 
be done multiple times for each exposure > outcome pair.

``` {r }


```


## Apoe Region removed step after prelimnary MR analysis (make sure to remove outliers before this analysis)

