---
title: "Mendelian Randomization mtDNA-CN on Alzheimer's Disease "
author: "Brian Alvarez"
date: "`r Sys.Date()`"
output: html_document
---

## Libraries

Loads the appropriate packages for MR analysis

``` {r eval=FALSE}
library(tidyverse) # Data wrangling 

library (usethis)

library(devtools)

library(ggman)

library(TwoSampleMR) # MR 

library(R.utils) #For the read_outcome_data step

library(RadialMR) # Radial MR sensitivity analysis 

library(phenoscanner)

library(LDlinkR) # LD and proxy snps

library(lhcMR)

library(janitor)

library(grid)

library(gridExtra)

library(cowplot)
```


## Importing and formatting Exposure

This step imports our GWAS Exposure data sets(mtDNAcn) and formats them up for MR analysis


```{r eval=FALSE}
# Define column types for summary statistics GWAS
coltypes = cols(
  ID = col_character(),
  CHROM = col_double(),
  POS = col_double(),
  REF = col_character(),
  ALT = col_character(),
  AF = col_double(),
  TRAIT = col_character(),
  BETA = col_double(),
  SE = col_double(),
  Z = col_double(),
  P = col_double(),
  N = col_double(),
  OR = col_double(),
  OR_L95 = col_double(),
  OR_U95 = col_double(),
  DIR = col_character(),
  G1000_ID = col_character(),
  G1000_VARIANT = col_character(),
  DBSNP_ID = col_character(),
  DBSNP_VARIANT = col_character(),
  OLD_ID = col_character(),
  OLD_VARIANT = col_character()
)

```


``` {r read_in_chong}

exposure_path_chong = "/wynton/group/andrews/data/summary_stats/results/Chong2022mtdnacn_eur.chrall.CPRA_b37.tsv"

mtDNAcn_chong <- read_tsv(exposure_path_chong,comment="#")


# To format the GWAS results for MR analysis Exposure...
# Rename right-side variables to match column names of your GWAS dataset 

Format_mtDNAcn_chong <- mtDNAcn_chong %>% format_data(.,
  type = "exposure",
  snps = NULL,
  header = TRUE,
  phenotype_col = "TRAIT",
  snp_col = "DBSNP_ID",
  beta_col = "BETA",
  se_col = "SE",
  eaf_col = "AF",
  effect_allele_col = "REF",
  other_allele_col = "ALT",
  pval_col = "P",
  units_col = "units",
  ncase_col = "N_CASES",
  ncontrol_col = "N_CTRLS",
  samplesize_col = "N",
  gene_col = "gene",
  min_pval = 1e-200,
  z_col = "Z",
  info_col = "info",
  chr_col = "CHROM",
  pos_col = "POS",
  log_pval = FALSE
)  %>%
  as_tibble()


Format_mtDNAcn_chong <- Format_mtDNAcn_chong %>%
mutate(exposure = "mtDNA_chong")



```


``` {r read_in_longchamps}

exposure_path_longchamps = "/wynton/group/andrews/data/summary_stats/results/Longchamps2021mtdnacn.chrall.CPRA_b37.tsv.gz"

mtDNAcn_longchamps <- read_tsv(exposure_path_longchamps,comment="#")


# To format the GWAS results for MR analysis Exposure. 
# Rename right-side variables to match column names of your GWAS dataset.
Format_mtDNAcn_longchamps <- mtDNAcn_longchamps %>% format_data(.,
  type = "exposure",
  snps = NULL,
  header = TRUE,
  phenotype_col = "TRAIT",
  snp_col = "DBSNP_ID",
  beta_col = "BETA",
  se_col = "SE",
  eaf_col = "AF",
  effect_allele_col = "REF",
  other_allele_col = "ALT",
  pval_col = "P",
  units_col = "units",
  ncase_col = "N_CASES",
  ncontrol_col = "N_CTRLS",
  samplesize_col = "N",
  gene_col = "gene",
  min_pval = 1e-200,
  z_col = "Z",
  info_col = "info",
  chr_col = "CHROM",
  pos_col = "POS",
  log_pval = FALSE
)  %>%
  as_tibble()


Format_mtDNAcn_longchamps <- Format_mtDNAcn_longchamps %>%
mutate(exposure = "mtDNA_longchamps")



```



``` {r read_in_hagg}

exposure_path_hagg = "/wynton/group/andrews/data/summary_stats/results/Hagg2020mtdnacn.chrall.CPRA_b37.tsv.gz"

mtDNAcn_hagg <- read_tsv(exposure_path_hagg,comment="#")


# To format the GWAS results for MR analysis Exposure. 
# Rename right-side variables to match column names of your GWAS dataset.
Format_mtDNAcn_hagg <- mtDNAcn_hagg %>% format_data(.,
  type = "exposure",
  snps = NULL,
  header = TRUE,
  phenotype_col = "TRAIT",
  snp_col = "DBSNP_ID",
  beta_col = "BETA",
  se_col = "SE",
  eaf_col = "AF",
  effect_allele_col = "REF",
  other_allele_col = "ALT",
  pval_col = "P",
  units_col = "units",
  ncase_col = "N_CASES",
  ncontrol_col = "N_CTRLS",
  samplesize_col = "N",
  gene_col = "gene",
  min_pval = 1e-200,
  z_col = "Z",
  info_col = "info",
  chr_col = "CHROM",
  pos_col = "POS",
  log_pval = FALSE
)  %>%
  as_tibble()


Format_mtDNAcn_hagg <- Format_mtDNAcn_hagg %>%
mutate(exposure = "mtDNA_hagg")

```


## Importing and formatting Outcome

This step imports our GWAS Outcome data sets and formats them up for MR analysis

``` {r read_in_Bellenguez}

outcome_path_bellenguez = "/wynton/group/andrews/data/summary_stats/results/Bellenguez2022load.chrall.CPRA_b37.tsv.gz"

# Import Code for outcome GWAS 
OutcomeAD_Bellenguez <- read_tsv(outcome_path_bellenguez,comment="#")

# Rename right-side variables to match column names of your GWAS dataset 
outcome_AD_Format_Bellenguez <- OutcomeAD_Bellenguez %>%
  format_data(.,
    type = "outcome",
    snps = NULL,
    header = TRUE,
    phenotype_col = "Phenotype",
    snp_col = "DBSNP_ID",
    beta_col = "BETA",
    se_col = "SE",
    eaf_col = "AF",
    effect_allele_col = "ALT",
    other_allele_col = "REF",
    pval_col = "P",
    samplesize_col = "N",
    z_col = "Z",
    chr_col = "CHROM",
    pos_col = "POS",
    log_pval = FALSE
) %>%
  as_tibble()

outcome_AD_Format_Bellenguez <- outcome_AD_Format_Bellenguez %>%
mutate(outcome = "bellenguez")

```



``` {r read_in_Kunkle}


outcome_path_kunkle = "/wynton/group/andrews/data/summary_stats/results/Kunkle2019load_stage123.chrall.CPRA_b37.tsv.gz"

# Import Code for outcome GWAS 
OutcomeAD_kunkle <- read_tsv(outcome_path_kunkle,comment="#")

# Rename right-side variables to match column names of your GWAS dataset 

outcome_AD_Format_kunkle <- OutcomeAD_kunkle %>%
  format_data(.,
    type = "outcome",
    snps = NULL,
    header = TRUE,
    phenotype_col = "Phenotype",
    snp_col = "DBSNP_ID",
    beta_col = "BETA",
    se_col = "SE",
    eaf_col = "AF",
    effect_allele_col = "ALT",
    other_allele_col = "REF",
    pval_col = "P",
    samplesize_col = "N",
    z_col = "Z",
    chr_col = "CHROM",
    pos_col = "POS",
    log_pval = FALSE
) %>%
  as_tibble()

outcome_AD_Format_kunkle <- outcome_AD_Format_kunkle %>%
mutate(outcome = "kunkle")

```


## Clumping Step
This step involving clumping and pruning to eliminate High linkage 
disequilibrium SNP (SNPS correlated with other SNPs). Removing High linkage 
disequilibrium SNP's are crucial to avoid "double counting" SNP's and
biasing our MR estimate.
The dataset will also be filter to remove a portion of insignificant SNP's for 
computational speed improvements.

``` {r eval=FALSE}

#Extract Instrumental variables from the exposure.
# Note that the lower Clump R^2 is, the more stringent, removes correlated 
# snps and retains the SNP with the lowest p-value. 
# Filter allows for more time efficient clumping, it 
# filters all of the significant SNPS and leaves all non significant 
# SNPs from the analysis.

# Filter Clumped SNPS will only include those SNPS whose P value is 
# Below 5e-8, Colname for P value is Pval.exposure in the filtered dataset.




# Chong Filter and Clumping
mtDNAcn_filter_chong <- filter(Format_mtDNAcn_chong , pval.exposure < 5e-8)


mtDNAcn_clump_chong <- clump_data(mtDNAcn_filter_chong, clump_kb = 10000,
                            clump_r2 = 0.001,
                            clump_p1 = 1,
                            clump_p2 = 1,
                            pop = "EUR"
)%>% 
  as_tibble()


# longchamps Filter and Clumping

mtDNAcn_filter_longchamps <- filter(Format_mtDNAcn_longchamps, pval.exposure < 5e-8)


mtDNAcn_clump_longchamps <- clump_data(mtDNAcn_filter_longchamps, clump_kb = 10000,
                            clump_r2 = 0.001,
                            clump_p1 = 1,
                            clump_p2 = 1,
                            pop = "EUR"
)%>% 
  as_tibble()


# hagg Filter and Clumping

mtDNAcn_filter_hagg <- filter(Format_mtDNAcn_hagg, pval.exposure < 5e-8)


mtDNAcn_clump_hagg <- clump_data(mtDNAcn_filter_hagg, clump_kb = 10000,
                            clump_r2 = 0.001,
                            clump_p1 = 1,
                            clump_p2 = 1,
                            pop = "EUR"
)%>% 
  as_tibble()



# Rename Data sets to be used in LD Proxy step

mtDNAcn_clump_chong_data <- mtDNAcn_clump_chong 

mtDNAcn_clump_longchamps_data <-  mtDNAcn_clump_longchamps

mtDNAcn_clump_hagg_data <-  mtDNAcn_clump_hagg

```


## LD Proxies
This attempts to retain any missing SNP's that are present in the 
exposure but not the outcome GWAS by utilizing LD proxies in their place. This step will 
be done multiple times for each exposure > outcome pair.

Note: If an error is encountered in the query request, check "exp_snps_wo"
to see if there are any snps to query, if none exist (will return 0 rows), then
there is no missing SNPS and no need to run LD proxies step for that 
exposure/outcome pair.
Just skip to run harmonization on that exposure outcome pair.
Example: No outcome SNPs were missing in the exposure for
longchamps_kunkle. So skip LD proxy step.


``` {r chong_kunkle_LDproxy}

# extract exposure SNPs present in outcome (outcome datasets,then exposure dataset)
outcome_clump <- semi_join(
  outcome_AD_Format_kunkle, mtDNAcn_clump_chong_data, by = "SNP"
)

# Exposure SNPs not present in outcome (exposure dataset, then outcome dataset)
exp_snps_wo <- anti_join(
  mtDNAcn_clump_chong_data, outcome_AD_Format_kunkle, by = "SNP"
)


# Use LDLinkR to identify proxy snps based on SNP dataset above
# that queries SNPs with no corresponding match in the outcome.
LDproxy_batch(exp_snps_wo$SNP, 
        pop = "CEU", 
        r2d = "r2", 
        token = 'a6deee62cc4a', 
        append = TRUE,
        genome_build = "grch37")
system("mv combined_query_snp_list_grch37.txt data/exposure_outcome_proxy_snps_chong_kunkle.txt")


# Munge proxy snp file
proxy_snps <- read_tsv("data/exposure_outcome_proxy_snps_chong_kunkle.txt", skip = 1, col_names = F) %>%
  rename(id = X1, func = X2, proxy_snp = X3, coord = X4, alleles = X5, maf = X6, 
         distance = X7, dprime = X8, rsq = X9, correlated_alleles = X10, FORGEdb = X11, RegulomeDB = X12) %>%
  separate(coord, c('chr', 'pos'), sep = ":") %>%
  mutate(snp = ifelse(id == 1, proxy_snp, NA), 
         chr = str_replace(chr, 'chr', ""), 
         chr = as.numeric(chr), 
         pos = as.numeric(pos)) %>%
  fill(snp, .direction = 'down') %>%
  relocate(snp, .before = proxy_snp) %>%
  dplyr::select(-id, -func, -FORGEdb, -RegulomeDB) %>%
  filter(rsq >= 0.8)

# Munge proxy snp and outcome data 
proxy_outcome <- left_join(
  proxy_snps, outcome_AD_Format_kunkle, by = c("proxy_snp" = "SNP")
) %>%
  separate(correlated_alleles, c("target_a1.outcome", "proxy_a1.outcome", 
                                 "target_a2.outcome", "proxy_a2.outcome"), sep = ",|=") %>%
  filter(!is.na(chr.outcome)) %>%
  arrange(snp, -rsq, abs(distance)) %>%
  group_by(snp) %>%
  slice(1) %>%
  ungroup() %>%
  mutate(
     proxy.outcome = TRUE,
     target_snp.outcome = snp,
     proxy_snp.outcome = proxy_snp, 
  ) %>% 
  mutate(
       new_effect_allele.outcome = case_when(
        proxy_a1.outcome == effect_allele.outcome & proxy_a2.outcome == other_allele.outcome ~ target_a1.outcome,
        proxy_a2.outcome == effect_allele.outcome & proxy_a1.outcome == other_allele.outcome ~ target_a2.outcome,
        TRUE ~ NA_character_
     ), 
      new_other_allele.outcome = case_when(
        proxy_a1.outcome == effect_allele.outcome & proxy_a2.outcome == other_allele.outcome ~ target_a2.outcome,
        proxy_a2.outcome == effect_allele.outcome & proxy_a1.outcome == other_allele.outcome ~ target_a1.outcome,
        TRUE ~ NA_character_
     ), 
     effect_allele.outcome = new_effect_allele.outcome, 
     other_allele.outcome = new_other_allele.outcome
  ) %>%
  dplyr::select(-proxy_snp, -chr, -pos, -alleles, -maf, -distance, -rsq, -dprime,  
         -new_effect_allele.outcome, -new_other_allele.outcome) %>%
  relocate(target_a1.outcome, proxy_a1.outcome, target_a2.outcome, proxy_a2.outcome, .after = proxy_snp.outcome) %>%
  rename(SNP = snp) %>%
  relocate(SNP, .after = samplesize.outcome)


# Merge outcome and proxy outcomes
chong_kunkle_dat_proxy <- bind_rows(
  outcome_clump, proxy_outcome
) %>% 
  arrange(chr.outcome, pos.outcome)

```

## No LD Proxies for longchamps_kunkle pair.
Code is shown for prosperity.

``` {r longchamps_kunkle_LDproxy}

# extract exposure SNPs present in outcome (outcome dataset,then exposure dataset)
outcome_clump <- semi_join(
  outcome_AD_Format_kunkle, mtDNAcn_clump_longchamps_data, by = "SNP"
)

# Exposure SNPs not present in outcome (exposure dataset, then outcome dataset)
exp_snps_wo <- anti_join(
  mtDNAcn_clump_longchamps_data, outcome_AD_Format_kunkle, by = "SNP"
)


# Use LDLinkR to identify proxy snps based on SNP dataset above
# that queries SNPs with no corresponding match in the outcome.
LDproxy_batch(exp_snps_wo$SNP, 
        pop = "CEU", 
        r2d = "r2", 
        token = 'a6deee62cc4a', 
        append = TRUE,
        genome_build = "grch37")
system("mv combined_query_snp_list_grch37.txt data/exposure_outcome_proxy_snps_longchamps_kunkle.txt")


# Munge proxy snp file
proxy_snps <- read_tsv("data/exposure_outcome_proxy_snps_longchamps_kunkle.txt", skip = 1, col_names = F) %>%
  rename(id = X1, func = X2, proxy_snp = X3, coord = X4, alleles = X5, maf = X6, 
         distance = X7, dprime = X8, rsq = X9, correlated_alleles = X10, FORGEdb = X11, RegulomeDB = X12) %>%
  separate(coord, c('chr', 'pos'), sep = ":") %>%
  mutate(snp = ifelse(id == 1, proxy_snp, NA), 
         chr = str_replace(chr, 'chr', ""), 
         chr = as.numeric(chr), 
         pos = as.numeric(pos)) %>%
  fill(snp, .direction = 'down') %>%
  relocate(snp, .before = proxy_snp) %>%
  dplyr::select(-id, -func, -FORGEdb, -RegulomeDB) %>%
  filter(rsq >= 0.8)

# Munge proxy snp and outcome data 
proxy_outcome <- left_join(
  proxy_snps, outcome_AD_Format_kunkle, by = c("proxy_snp" = "SNP")
) %>%
  separate(correlated_alleles, c("target_a1.outcome", "proxy_a1.outcome", 
                                 "target_a2.outcome", "proxy_a2.outcome"), sep = ",|=") %>%
  filter(!is.na(chr.outcome)) %>%
  arrange(snp, -rsq, abs(distance)) %>%
  group_by(snp) %>%
  slice(1) %>%
  ungroup() %>%
  mutate(
     proxy.outcome = TRUE,
     target_snp.outcome = snp,
     proxy_snp.outcome = proxy_snp, 
  ) %>% 
  mutate(
       new_effect_allele.outcome = case_when(
        proxy_a1.outcome == effect_allele.outcome & proxy_a2.outcome == other_allele.outcome ~ target_a1.outcome,
        proxy_a2.outcome == effect_allele.outcome & proxy_a1.outcome == other_allele.outcome ~ target_a2.outcome,
        TRUE ~ NA_character_
     ), 
      new_other_allele.outcome = case_when(
        proxy_a1.outcome == effect_allele.outcome & proxy_a2.outcome == other_allele.outcome ~ target_a2.outcome,
        proxy_a2.outcome == effect_allele.outcome & proxy_a1.outcome == other_allele.outcome ~ target_a1.outcome,
        TRUE ~ NA_character_
     ), 
     effect_allele.outcome = new_effect_allele.outcome, 
     other_allele.outcome = new_other_allele.outcome
  ) %>%
  dplyr::select(-proxy_snp, -chr, -pos, -alleles, -maf, -distance, -rsq, -dprime,  
         -new_effect_allele.outcome, -new_other_allele.outcome) %>%
  relocate(target_a1.outcome, proxy_a1.outcome, target_a2.outcome, proxy_a2.outcome, .after = proxy_snp.outcome) %>%
  rename(SNP = snp) %>%
  relocate(SNP, .after = samplesize.outcome)


# Merge outcome and proxy outcomes
longchamps_kunkle_dat_proxy <- bind_rows(
  outcome_clump, proxy_outcome
) %>% 
  arrange(chr.outcome, pos.outcome)

```



``` {r hagg_kunkle_LDproxy}


# extract exposure SNPs present in outcome (outcome datasets,then exposure dataset)
outcome_clump <- semi_join(
  outcome_AD_Format_kunkle, mtDNAcn_clump_hagg_data, by = "SNP"
)

# Exposure SNPs not present in outcome (exposure dataset, then outcome dataset)
exp_snps_wo <- anti_join(
  mtDNAcn_clump_hagg_data, outcome_AD_Format_kunkle, by = "SNP"
)


# Use LDLinkR to identify proxy snps based on SNP dataset above
# that queries SNPs with no corresponding match in the outcome.
LDproxy_batch(exp_snps_wo$SNP, 
        pop = "CEU", 
        r2d = "r2", 
        token = 'a6deee62cc4a', 
        append = TRUE,
        genome_build = "grch37")
system("mv combined_query_snp_list_grch37.txt data/exposure_outcome_proxy_snps_hagg_kunkle.txt")


# Munge proxy snp file
proxy_snps <- read_tsv("data/exposure_outcome_proxy_snps_hagg_kunkle.txt", skip = 1, col_names = F) %>%
  rename(id = X1, func = X2, proxy_snp = X3, coord = X4, alleles = X5, maf = X6, 
         distance = X7, dprime = X8, rsq = X9, correlated_alleles = X10, FORGEdb = X11, RegulomeDB = X12) %>%
  separate(coord, c('chr', 'pos'), sep = ":") %>%
  mutate(snp = ifelse(id == 1, proxy_snp, NA), 
         chr = str_replace(chr, 'chr', ""), 
         chr = as.numeric(chr), 
         pos = as.numeric(pos)) %>%
  fill(snp, .direction = 'down') %>%
  relocate(snp, .before = proxy_snp) %>%
  dplyr::select(-id, -func, -FORGEdb, -RegulomeDB) %>%
  filter(rsq >= 0.8)

# Munge proxy snp and outcome data 
proxy_outcome <- left_join(
  proxy_snps, outcome_AD_Format_kunkle, by = c("proxy_snp" = "SNP")
) %>%
  separate(correlated_alleles, c("target_a1.outcome", "proxy_a1.outcome", 
                                 "target_a2.outcome", "proxy_a2.outcome"), sep = ",|=") %>%
  filter(!is.na(chr.outcome)) %>%
  arrange(snp, -rsq, abs(distance)) %>%
  group_by(snp) %>%
  slice(1) %>%
  ungroup() %>%
  mutate(
     proxy.outcome = TRUE,
     target_snp.outcome = snp,
     proxy_snp.outcome = proxy_snp, 
  ) %>% 
  mutate(
       new_effect_allele.outcome = case_when(
        proxy_a1.outcome == effect_allele.outcome & proxy_a2.outcome == other_allele.outcome ~ target_a1.outcome,
        proxy_a2.outcome == effect_allele.outcome & proxy_a1.outcome == other_allele.outcome ~ target_a2.outcome,
        TRUE ~ NA_character_
     ), 
      new_other_allele.outcome = case_when(
        proxy_a1.outcome == effect_allele.outcome & proxy_a2.outcome == other_allele.outcome ~ target_a2.outcome,
        proxy_a2.outcome == effect_allele.outcome & proxy_a1.outcome == other_allele.outcome ~ target_a1.outcome,
        TRUE ~ NA_character_
     ), 
     effect_allele.outcome = new_effect_allele.outcome, 
     other_allele.outcome = new_other_allele.outcome
  ) %>%
  dplyr::select(-proxy_snp, -chr, -pos, -alleles, -maf, -distance, -rsq, -dprime,  
         -new_effect_allele.outcome, -new_other_allele.outcome) %>%
  relocate(target_a1.outcome, proxy_a1.outcome, target_a2.outcome, proxy_a2.outcome, .after = proxy_snp.outcome) %>%
  rename(SNP = snp) %>%
  relocate(SNP, .after = samplesize.outcome)


# Merge outcome and proxy outcomes
hagg_kunkle_dat_proxy <- bind_rows(
  outcome_clump, proxy_outcome
) %>% 
  arrange(chr.outcome, pos.outcome)

```


``` {r chong_bellenguez_LDproxy}


# extract exposure SNPs present in outcome (outcome datasets,then exposure dataset)
outcome_clump <- semi_join(
  outcome_AD_Format_bellenguez, mtDNAcn_clump_chong_data, by = "SNP"
)

# Exposure SNPs not present in outcome (exposure dataset, then outcome dataset)
exp_snps_wo <- anti_join(
  mtDNAcn_clump_chong_data, outcome_AD_Format_bellenguez, by = "SNP"
)


# Use LDLinkR to identify proxy snps based on SNP dataset above
# that queries SNPs with no corresponding match in the outcome.
LDproxy_batch(exp_snps_wo$SNP, 
        pop = "CEU", 
        r2d = "r2", 
        token = 'a6deee62cc4a', 
        append = TRUE,
        genome_build = "grch37")
system("mv combined_query_snp_list_grch37.txt data/exposure_outcome_proxy_snps_chong_bellenguez.txt")


# Munge proxy snp file
proxy_snps <- read_tsv("data/exposure_outcome_proxy_snps_chong_bellenguez.txt", skip = 1, col_names = F) %>%
  rename(id = X1, func = X2, proxy_snp = X3, coord = X4, alleles = X5, maf = X6, 
         distance = X7, dprime = X8, rsq = X9, correlated_alleles = X10, FORGEdb = X11, RegulomeDB = X12) %>%
  separate(coord, c('chr', 'pos'), sep = ":") %>%
  mutate(snp = ifelse(id == 1, proxy_snp, NA), 
         chr = str_replace(chr, 'chr', ""), 
         chr = as.numeric(chr), 
         pos = as.numeric(pos)) %>%
  fill(snp, .direction = 'down') %>%
  relocate(snp, .before = proxy_snp) %>%
  dplyr::select(-id, -func, -FORGEdb, -RegulomeDB) %>%
  filter(rsq >= 0.8)

# Munge proxy snp and outcome data 
proxy_outcome <- left_join(
  proxy_snps, outcome_AD_Format_bellenguez, by = c("proxy_snp" = "SNP")
) %>%
  separate(correlated_alleles, c("target_a1.outcome", "proxy_a1.outcome", 
                                 "target_a2.outcome", "proxy_a2.outcome"), sep = ",|=") %>%
  filter(!is.na(chr.outcome)) %>%
  arrange(snp, -rsq, abs(distance)) %>%
  group_by(snp) %>%
  slice(1) %>%
  ungroup() %>%
  mutate(
     proxy.outcome = TRUE,
     target_snp.outcome = snp,
     proxy_snp.outcome = proxy_snp, 
  ) %>% 
  mutate(
       new_effect_allele.outcome = case_when(
        proxy_a1.outcome == effect_allele.outcome & proxy_a2.outcome == other_allele.outcome ~ target_a1.outcome,
        proxy_a2.outcome == effect_allele.outcome & proxy_a1.outcome == other_allele.outcome ~ target_a2.outcome,
        TRUE ~ NA_character_
     ), 
      new_other_allele.outcome = case_when(
        proxy_a1.outcome == effect_allele.outcome & proxy_a2.outcome == other_allele.outcome ~ target_a2.outcome,
        proxy_a2.outcome == effect_allele.outcome & proxy_a1.outcome == other_allele.outcome ~ target_a1.outcome,
        TRUE ~ NA_character_
     ), 
     effect_allele.outcome = new_effect_allele.outcome, 
     other_allele.outcome = new_other_allele.outcome
  ) %>%
  dplyr::select(-proxy_snp, -chr, -pos, -alleles, -maf, -distance, -rsq, -dprime,  
         -new_effect_allele.outcome, -new_other_allele.outcome) %>%
  relocate(target_a1.outcome, proxy_a1.outcome, target_a2.outcome, proxy_a2.outcome, .after = proxy_snp.outcome) %>%
  rename(SNP = snp) %>%
  relocate(SNP, .after = samplesize.outcome)


# Merge outcome and proxy outcomes
chong_bellenguez_dat_proxy <- bind_rows(
  outcome_clump, proxy_outcome
) %>% 
  arrange(chr.outcome, pos.outcome)

```


``` {r longchamps_bellenguez_LDproxy}


# extract exposure SNPs present in outcome (outcome datasets,then exposure dataset)
outcome_clump <- semi_join(
  outcome_AD_Format_bellenguez, mtDNAcn_clump_longchamps_data, by = "SNP"
)

# Exposure SNPs not present in outcome (exposure dataset, then outcome dataset)
exp_snps_wo <- anti_join(
  mtDNAcn_clump_longchamps_data, outcome_AD_Format_bellenguez, by = "SNP"
)


# Use LDLinkR to identify proxy snps based on SNP dataset above
# that queries SNPs with no corresponding match in the outcome.
LDproxy_batch(exp_snps_wo$SNP, 
        pop = "CEU", 
        r2d = "r2", 
        token = 'a6deee62cc4a', 
        append = TRUE,
        genome_build = "grch37")
system("mv combined_query_snp_list_grch37.txt data/exposure_outcome_proxy_snps_longchamps_bellenguez.txt")


# Munge proxy snp file
proxy_snps <- read_tsv("data/exposure_outcome_proxy_snps_longchamps_bellenguez.txt", skip = 1, col_names = F) %>%
  rename(id = X1, func = X2, proxy_snp = X3, coord = X4, alleles = X5, maf = X6, 
         distance = X7, dprime = X8, rsq = X9, correlated_alleles = X10, FORGEdb = X11, RegulomeDB = X12) %>%
  separate(coord, c('chr', 'pos'), sep = ":") %>%
  mutate(snp = ifelse(id == 1, proxy_snp, NA), 
         chr = str_replace(chr, 'chr', ""), 
         chr = as.numeric(chr), 
         pos = as.numeric(pos)) %>%
  fill(snp, .direction = 'down') %>%
  relocate(snp, .before = proxy_snp) %>%
  dplyr::select(-id, -func, -FORGEdb, -RegulomeDB) %>%
  filter(rsq >= 0.8)

# Munge proxy snp and outcome data 
proxy_outcome <- left_join(
  proxy_snps, outcome_AD_Format_bellenguez, by = c("proxy_snp" = "SNP")
) %>%
  separate(correlated_alleles, c("target_a1.outcome", "proxy_a1.outcome", 
                                 "target_a2.outcome", "proxy_a2.outcome"), sep = ",|=") %>%
  filter(!is.na(chr.outcome)) %>%
  arrange(snp, -rsq, abs(distance)) %>%
  group_by(snp) %>%
  slice(1) %>%
  ungroup() %>%
  mutate(
     proxy.outcome = TRUE,
     target_snp.outcome = snp,
     proxy_snp.outcome = proxy_snp, 
  ) %>% 
  mutate(
       new_effect_allele.outcome = case_when(
        proxy_a1.outcome == effect_allele.outcome & proxy_a2.outcome == other_allele.outcome ~ target_a1.outcome,
        proxy_a2.outcome == effect_allele.outcome & proxy_a1.outcome == other_allele.outcome ~ target_a2.outcome,
        TRUE ~ NA_character_
     ), 
      new_other_allele.outcome = case_when(
        proxy_a1.outcome == effect_allele.outcome & proxy_a2.outcome == other_allele.outcome ~ target_a2.outcome,
        proxy_a2.outcome == effect_allele.outcome & proxy_a1.outcome == other_allele.outcome ~ target_a1.outcome,
        TRUE ~ NA_character_
     ), 
     effect_allele.outcome = new_effect_allele.outcome, 
     other_allele.outcome = new_other_allele.outcome
  ) %>%
  dplyr::select(-proxy_snp, -chr, -pos, -alleles, -maf, -distance, -rsq, -dprime,  
         -new_effect_allele.outcome, -new_other_allele.outcome) %>%
  relocate(target_a1.outcome, proxy_a1.outcome, target_a2.outcome, proxy_a2.outcome, .after = proxy_snp.outcome) %>%
  rename(SNP = snp) %>%
  relocate(SNP, .after = samplesize.outcome)


# Merge outcome and proxy outcomes
longchamps_bellenguez_dat_proxy <- bind_rows(
  outcome_clump, proxy_outcome
) %>% 
  arrange(chr.outcome, pos.outcome)

```


``` {r hagg_bellenguez_LDproxy}


# extract exposure SNPs present in outcome (outcome datasets,then exposure dataset)
outcome_clump <- semi_join(
  outcome_AD_Format_bellenguez, mtDNAcn_clump_hagg_data, by = "SNP"
)

# Exposure SNPs not present in outcome (exposure dataset, then outcome dataset)
exp_snps_wo <- anti_join(
  mtDNAcn_clump_hagg_data, outcome_AD_Format_bellenguez, by = "SNP"
)


# Use LDLinkR to identify proxy snps based on SNP dataset above
# that queries SNPs with no corresponding match in the outcome.
LDproxy_batch(exp_snps_wo$SNP, 
        pop = "CEU", 
        r2d = "r2", 
        token = 'a6deee62cc4a', 
        append = TRUE,
        genome_build = "grch37")
system("mv combined_query_snp_list_grch37.txt data/exposure_outcome_proxy_snps_hagg_bellenguez.txt")


# Munge proxy snp file
proxy_snps <- read_tsv("data/exposure_outcome_proxy_snps_hagg_bellenguez.txt", skip = 1, col_names = F) %>%
  rename(id = X1, func = X2, proxy_snp = X3, coord = X4, alleles = X5, maf = X6, 
         distance = X7, dprime = X8, rsq = X9, correlated_alleles = X10, FORGEdb = X11, RegulomeDB = X12) %>%
  separate(coord, c('chr', 'pos'), sep = ":") %>%
  mutate(snp = ifelse(id == 1, proxy_snp, NA), 
         chr = str_replace(chr, 'chr', ""), 
         chr = as.numeric(chr), 
         pos = as.numeric(pos)) %>%
  fill(snp, .direction = 'down') %>%
  relocate(snp, .before = proxy_snp) %>%
  dplyr::select(-id, -func, -FORGEdb, -RegulomeDB) %>%
  filter(rsq >= 0.8)

# Munge proxy snp and outcome data 
proxy_outcome <- left_join(
  proxy_snps, outcome_AD_Format_bellenguez, by = c("proxy_snp" = "SNP")
) %>%
  separate(correlated_alleles, c("target_a1.outcome", "proxy_a1.outcome", 
                                 "target_a2.outcome", "proxy_a2.outcome"), sep = ",|=") %>%
  filter(!is.na(chr.outcome)) %>%
  arrange(snp, -rsq, abs(distance)) %>%
  group_by(snp) %>%
  slice(1) %>%
  ungroup() %>%
  mutate(
     proxy.outcome = TRUE,
     target_snp.outcome = snp,
     proxy_snp.outcome = proxy_snp, 
  ) %>% 
  mutate(
       new_effect_allele.outcome = case_when(
        proxy_a1.outcome == effect_allele.outcome & proxy_a2.outcome == other_allele.outcome ~ target_a1.outcome,
        proxy_a2.outcome == effect_allele.outcome & proxy_a1.outcome == other_allele.outcome ~ target_a2.outcome,
        TRUE ~ NA_character_
     ), 
      new_other_allele.outcome = case_when(
        proxy_a1.outcome == effect_allele.outcome & proxy_a2.outcome == other_allele.outcome ~ target_a2.outcome,
        proxy_a2.outcome == effect_allele.outcome & proxy_a1.outcome == other_allele.outcome ~ target_a1.outcome,
        TRUE ~ NA_character_
     ), 
     effect_allele.outcome = new_effect_allele.outcome, 
     other_allele.outcome = new_other_allele.outcome
  ) %>%
  dplyr::select(-proxy_snp, -chr, -pos, -alleles, -maf, -distance, -rsq, -dprime,  
         -new_effect_allele.outcome, -new_other_allele.outcome) %>%
  relocate(target_a1.outcome, proxy_a1.outcome, target_a2.outcome, proxy_a2.outcome, .after = proxy_snp.outcome) %>%
  rename(SNP = snp) %>%
  relocate(SNP, .after = samplesize.outcome)


# Merge outcome and proxy outcomes
hagg_bellenguez_dat_proxy <- bind_rows(
  outcome_clump, proxy_outcome
) %>% 
  arrange(chr.outcome, pos.outcome)

```



## Harmonization and appending Step

Harmonize and combine all six pairs of analysis.

Chong_Kunkle
Longchamps_Kunkle
Hagg_Kunkle
Chong_Bellenguez
Longchamps_Bellenguez
Hagg_Bellenguez


``` {r eval=FALSE}
#  Change the two datasets names to ones created
# while performing clumping and LD proxies. Put LD proxy dataset in outcome slot (right)

# chong_kunkle_harmonization
mtDNAcn_AD_Harm_chong_kunkle <- harmonise_data(mtDNAcn_clump_chong, chong_kunkle_dat_proxy,action=2)


# longchamps_kunkle_harmonization (harmonization done with formatting dataset, no LD proxies found.)

mtDNAcn_AD_Harm_longchamps_kunkle <- harmonise_data(mtDNAcn_clump_longchamps, outcome_AD_Format_kunkle,action=2)

# hagg_kunkle_harmonization


mtDNAcn_AD_Harm_hagg_kunkle <- harmonise_data(mtDNAcn_clump_hagg, hagg_kunkle_dat_proxy,action=2)


# chong_bellenguez_harmonization

mtDNAcn_AD_Harm_chong_bellenguez <- harmonise_data(mtDNAcn_clump_chong, chong_bellenguez_dat_proxy,action=2)

# longchamps_bellenguez_harmonization

mtDNAcn_AD_Harm_longchamps_bellenguez <- harmonise_data(mtDNAcn_clump_longchamps, longchamps_bellenguez_dat_proxy,action=2)

# hagg_bellenguez_harmonization

mtDNAcn_AD_Harm_hagg_bellenguez <- harmonise_data(mtDNAcn_clump_hagg, hagg_bellenguez_dat_proxy,action=2)
```




## Radial MR to Remove Outliers in each dataset pair.

Radial MR will be conducted to identify outliers.

``` {r}
#  Correction for multiple testing bonferroni correction .05/nrow

# Chong Kunkle pair

AD_radial_ivw_mtDNAcn_chong_kunkle <- ivw_radial(mtDNAcn_AD_Harm_chong_kunkle, .05/nrow(mtDNAcn_AD_Harm_chong_kunkle), 2, .0001, TRUE)




#To view radial MR

AD_radial_ivw_mtDNAcn_chong_kunkle 


# Static plot results of Radial MR

radial_p_chong_kunkle_ivw <- plot_radial(AD_radial_ivw_mtDNAcn_chong_kunkle, radial_scale = F, show_outliers = T)

radial_p_chong_kunkle_ivw


# longchamps Kunkle pair

AD_radial_ivw_mtDNAcn_longchamps_kunkle <- ivw_radial(mtDNAcn_AD_Harm_longchamps_kunkle, .05/nrow(mtDNAcn_AD_Harm_longchamps_kunkle), 2, .0001, TRUE)



AD_radial_ivw_mtDNAcn_longchamps_kunkle 


radial_p_longchamps_kunkle_ivw <- plot_radial(AD_radial_ivw_mtDNAcn_longchamps_kunkle, radial_scale = F, show_outliers = T)

radial_p_longchamps_kunkle_ivw


# Hagg Kunkle pair

AD_radial_ivw_mtDNAcn_hagg_kunkle <- ivw_radial(mtDNAcn_AD_Harm_hagg_kunkle, .05/nrow(mtDNAcn_AD_Harm_hagg_kunkle), 2, .0001, TRUE)



AD_radial_ivw_mtDNAcn_hagg_kunkle 


radial_p_hagg_kunkle_ivw <- plot_radial(AD_radial_ivw_mtDNAcn_hagg_kunkle, radial_scale = F, show_outliers = T)

radial_p_hagg_kunkle_ivw



# Chong Bellenguez pair

AD_radial_ivw_mtDNAcn_chong_bellenguez <- ivw_radial(mtDNAcn_AD_Harm_chong_bellenguez, .05/nrow(mtDNAcn_AD_Harm_chong_bellenguez), 2, .0001, TRUE)



AD_radial_ivw_mtDNAcn_chong_bellenguez 


radial_p_chong_bellenguez_ivw <- plot_radial(AD_radial_ivw_mtDNAcn_chong_bellenguez, radial_scale = F, show_outliers = T)

radial_p_chong_bellenguez_ivw


# longchamps Bellenguez pair

AD_radial_ivw_mtDNAcn_longchamps_bellenguez <- ivw_radial(mtDNAcn_AD_Harm_longchamps_bellenguez, .05/nrow(mtDNAcn_AD_Harm_longchamps_bellenguez), 2, .0001, TRUE)



AD_radial_ivw_mtDNAcn_longchamps_bellenguez 


radial_p_longchamps_bellenguez_ivw <- plot_radial(AD_radial_ivw_mtDNAcn_longchamps_bellenguez, radial_scale = F, show_outliers = T)

radial_p_longchamps_bellenguez_ivw


# Hagg Bellenguez pair

AD_radial_ivw_mtDNAcn_hagg_bellenguez <- ivw_radial(mtDNAcn_AD_Harm_hagg_bellenguez, .05/nrow(mtDNAcn_AD_Harm_hagg_bellenguez), 2, .0001, TRUE)


AD_radial_ivw_mtDNAcn_hagg_bellenguez 


radial_p_hagg_bellenguez_ivw <- plot_radial(AD_radial_ivw_mtDNAcn_hagg_bellenguez, radial_scale = F, show_outliers = T)

radial_p_hagg_bellenguez_ivw


outlier_plots_set1<- grid.arrange(radial_p_chong_kunkle_ivw, radial_p_longchamps_kunkle_ivw, radial_p_hagg_kunkle_ivw, 
nrow = 1,
top = "longchamps_kunkle",
left = "chong_kunkle",
right= "hagg_kunkle"
) 

outlier_plots_set2 <- grid.arrange(radial_p_chong_bellenguez_ivw, radial_p_longchamps_bellenguez_ivw, radial_p_hagg_bellenguez_ivw, 
nrow = 1,
(arrangeGrob(p, top = 'title 1'), arrangeGrob(p, top = 'title 2'))
) 


outlier_plots_set2 <- grid.arrange(plot(radial_p_chong_bellenguez_ivw, main = 'chong_bellenguez'),
plot(radial_p_longchamps_bellenguez_ivw, main = 'title2'), plot(radial_p_hagg_bellenguez_ivw, main = 'title3'))

#Brian try making this work, if does not, just make to seprate figures for bellenguez/ kunkle
outlier_plots_set1 <- grid.arrange( arrangeGrob(radial_p_chong_kunkle_ivw, top = grid::textGrob("Chong_Kunkle", x = 0.09, y = .5, hjust = 0, gp=gpar(fontsize=10))), arrangeGrob(radial_p_longchamps_kunkle_ivw, top = grid::textGrob("Longchamps_Kunkle", x = 0.09, y = .5, hjust = 0, gp=gpar(fontsize=10))), arrangeGrob(radial_p_hagg_kunkle_ivw, top = grid::textGrob("Hagg_Kunkle", x = 0.09, y = 0.5, hjust = 0, gp=gpar(fontsize=10))), nrow=2, ncol=3)



outlier_plots_set2 <- grid.arrange( arrangeGrob(radial_p_chong_bellenguez_ivw, top = grid::textGrob("Chong_Bellenguez", x = 0.25, y = -4, hjust = 0)), arrangeGrob(radial_p_longchamps_bellenguez_ivw, top = grid::textGrob("Longchamps_Bellenguez", x = 0.25, y = -4, hjust = 0)), arrangeGrob(radial_p_hagg_bellenguez_ivw, top = grid::textGrob("Hagg_Bellenguez", x = 0.3, y = -4, hjust = 0 )), nrow=1) 



arrangeGrob(radial_p_chong_bellenguez_ivw, top = grid::textGrob("Chong_Bellenguez", x = 0.25, y = .5, hjust = 0, gp=gpar(fontsize=10))), arrangeGrob(radial_p_longchamps_bellenguez_ivw, top = grid::textGrob("Longchamps_Bellenguez", x = 0.25, y = .5, hjust = 0, gp=gpar(fontsize=10))), arrangeGrob(radial_p_hagg_bellenguez_ivw, top = grid::textGrob("Hagg_Bellenguez", x = 0.3, y = 0.5, hjust = 0, gp=gpar(fontsize=10) )), nrow=2, ncol=3)


```
## Query outliers detected by Radial MR
We use Phenoscanner to obtain more information on these variants.

``` {r outlier_details}

# Phenoscanner Chong_Kunkle

phewas_data_apoe_retained_C_K <- phenoscanner(
  snpquery=AD_radial_ivw_mtDNAcn_chong_kunkle$outliers$SNP, 
  proxies='EUR', 
  r2 = 0.8
  ) 

view(phewas_data_apoe_retained_C_K)


# Phenoscanner Longchamps_Kunkle

phewas_data_apoe_retained_L_K <- phenoscanner(
  snpquery=AD_radial_ivw_mtDNAcn_longchamps_kunkle$outliers$SNP, 
  proxies='EUR', 
  r2 = 0.8
  ) 

view(phewas_data_apoe_retained_L_K)

# Phenoscanner Hagg_Kunkle

phewas_data_apoe_retained_H_K <- phenoscanner(
  snpquery=AD_radial_ivw_mtDNAcn_hagg_kunkle$outliers$SNP, 
  proxies='EUR', 
  r2 = 0.8
  ) 

view(phewas_data_apoe_retained_H_K)


# Phenoscanner Chong_Bellenguez

phewas_data_apoe_retained_C_B <- phenoscanner(
  snpquery=AD_radial_ivw_mtDNAcn_chong_bellenguez$outliers$SNP, 
  proxies='EUR', 
  r2 = 0.8
  ) 
 

# Phenoscanner Longchamps_Bellenguez

phewas_data_apoe_retained_L_B <- phenoscanner(
  snpquery=AD_radial_ivw_mtDNAcn_longchamps_bellenguez$outliers$SNP, 
  proxies='EUR', 
  r2 = 0.8
  ) 

view(phewas_data_apoe_retained_L_B)


# No outliers for  Hagg_Bellenguez

# Unique Traits associated with outliers

# Chong Kunkle

Traits_C_K <- phewas_data_apoe_retained_C_K$results

view (Traits_C_K) 

length(unique(Traits_C_K$trait))

# Longchamps Kunkle

Traits_L_K <- phewas_data_apoe_retained_L_K$results

view (Traits_L_K) 

length(unique(Traits_L_K$trait))

# Hagg Kunkle

Traits_H_K <- phewas_data_apoe_retained_H_K$results

view (Traits_H_K) 

length(unique(Traits_H_K$trait))

# Chong Bellenguez (Multiple outliers requires grouping)

Traits_C_B <- phewas_data_apoe_retained_C_B$results


Traits_C_B_uni_count <- Traits_C_B %>%
  group_by(snp) %>%
  summarize(count = n_distinct(trait))

view(Traits_C_B_uni_count)

view (Traits_C_B) 


# Longchamps Bellenguez 

Traits_L_B <- phewas_data_apoe_retained_L_B$results

Traits_L_B_uni_count <- Traits_L_B %>%
  group_by(snp) %>%
  summarize(count = n_distinct(trait))

length(unique(Traits_L_B$trait))



view (Traits_L_B) 



# No outliers for Hagg Bellenguez 


# Export Outlier details tables.

write_csv(Traits_C_K, 'Casual_Estimate_mtDNAcn_on_Alzheimer/results/Outliers_Plots_All/Traits_C_K.csv')

write_csv(Traits_L_K, 'Casual_Estimate_mtDNAcn_on_Alzheimer/results/Outliers_Plots_All/Traits_L_K.csv')

write_csv(Traits_H_K, 'Casual_Estimate_mtDNAcn_on_Alzheimer/results/Outliers_Plots_All/Traits_H_K.csv')

write_csv(Traits_C_B, 'Casual_Estimate_mtDNAcn_on_Alzheimer/results/Outliers_Plots_All/Traits_C_B.csv')

write_csv(Traits_L_B, 'Casual_Estimate_mtDNAcn_on_Alzheimer/results/Outliers_Plots_All/Traits_L_B.csv')



```

## Export Radial Plots for all pairs

``` {r radial_outliers_export}


ggsave('results/Outliers_Plots_All/mr_radial_IVW_chong_kunkle.png', plot = radial_p_chong_kunkle_ivw, units = 'in', height = 4, width = 9)

ggsave('results/Outliers_Plots_All/mr_radial_IVW_longchamps_kunkle.png', plot = radial_p_longchamps_kunkle_ivw, units = 'in', height = 4, width = 9)

ggsave('results/Outliers_Plots_All/mr_radial_IVW_hagg_kunkle.png', plot = radial_p_hagg_kunkle_ivw, units = 'in', height = 4, width = 9)

ggsave('results/Outliers_Plots_All/mr_radial_IVW_chong_bellenguez.png', plot = radial_p_chong_bellenguez_ivw, units = 'in', height = 4, width = 9)

ggsave('results/Outliers_Plots_All/mr_radial_IVW_longchamps_bellenguez.png', plot = radial_p_longchamps_bellenguez_ivw, units = 'in', height = 4, width = 9)

ggsave('results/Outliers_Plots_All/mr_radial_IVW_hagg_bellenguez.png', plot = radial_p_hagg_bellenguez_ivw, units = 'in', height = 4, width = 9)

ggsave('results/Outliers_Plots_All/mr_radial_IVW_set1.png', plot = outlier_plots_set1, units = 'in', height = 4, width = 9)

ggsave('results/Outliers_Plots_All/mr_radial_IVW_set2.png', plot = outlier_plots_set2, units = 'in', height = 4, width = 9)



```


## Combining Radial data and harmonized
``` {r radial_combination}

# This step creates a new dataset combining our radial data with our harmonized
# data, creating a dichotomous variable (Columns) of "outlier",
# so when conducting  MR on this dataset, omission of outlier 
# SNPS that may violate MR assumptions is possible.

# chong_kunkle

mtDNAcn_AD_Harm_Outlier_chong_kunkle <- mtDNAcn_AD_Harm_chong_kunkle %>%
  left_join(AD_radial_ivw_mtDNAcn_chong_kunkle$data) %>%
  mutate( gws.outcome = ifelse(pval.outcome <5e-8, TRUE, FALSE),
    mr_keep = ifelse(mr_keep == FALSE | Outliers == "Outlier" | gws.outcome == TRUE, FALSE, TRUE)
  )


# longchamps_kunkle

mtDNAcn_AD_Harm_Outlier_longchamps_kunkle <- mtDNAcn_AD_Harm_longchamps_kunkle %>%
  left_join(AD_radial_ivw_mtDNAcn_longchamps_kunkle$data) %>%
  mutate( gws.outcome = ifelse(pval.outcome <5e-8, TRUE, FALSE),
    mr_keep = ifelse(mr_keep == FALSE | Outliers == "Outlier" | gws.outcome == TRUE, FALSE, TRUE)
  )

# hagg_kunkle

mtDNAcn_AD_Harm_Outlier_hagg_kunkle <- mtDNAcn_AD_Harm_hagg_kunkle %>%
  left_join(AD_radial_ivw_mtDNAcn_hagg_kunkle$data) %>%
  mutate( gws.outcome = ifelse(pval.outcome <5e-8, TRUE, FALSE),
    mr_keep = ifelse(mr_keep == FALSE | Outliers == "Outlier" | gws.outcome == TRUE, FALSE, TRUE)
  )


# chong_bellenguez

mtDNAcn_AD_Harm_Outlier_chong_bellenguez <- mtDNAcn_AD_Harm_chong_bellenguez %>%
  left_join(AD_radial_ivw_mtDNAcn_chong_bellenguez$data) %>%
  mutate( gws.outcome = ifelse(pval.outcome <5e-8, TRUE, FALSE),
    mr_keep = ifelse(mr_keep == FALSE | Outliers == "Outlier" | gws.outcome == TRUE, FALSE, TRUE)
  )

# longchamps_bellenguez

mtDNAcn_AD_Harm_Outlier_longchamps_bellenguez <- mtDNAcn_AD_Harm_longchamps_bellenguez %>%
  left_join(AD_radial_ivw_mtDNAcn_longchamps_bellenguez$data) %>%
  mutate( gws.outcome = ifelse(pval.outcome <5e-8, TRUE, FALSE),
    mr_keep = ifelse(mr_keep == FALSE | Outliers == "Outlier" | gws.outcome == TRUE, FALSE, TRUE)
  )

# hagg_bellenguez

mtDNAcn_AD_Harm_Outlier_hagg_bellenguez <- mtDNAcn_AD_Harm_hagg_bellenguez %>%
  left_join(AD_radial_ivw_mtDNAcn_hagg_bellenguez$data) %>%
  mutate( gws.outcome = ifelse(pval.outcome <5e-8, TRUE, FALSE),
    mr_keep = ifelse(mr_keep == FALSE | Outliers == "Outlier" | gws.outcome == TRUE, FALSE, TRUE)
  )

```


# Appending Step

``` {r append_all_data}

mtDNAcn_AD_harm_appended <- bind_rows(mtDNAcn_AD_Harm_Outlier_chong_kunkle, mtDNAcn_AD_Harm_Outlier_longchamps_kunkle, mtDNAcn_AD_Harm_Outlier_hagg_kunkle, mtDNAcn_AD_Harm_Outlier_chong_bellenguez, mtDNAcn_AD_Harm_Outlier_longchamps_bellenguez, mtDNAcn_AD_Harm_Outlier_hagg_bellenguez)


view(mtDNAcn_AD_harm_appended)

```

```{r export_harmonized_data}
write_csv(mtDNAcn_AD_harm_appended, 'data/Harmonized_Datasets/harmonized_data_appended.csv')
# mtDNAcn_AD_harm_appended <- read_csv('data/Harmonized_Datasets/harmonized_data_appended.csv')

 
```


## Preliminary MR 

We run a Initial MR analysis for at a glance analysis on the SNPS being used in
MR analysis. Potential outliers not removed from this analysis

```{r MR_analysis_all_dataset_pairs}

Mr_mtDNAcn_AD_appended_Prime <- mr(mtDNAcn_AD_harm_appended, method_list = c("mr_ivw_fe", "mr_egger_regression", "mr_weighted_median", "mr_weighted_mode"))

Mr_mtDNAcn_AD_appended_Prime_OR <- generate_odds_ratios(Mr_mtDNAcn_AD_appended_Prime)

view (Mr_mtDNAcn_AD_appended_Prime_OR )


```

## Perform Outlier Removed Sensentivity analysis

``` {r sensitivity}
# Heterogeneity (Test for Overall Violations of MR assumptions)
Mr_H_Test_prime <- mr_heterogeneity(mtDNAcn_AD_harm_appended)

View(Mr_H_Test_prime)

# (Test for Violations of MR assumptions via Horizontal pleiotrophy)
Mr_Horizontal_Test_prime <- mr_pleiotropy_test(mtDNAcn_AD_harm_appended)

View(Mr_Horizontal_Test_prime)




```

## Preliminary Plots
Using separate datasets to generate plots for each exposure outcome pair.

``` {r plots_}
# Chong_kunkle Preparation

MrAD_single_mtDNAcn_Prime_c_k <- mr_singlesnp(mtDNAcn_AD_Harm_Outlier_chong_kunkle)

MrAD_loo_mtDNAcn_Prime_c_k <- mr_leaveoneout(mtDNAcn_AD_Harm_Outlier_chong_kunkle)

#longchamps_kunkle Preparation

MrAD_single_mtDNAcn_Prime_l_k <- mr_singlesnp(mtDNAcn_AD_Harm_Outlier_longchamps_kunkle)

MrAD_loo_mtDNAcn_Prime_l_k <- mr_leaveoneout(mtDNAcn_AD_Harm_Outlier_longchamps_kunkle)


# hagg_kunkle Preparation

MrAD_single_mtDNAcn_Prime_h_k <- mr_singlesnp(mtDNAcn_AD_Harm_Outlier_hagg_kunkle)

MrAD_loo_mtDNAcn_Prime_h_k <- mr_leaveoneout(mtDNAcn_AD_Harm_Outlier_hagg_kunkle)

# chong_bellenguez Preparation
MrAD_single_mtDNAcn_Prime_c_b <- mr_singlesnp(mtDNAcn_AD_Harm_Outlier_chong_bellenguez)

MrAD_loo_mtDNAcn_Prime_c_b <- mr_leaveoneout(mtDNAcn_AD_Harm_Outlier_chong_bellenguez)

# longchamps_bellenguez Preparation

MrAD_single_mtDNAcn_Prime_l_b <- mr_singlesnp(mtDNAcn_AD_Harm_Outlier_longchamps_bellenguez)

MrAD_loo_mtDNAcn_Prime_l_b <- mr_leaveoneout(mtDNAcn_AD_Harm_Outlier_longchamps_bellenguez)

# hagg_bellenguez Preparation

MrAD_single_mtDNAcn_Prime_h_b <- mr_singlesnp(mtDNAcn_AD_Harm_Outlier_hagg_bellenguez)

MrAD_loo_mtDNAcn_Prime_h_b <- mr_leaveoneout(mtDNAcn_AD_Harm_Outlier_hagg_bellenguez)

# Generates Plots of single and leave one out anlysis MR

# Chong_kunkle Plots


p1_Prime_Single_c_k <- mr_forest_plot(MrAD_single_mtDNAcn_Prime_c_k )
p1_Prime_Single_c_k [[1]]

P2_Prime_Funnel_single_c_k <- mr_funnel_plot(MrAD_single_mtDNAcn_Prime_c_k)
P2_Prime_Funnel_single_c_k[[1]]

p3_Prime_loo_c_k <- mr_leaveoneout_plot(MrAD_loo_mtDNAcn_Prime_c_k)
p3_Prime_loo_c_k[[1]]


# Note, indivdual MR must be run in this step to create the plot

MR_scatter_mtDNAcn_ad <- mr(mtDNAcn_AD_Harm_Outlier_chong_kunkle)

p4_Prime_scatter <- mr_scatter_plot(MR_scatter_mtDNAcn_ad, mtDNAcn_AD_Harm_Outlier_chong_kunkle)

p4_scatter_prime_format_c_k <- p4_Prime_scatter[[1]] + theme_bw() + 
  guides(color=guide_legend(ncol =1)) + 
  theme(
    text = element_text(size = 8), 
  )
p4_scatter_prime_format_c_k


# longchamps_kunkle Plots

p1_Prime_Single_l_k <- mr_forest_plot(MrAD_single_mtDNAcn_Prime_l_k )
p1_Prime_Single_l_k [[1]]

P2_Prime_Funnel_single_l_k <- mr_funnel_plot(MrAD_single_mtDNAcn_Prime_l_k)
P2_Prime_Funnel_single_l_k[[1]]

p3_Prime_loo_l_k <- mr_leaveoneout_plot(MrAD_loo_mtDNAcn_Prime_l_k)
p3_Prime_loo_l_k[[1]]


MR_scatter_mtDNAcn_ad <- mr(mtDNAcn_AD_Harm_Outlier_longchamps_kunkle)

p4_Prime_scatter <- mr_scatter_plot(MR_scatter_mtDNAcn_ad, mtDNAcn_AD_Harm_Outlier_longchamps_kunkle)

p4_scatter_prime_format_l_k <- p4_Prime_scatter[[1]] + theme_bw() + 
  guides(color=guide_legend(ncol =1)) + 
  theme(
    text = element_text(size = 8), 
  )
p4_scatter_prime_format_l_k


# hagg_kunkle Plots

p1_Prime_Single_h_k <- mr_forest_plot(MrAD_single_mtDNAcn_Prime_h_k )
p1_Prime_Single_h_k [[1]]

P2_Prime_Funnel_single_h_k <- mr_funnel_plot(MrAD_single_mtDNAcn_Prime_h_k)
P2_Prime_Funnel_single_h_k[[1]]

p3_Prime_loo_h_k <- mr_leaveoneout_plot(MrAD_loo_mtDNAcn_Prime_h_k)
p3_Prime_loo_h_k[[1]]


MR_scatter_mtDNAcn_ad <- mr(mtDNAcn_AD_Harm_Outlier_hagg_kunkle)

p4_Prime_scatter <- mr_scatter_plot(MR_scatter_mtDNAcn_ad, mtDNAcn_AD_Harm_Outlier_hagg_kunkle)

p4_scatter_prime_format_h_k <- p4_Prime_scatter[[1]] + theme_bw() + 
  guides(color=guide_legend(ncol =1)) + 
  theme(
    text = element_text(size = 8), 
  )
p4_scatter_prime_format_h_k

# Chong_bellenguez Plots


p1_Prime_Single_c_b <- mr_forest_plot(MrAD_single_mtDNAcn_Prime_c_b )
p1_Prime_Single_c_b [[1]]

P2_Prime_Funnel_single_c_b <- mr_funnel_plot(MrAD_single_mtDNAcn_Prime_c_b)
P2_Prime_Funnel_single_c_b[[1]]

p3_Prime_loo_c_b <- mr_leaveoneout_plot(MrAD_loo_mtDNAcn_Prime_c_b)
p3_Prime_loo_c_b[[1]]

MR_scatter_mtDNAcn_ad <- mr(mtDNAcn_AD_Harm_Outlier_chong_bellenguez)

p4_Prime_scatter <- mr_scatter_plot(MR_scatter_mtDNAcn_ad, mtDNAcn_AD_Harm_Outlier_chong_bellenguez)

p4_scatter_prime_format_c_b <- p4_Prime_scatter[[1]] + theme_bw() + 
  guides(color=guide_legend(ncol =1)) + 
  theme(
    text = element_text(size = 8), 
  )
p4_scatter_prime_format_c_b

# longchamps_bellenguez Plots

p1_Prime_Single_l_b <- mr_forest_plot(MrAD_single_mtDNAcn_Prime_l_b )
p1_Prime_Single_l_b [[1]]

P2_Prime_Funnel_single_l_b <- mr_funnel_plot(MrAD_single_mtDNAcn_Prime_l_b)
P2_Prime_Funnel_single_l_b[[1]]

p3_Prime_loo_l_b <- mr_leaveoneout_plot(MrAD_loo_mtDNAcn_Prime_l_b)
p3_Prime_loo_l_b[[1]]


MR_scatter_mtDNAcn_ad <- mr(mtDNAcn_AD_Harm_Outlier_longchamps_bellenguez)

p4_Prime_scatter <- mr_scatter_plot(MR_scatter_mtDNAcn_ad, mtDNAcn_AD_Harm_Outlier_longchamps_bellenguez)

p4_scatter_prime_format_l_b <- p4_Prime_scatter[[1]] + theme_bw() + 
  guides(color=guide_legend(ncol =1)) + 
  theme(
    text = element_text(size = 8), 
  )
p4_scatter_prime_format_l_b

# hagg_bellenguez Plots

p1_Prime_Single_h_b <- mr_forest_plot(MrAD_single_mtDNAcn_Prime_h_b )
p1_Prime_Single_h_b [[1]]

P2_Prime_Funnel_single_h_b <- mr_funnel_plot(MrAD_single_mtDNAcn_Prime_h_b)
P2_Prime_Funnel_single_h_b[[1]]

p3_Prime_loo_h_b <- mr_leaveoneout_plot(MrAD_loo_mtDNAcn_Prime_h_b)
p3_Prime_loo_h_b[[1]]


MR_scatter_mtDNAcn_ad <- mr(mtDNAcn_AD_Harm_Outlier_hagg_bellenguez)

p4_Prime_scatter <- mr_scatter_plot(MR_scatter_mtDNAcn_ad, mtDNAcn_AD_Harm_Outlier_hagg_bellenguez)

p4_scatter_prime_format_h_b <- p4_Prime_scatter[[1]] + theme_bw() + 
  guides(color=guide_legend(ncol =1)) + 
  theme(
    text = element_text(size = 8), 
  )
p4_scatter_prime_format_h_b



# Combines somes plots onto one page.

leaveone_combine <- list(p3_Prime_loo_c_k, p3_Prime_loo_l_k, p3_Prime_loo_h_k, p3_Prime_loo_c_b, p3_Prime_loo_l_b, p3_Prime_loo_h_b)

leaveone_combine = lapply(leaveone_combine, ggplotGrob)

leaveoneout_plots_set<- grid.arrange(leaveone_combine,
nrow = 2 ,
top = "Leave_one_out_plots"
) 



p1_Prime_Single_c_k<- as_grob(p1_Prime_Single_c_k, device = NULL)

boxC <- list(box1,box2)
```

## Export Plots MR analsis

``` {r export_dataset_pairs}
# Plots for individual dataset pairs
# Single SNP plots

ggsave('results/Plots_all_data/mr_single_Forest_Prime_c_k.png', plot = p1_Prime_Single_c_k [[1]], units = 'in', height = 4, width = 9)

ggsave('results/Plots_all_data/mr_single_Forest_Prime_l_k.png', plot = p1_Prime_Single_l_k [[1]], units = 'in', height = 4, width = 9)


ggsave('results/Plots_all_data/mr_single_Forest_Prime_h_k.png', plot = p1_Prime_Single_h_k [[1]], units = 'in', height = 4, width = 9)


ggsave('results/Plots_all_data/mr_single_Forest_Prime_c_b.png', plot = p1_Prime_Single_c_b [[1]], units = 'in', height = 4, width = 9)


ggsave('results/Plots_all_data/mr_single_Forest_Prime_l_b.png', plot = p1_Prime_Single_l_b [[1]], units = 'in', height = 4, width = 9)


ggsave('results/Plots_all_data/mr_single_Forest_Prime_h_b.png', plot = p1_Prime_Single_h_b [[1]], units = 'in', height = 4, width = 9)




# Plots Leave one out for individual dataset pairs.

ggsave('results/Plots_all_data/mr_forest_loo_Prime_c_k.png', plot =p3_Prime_loo_c_k[[1]] + theme(text = element_text(size = 8)), units = 'in', height = 9, width = 9)


ggsave('results/Plots_all_data/mr_forest_loo_Prime_l_k.png', plot =p3_Prime_loo_l_k[[1]] + theme(text = element_text(size = 8)), units = 'in', height = 9, width = 9)


ggsave('results/Plots_all_data/mr_forest_loo_Prime_h_k.png', plot =p3_Prime_loo_h_k[[1]] + theme(text = element_text(size = 8)), units = 'in', height = 9, width = 9)



ggsave('results/Plots_all_data/mr_forest_loo_Prime_c_b.png', plot =p3_Prime_loo_c_b[[1]] + theme(text = element_text(size = 8)), units = 'in', height = 9, width = 9)


ggsave('results/Plots_all_data/mr_forest_loo_Prime_l_b.png', plot =p3_Prime_loo_l_b[[1]] + theme(text = element_text(size = 8)), units = 'in', height = 9, width = 9)


ggsave('results/Plots_all_data/mr_forest_loo_Prime_h_b.png', plot =p3_Prime_loo_h_b[[1]] + theme(text = element_text(size = 8)), units = 'in', height = 9, width = 9)

# Funnel plots for dataset pairs

ggsave('results/Plots_all_data/mr_funnel_Single_Prime_c_k.png', plot = P2_Prime_Funnel_single_c_k[[1]], units = 'in', height = 4, width = 9)


ggsave('results/Plots_all_data/mr_funnel_Single_Prime_l_k.png', plot = P2_Prime_Funnel_single_l_k[[1]], units = 'in', height = 4, width = 9)

ggsave('results/Plots_all_data/mr_funnel_Single_Prime_h_k.png', plot = P2_Prime_Funnel_single_h_k[[1]], units = 'in', height = 4, width = 9)

ggsave('results/Plots_all_data/mr_funnel_Single_Prime_c_b.png', plot = P2_Prime_Funnel_single_c_b[[1]], units = 'in', height = 4, width = 9)


ggsave('results/Plots_all_data/mr_funnel_Single_Prime_l_b.png', plot = P2_Prime_Funnel_single_l_b[[1]], units = 'in', height = 4, width = 9)


ggsave('results/Plots_all_data/mr_funnel_Single_Prime_h_b.png', plot = P2_Prime_Funnel_single_h_b[[1]], units = 'in', height = 4, width = 9)

# Plots Scatter for dataset pairs


ggsave('results/Plots_all_data/mr_scatter_Prime_c_k.png', plot = p4_scatter_prime_format_c_k, units = 'in', height = 4, width = 9)


ggsave('results/Plots_all_data/mr_scatter_Prime_l_k.png', plot = p4_scatter_prime_format_l_k, units = 'in', height = 4, width = 9)

ggsave('results/Plots_all_data/mr_scatter_Prime_h_k.png', plot = p4_scatter_prime_format_h_k, units = 'in', height = 4, width = 9)

ggsave('results/Plots_all_data/mr_scatter_Prime_c_b.png', plot = p4_scatter_prime_format_c_b, units = 'in', height = 4, width = 9)


ggsave('results/Plots_all_data/mr_scatter_Prime_l_b.png', plot = p4_scatter_prime_format_l_b, units = 'in', height = 4, width = 9)


ggsave('results/Plots_all_data/mr_scatter_Prime_h_b.png', plot = p4_scatter_prime_format_h_b, units = 'in', height = 4, width = 9)

```



## Export preliminary MR Results

```{r preliminary_results}
# MR Analysis
write_csv(Mr_mtDNAcn_AD_appended_Prime_OR , 'results/MR_Results_all_datasets/MR_Prime.csv')

# Sensitivity Tests

write_csv(Mr_H_Test_prime , 'results/Sensitivity_Tests_all_datasets/mr_het_res_Prime.csv')
write_csv(Mr_Horizontal_Test_prime , 'results/Sensitivity_Tests_all_datasets/mr_hor_plt_res_Prime.csv')



# Plots

ggsave('results/Plots_all_data/mr_single_Forest_Prime.png', plot = p1_Prime_Single [[1]], units = 'in', height = 4, width = 9)

ggsave('results/Plots_all_data/mr_funnel_Single_Prime.png', plot = P2_Prime_Funnel_single[[1]], units = 'in', height = 4, width = 9)
ggsave('results/Plots_all_data/mr_forest_loo_Prime.png', plot =p3_Prime_loo[[1]] + theme(text = element_text(size = 8)), units = 'in', height = 9, width = 9)
ggsave('results/Plots_all_data/mr_scatter_Prime.png', plot = p4_scatter_prime_format, units = 'in', height = 4, width = 9)

```




## APOE Region removed MR.
Removal of the Apoe region from both the Exposure and Outcome datasets before Radial MR
and MR analysis. 
Creates a new Binary variable apoe_region and filters dataset based on that column.
(Brian,Change all names to radial versions of the names.)

``` {r filter_exposures}

# Run the these to add binary columns marking SNP's as existing in the apoe region if they
# meet the specified creteria. These Apoe SNP's will be removed in the subsquent MR analysis.

# For Chong Kunkle

mtDNAcn_AD_Harm_chong_kunkle_apoeomit <-  filter(mtDNAcn_AD_Harm_chong_kunkle, !(chr.exposure == 19 & between(pos.exposure, 44912079, 45912079))) 

# For Longchamps Kunkle
mtDNAcn_AD_Harm_longchamps_kunkle_apoeomit <-  filter(mtDNAcn_AD_Harm_longchamps_kunkle, !(chr.exposure == 19 & between(pos.exposure, 44912079, 45912079))) 

# For Hagg Kunkle
mtDNAcn_AD_Harm_hagg_kunkle_apoeomit <-  filter(mtDNAcn_AD_Harm_hagg_kunkle, !(chr.exposure == 19 & between(pos.exposure, 44912079, 45912079))) 

# For Chong Bellenguez
mtDNAcn_AD_Harm_chong_bellenguez_apoeomit <-  filter(mtDNAcn_AD_Harm_chong_bellenguez, !(chr.exposure == 19 & between(pos.exposure, 44912079, 45912079))) 

# For Longchamps Bellenguez
mtDNAcn_AD_Harm_longchamps_bellenguez_apoeomit <-  filter(mtDNAcn_AD_Harm_longchamps_bellenguez, !(chr.exposure == 19 & between(pos.exposure, 44912079, 45912079))) 

# For Hagg Bellenguez
mtDNAcn_AD_Harm_hagg_bellenguez_apoeomit <-  filter(mtDNAcn_AD_Harm_hagg_bellenguez, !(chr.exposure == 19 & between(pos.exposure, 44912079, 45912079))) 




```

## Radial MR on apoe removed datasets to Remove Outliers in each dataset pair.

Radial MR will be ran first to Remove the presence of outliers.

``` {r}
#  Correction for Multiple testing bonferroni correction .05/nrow

# Chong Kunkle pair

AD_radial_ivw_mtDNAcn_chong_kunkle_apoeomit <- ivw_radial(mtDNAcn_AD_Harm_chong_kunkle_apoeomit, .05/nrow(mtDNAcn_AD_Harm_chong_kunkle_apoeomit), 2, .0001, TRUE)




#To view radial MR

AD_radial_ivw_mtDNAcn_chong_kunkle_apoeomit


# Static plot results of Radial MR

radial_p_chong_kunkle_ivw_apoeomit <- plot_radial(AD_radial_ivw_mtDNAcn_chong_kunkle_apoeomit, radial_scale = F, show_outliers = T)

radial_p_chong_kunkle_ivw_apoeomit


# longchamps Kunkle pair

AD_radial_ivw_mtDNAcn_longchamps_kunkle_apoeomit <- ivw_radial(mtDNAcn_AD_Harm_longchamps_kunkle_apoeomit, .05/nrow(mtDNAcn_AD_Harm_longchamps_kunkle_apoeomit), 2, .0001, TRUE)



AD_radial_ivw_mtDNAcn_longchamps_kunkle_apoeomit 


radial_p_longchamps_kunkle_ivw_apoeomit <- plot_radial(AD_radial_ivw_mtDNAcn_longchamps_kunkle_apoeomit, radial_scale = F, show_outliers = T)

radial_p_longchamps_kunkle_ivw_apoeomit


# Hagg Kunkle pair

AD_radial_ivw_mtDNAcn_hagg_kunkle_apoeomit <- ivw_radial(mtDNAcn_AD_Harm_hagg_kunkle_apoeomit, .05/nrow(mtDNAcn_AD_Harm_hagg_kunkle_apoeomit), 2, .0001, TRUE)



AD_radial_ivw_mtDNAcn_hagg_kunkle_apoeomit 


radial_p_hagg_kunkle_ivw_apoeomit <- plot_radial(AD_radial_ivw_mtDNAcn_hagg_kunkle_apoeomit, radial_scale = F, show_outliers = T)

radial_p_hagg_kunkle_ivw_apoeomit



# Chong Bellenguez pair

AD_radial_ivw_mtDNAcn_chong_bellenguez_apoeomit <- ivw_radial(mtDNAcn_AD_Harm_chong_bellenguez_apoeomit, .05/nrow(mtDNAcn_AD_Harm_chong_bellenguez_apoeomit), 2, .0001, TRUE)



AD_radial_ivw_mtDNAcn_chong_bellenguez_apoeomit 


radial_p_chong_bellenguez_ivw_apoeomit <- plot_radial(AD_radial_ivw_mtDNAcn_chong_bellenguez_apoeomit, radial_scale = F, show_outliers = T)

radial_p_chong_bellenguez_ivw_apoeomit


# longchamps Bellenguez pair

AD_radial_ivw_mtDNAcn_longchamps_bellenguez_apoeomit <- ivw_radial(mtDNAcn_AD_Harm_longchamps_bellenguez_apoeomit, .05/nrow(mtDNAcn_AD_Harm_longchamps_bellenguez_apoeomit), 2, .0001, TRUE)



AD_radial_ivw_mtDNAcn_longchamps_bellenguez_apoeomit 


radial_p_longchamps_bellenguez_ivw_apoeomit <- plot_radial(AD_radial_ivw_mtDNAcn_longchamps_bellenguez_apoeomit, radial_scale = F, show_outliers = T)

radial_p_longchamps_bellenguez_ivw_apoeomit


# Hagg Bellenguez pair

AD_radial_ivw_mtDNAcn_hagg_bellenguez_apoeomit <- ivw_radial(mtDNAcn_AD_Harm_hagg_bellenguez_apoeomit, .05/nrow(mtDNAcn_AD_Harm_hagg_bellenguez_apoeomit), 2, .0001, TRUE)



AD_radial_ivw_mtDNAcn_hagg_bellenguez_apoeomit 


radial_p_hagg_bellenguez_ivw_apoeomit <- plot_radial(AD_radial_ivw_mtDNAcn_hagg_bellenguez_apoeomit, radial_scale = F, show_outliers = T)

radial_p_hagg_bellenguez_ivw_apoeomit


outlier_plots_set1_apoeomit <- grid.arrange(radial_p_chong_kunkle_ivw_apoeomit, radial_p_longchamps_kunkle_ivw_apoeomit, radial_p_hagg_kunkle_ivw_apoeomit, 
nrow = 1,
top = "longchamps_kunkle",
left = "chong_kunkle",
right= "hagg_kunkle"
) 

outlier_plots_set2_apoeomit <- grid.arrange(radial_p_chong_bellenguez_ivw_apoeomit, radial_p_longchamps_bellenguez_ivw_apoeomit, radial_p_hagg_bellenguez_ivw_apoeomit, 
nrow = 1,
top = "longchamps_bellenguez",
left = "chong_bellenguez",
right= "hagg_bellenguez"
) 



```


## Export Radial Plots for all pairs

``` {r radial_outliers_export}


ggsave('results/Outliers_Plots_All/mr_radial_IVW_chong_kunkle_apoeomit.png', plot = radial_p_chong_kunkle_ivw_apoeomit, units = 'in', height = 4, width = 9)

ggsave('results/Outliers_Plots_All/mr_radial_IVW_longchamps_kunkle_apoeomit.png', plot = radial_p_longchamps_kunkle_ivw_apoeomit, units = 'in', height = 4, width = 9)

ggsave('results/Outliers_Plots_All/mr_radial_IVW_hagg_kunkle_apoeomit.png', plot = radial_p_hagg_kunkle_ivw_apoeomit, units = 'in', height = 4, width = 9)

ggsave('results/Outliers_Plots_All/mr_radial_IVW_chong_bellenguez_apoeomit.png', plot = radial_p_chong_bellenguez_ivw_apoeomit, units = 'in', height = 4, width = 9)

ggsave('results/Outliers_Plots_All/mr_radial_IVW_longchamps_bellenguez_apoeomit.png', plot = radial_p_longchamps_bellenguez_ivw_apoeomit, units = 'in', height = 4, width = 9)

ggsave('results/Outliers_Plots_All/mr_radial_IVW_hagg_bellenguez_apoeomit.png', plot = radial_p_hagg_bellenguez_ivw_apoeomit, units = 'in', height = 4, width = 9)

ggsave('results/Outliers_Plots_All/mr_radial_IVW_set1_apoeomit.png', plot = outlier_plots_set1_apoeomit, units = 'in', height = 4, width = 9)

ggsave('results/Outliers_Plots_All/mr_radial_IVW_set2_apoeomit.png', plot = outlier_plots_set2_apoeomit, units = 'in', height = 4, width = 9)

```




## Combining Radial data and harmonized with APOE regions removed.
(Brian, change the left join dataset from AD_radial_ivw_mtDNAcn_chong_kunkle to
data to AD_radial_ivw_mtDNAcn_chong_kunkle_apoeomit$data)
note it says apoe omit, change the name of the radial with apoe removed to that.

Also change mtDNAcn_AD_Harm_chong_kunkle to dataset with the apoe regions removed

``` {r radial_combination}

# This step creates a new dataset combining our radial data with our harmonized
# data, creating a dichotomous variable (Columns) of "outlier",
# so when conducting  MR on this dataset, omission of outlier 
# SNPS that may violate MR assumptions is possible.

# chong_kunkle

mtDNAcn_AD_Harm_Outlier_chong_kunkle_apoeomit <- mtDNAcn_AD_Harm_chong_kunkle_apoeomit %>%
  left_join(AD_radial_ivw_mtDNAcn_chong_kunkle_apoeomit$data) %>%
  mutate( gws.outcome = ifelse(pval.outcome <5e-8, TRUE, FALSE),
    mr_keep = ifelse(mr_keep == FALSE | Outliers == "Outlier" | gws.outcome == TRUE, FALSE, TRUE)
  )


# longchamps_kunkle

mtDNAcn_AD_Harm_Outlier_longchamps_kunkle_apoeomit <- mtDNAcn_AD_Harm_longchamps_kunkle_apoeomit %>%
  left_join(AD_radial_ivw_mtDNAcn_longchamps_kunkle_apoeomit$data) %>%
  mutate( gws.outcome = ifelse(pval.outcome <5e-8, TRUE, FALSE),
    mr_keep = ifelse(mr_keep == FALSE | Outliers == "Outlier" | gws.outcome == TRUE, FALSE, TRUE)
  )

# hagg_kunkle

mtDNAcn_AD_Harm_Outlier_hagg_kunkle_apoeomit <- mtDNAcn_AD_Harm_hagg_kunkle_apoeomit %>%
  left_join(AD_radial_ivw_mtDNAcn_hagg_kunkle_apoeomit$data) %>%
  mutate( gws.outcome = ifelse(pval.outcome <5e-8, TRUE, FALSE),
    mr_keep = ifelse(mr_keep == FALSE | Outliers == "Outlier" | gws.outcome == TRUE, FALSE, TRUE)
  )


# chong_bellenguez

mtDNAcn_AD_Harm_Outlier_chong_bellenguez_apoeomit <- mtDNAcn_AD_Harm_chong_bellenguez_apoeomit %>%
  left_join(AD_radial_ivw_mtDNAcn_chong_bellenguez_apoeomit$data) %>%
  mutate( gws.outcome = ifelse(pval.outcome <5e-8, TRUE, FALSE),
    mr_keep = ifelse(mr_keep == FALSE | Outliers == "Outlier" | gws.outcome == TRUE, FALSE, TRUE)
  )

# longchamps_bellenguez

mtDNAcn_AD_Harm_Outlier_longchamps_bellenguez_apoeomit <- mtDNAcn_AD_Harm_longchamps_bellenguez_apoeomit %>%
  left_join(AD_radial_ivw_mtDNAcn_longchamps_bellenguez_apoeomit$data) %>%
  mutate( gws.outcome = ifelse(pval.outcome <5e-8, TRUE, FALSE),
    mr_keep = ifelse(mr_keep == FALSE | Outliers == "Outlier" | gws.outcome == TRUE, FALSE, TRUE)
  )

# hagg_bellenguez

mtDNAcn_AD_Harm_Outlier_hagg_bellenguez_apoeomit <- mtDNAcn_AD_Harm_hagg_bellenguez_apoeomit %>%
  left_join(AD_radial_ivw_mtDNAcn_hagg_bellenguez_apoeomit$data) %>%
  mutate( gws.outcome = ifelse(pval.outcome <5e-8, TRUE, FALSE),
    mr_keep = ifelse(mr_keep == FALSE | Outliers == "Outlier" | gws.outcome == TRUE, FALSE, TRUE)
  )

```



# Appending Step apoe

``` {r append_all_data}

mtDNAcn_AD_harm_appended_apoe <- bind_rows(mtDNAcn_AD_Harm_Outlier_chong_kunkle_apoeomit, mtDNAcn_AD_Harm_Outlier_longchamps_kunkle_apoeomit, mtDNAcn_AD_Harm_Outlier_hagg_kunkle_apoeomit, mtDNAcn_AD_Harm_Outlier_chong_bellenguez_apoeomit, mtDNAcn_AD_Harm_Outlier_longchamps_bellenguez_apoeomit, mtDNAcn_AD_Harm_Outlier_hagg_bellenguez_apoeomit)


view(mtDNAcn_AD_harm_appended_apoe)

```

```{r export_harmonized_data}
write_csv(mtDNAcn_AD_harm_appended_apoe, 'data/Harmonized_Datasets/harmonized_data_appended_apoe_omit.csv')
# mtDNAcn_AD_harm_appended_apoe <- read_csv('data/Harmonized_Datasets/harmonized_data_appended_apoe_omit.csv')


```

## Apoe MR 

We run a Initial MR analysis for at a glance analysis on the SNPS being used in
MR analysis. Potential outliers not removed from this analysis

```{r MR_analysis_all_dataset_pairs}

Mr_mtDNAcn_AD_appended_apoe <- mr(mtDNAcn_AD_harm_appended_apoe, method_list = c("mr_ivw_fe", "mr_egger_regression", "mr_weighted_median", "mr_weighted_mode"))

Mr_mtDNAcn_AD_appended_apoe_OR <- generate_odds_ratios(Mr_mtDNAcn_AD_appended_apoe)

view (Mr_mtDNAcn_AD_appended_apoe_OR )



```

## Perform apoe Removed Sensentivity analysis

``` {r sensitivity}
# Heterogeneity (Test for Overall Violations of MR assumptions)
Mr_H_Test_apoe <- mr_heterogeneity(mtDNAcn_AD_harm_appended_apoe)

View(Mr_H_Test_apoe)

# (Test for Violations of MR assumptions via Horizontal pleiotrophy)
Mr_Horizontal_Test_apoe <- mr_pleiotropy_test(mtDNAcn_AD_harm_appended_apoe)

View(Mr_Horizontal_Test_apoe)






```


##  Plots MR APOE region removed.
Using seperate datasets to generate plots for each exposure outcome pair.

``` {r plots_}
# Chong_kunkle Preparation

MrAD_single_mtDNAcn_apoeomit_c_k <- mr_singlesnp(mtDNAcn_AD_Harm_Outlier_chong_kunkle_apoeomit)

MrAD_loo_mtDNAcn_apoeomit_c_k <- mr_leaveoneout(mtDNAcn_AD_Harm_Outlier_chong_kunkle_apoeomit)

#longchamps_kunkle Preparation

MrAD_single_mtDNAcn_apoeomit_l_k <- mr_singlesnp(mtDNAcn_AD_Harm_Outlier_longchamps_kunkle_apoeomit)

MrAD_loo_mtDNAcn_apoeomit_l_k <- mr_leaveoneout(mtDNAcn_AD_Harm_Outlier_longchamps_kunkle_apoeomit)


# hagg_kunkle Preparation

MrAD_single_mtDNAcn_apoeomit_h_k <- mr_singlesnp(mtDNAcn_AD_Harm_Outlier_hagg_kunkle_apoeomit)

MrAD_loo_mtDNAcn_apoeomit_h_k <- mr_leaveoneout(mtDNAcn_AD_Harm_Outlier_hagg_kunkle_apoeomit)

# chong_bellenguez Preparation
MrAD_single_mtDNAcn_apoeomit_c_b <- mr_singlesnp(mtDNAcn_AD_Harm_Outlier_chong_bellenguez_apoeomit)

MrAD_loo_mtDNAcn_apoeomit_c_b <- mr_leaveoneout(mtDNAcn_AD_Harm_Outlier_chong_bellenguez_apoeomit)

# longchamps_bellenguez Preparation

MrAD_single_mtDNAcn_apoeomit_l_b <- mr_singlesnp(mtDNAcn_AD_Harm_Outlier_longchamps_bellenguez_apoeomit)

MrAD_loo_mtDNAcn_apoeomit_l_b <- mr_leaveoneout(mtDNAcn_AD_Harm_Outlier_longchamps_bellenguez_apoeomit)

# hagg_bellenguez Preparation

MrAD_single_mtDNAcn_apoeomit_h_b <- mr_singlesnp(mtDNAcn_AD_Harm_Outlier_hagg_bellenguez_apoeomit)

MrAD_loo_mtDNAcn_apoeomit_h_b <- mr_leaveoneout(mtDNAcn_AD_Harm_Outlier_hagg_bellenguez_apoeomit)

# Generates Plots of single and leave one out anlysis MR

# Chong_kunkle Plots


p1_apoeomit_Single_c_k <- mr_forest_plot(MrAD_single_mtDNAcn_apoeomit_c_k )
p1_apoeomit_Single_c_k [[1]]

P2_apoeomit_Funnel_single_c_k <- mr_funnel_plot(MrAD_single_mtDNAcn_apoeomit_c_k)
P2_apoeomit_Funnel_single_c_k[[1]]

p3_apoeomit_loo_c_k <- mr_leaveoneout_plot(MrAD_loo_mtDNAcn_apoeomit_c_k)
p3_apoeomit_loo_c_k[[1]]


# Note, indivdual MR must be run in this step to create the plot

MR_scatter_mtDNAcn_ad <- mr(mtDNAcn_AD_Harm_Outlier_chong_kunkle_apoeomit)

p4_Prime_scatter_apoeomit <- mr_scatter_plot(MR_scatter_mtDNAcn_ad, mtDNAcn_AD_Harm_Outlier_chong_kunkle_apoeomit)

p4_scatter_apoeomit_format_c_k <- p4_Prime_scatter_apoeomit[[1]] + theme_bw() + 
  guides(color=guide_legend(ncol =1)) + 
  theme(
    text = element_text(size = 8), 
  )
p4_scatter_apoeomit_format_c_k


# longchamps_kunkle Plots

p1_apoeomit_Single_l_k <- mr_forest_plot(MrAD_single_mtDNAcn_apoeomit_l_k )
p1_apoeomit_Single_l_k [[1]]

P2_apoeomit_Funnel_single_l_k <- mr_funnel_plot(MrAD_single_mtDNAcn_apoeomit_l_k)
P2_apoeomit_Funnel_single_l_k[[1]]

p3_apoeomit_loo_l_k <- mr_leaveoneout_plot(MrAD_loo_mtDNAcn_apoeomit_l_k)
p3_apoeomit_loo_l_k[[1]]


MR_scatter_mtDNAcn_ad <- mr(mtDNAcn_AD_Harm_Outlier_longchamps_kunkle_apoeomit)

p4_apoeomit_scatter <- mr_scatter_plot(MR_scatter_mtDNAcn_ad, mtDNAcn_AD_Harm_Outlier_longchamps_kunkle_apoeomit)

p4_scatter_apoeomit_format_l_k <- p4_apoeomit_scatter[[1]] + theme_bw() + 
  guides(color=guide_legend(ncol =1)) + 
  theme(
    text = element_text(size = 8), 
  )
p4_scatter_apoeomit_format_l_k


# hagg_kunkle Plots

p1_apoeomit_Single_h_k <- mr_forest_plot(MrAD_single_mtDNAcn_apoeomit_h_k )
p1_apoeomit_Single_h_k [[1]]

P2_apoeomit_Funnel_single_h_k <- mr_funnel_plot(MrAD_single_mtDNAcn_apoeomit_h_k)
P2_apoeomit_Funnel_single_h_k[[1]]

p3_apoeomit_loo_h_k <- mr_leaveoneout_plot(MrAD_loo_mtDNAcn_apoeomit_h_k)
p3_apoeomit_loo_h_k[[1]]


MR_scatter_mtDNAcn_ad <- mr(mtDNAcn_AD_Harm_Outlier_hagg_kunkle_apoeomit)

p4_apoeomit_scatter <- mr_scatter_plot(MR_scatter_mtDNAcn_ad, mtDNAcn_AD_Harm_Outlier_hagg_kunkle_apoeomit)

p4_scatter_apoeomit_format_h_k <- p4_apoeomit_scatter[[1]] + theme_bw() + 
  guides(color=guide_legend(ncol =1)) + 
  theme(
    text = element_text(size = 8), 
  )
p4_scatter_apoeomit_format_h_k

# Chong_bellenguez Plots

p1_apoeomit_Single_c_b <- mr_forest_plot(MrAD_single_mtDNAcn_apoeomit_c_b )
p1_apoeomit_Single_c_b [[1]]

P2_apoeomit_Funnel_single_c_b <- mr_funnel_plot(MrAD_single_mtDNAcn_apoeomit_c_b)
P2_apoeomit_Funnel_single_c_b[[1]]

p3_apoeomit_loo_c_b <- mr_leaveoneout_plot(MrAD_loo_mtDNAcn_apoeomit_c_b)
p3_apoeomit_loo_c_b[[1]]

MR_scatter_mtDNAcn_ad <- mr(mtDNAcn_AD_Harm_Outlier_chong_bellenguez_apoeomit)

p4_Prime_scatter <- mr_scatter_plot(MR_scatter_mtDNAcn_ad, mtDNAcn_AD_Harm_Outlier_chong_bellenguez_apoeomit)

p4_scatter_apoeomit_format_c_b <- p4_Prime_scatter[[1]] + theme_bw() + 
  guides(color=guide_legend(ncol =1)) + 
  theme(
    text = element_text(size = 8), 
  )
p4_scatter_apoeomit_format_c_b

# longchamps_bellenguez Plots

p1_apoeomit_Single_l_b <- mr_forest_plot(MrAD_single_mtDNAcn_apoeomit_l_b )
p1_apoeomit_Single_l_b [[1]]

P2_apoeomit_Funnel_single_l_b <- mr_funnel_plot(MrAD_single_mtDNAcn_apoeomit_l_b)
P2_apoeomit_Funnel_single_l_b[[1]]

p3_apoeomit_loo_l_b <- mr_leaveoneout_plot(MrAD_loo_mtDNAcn_apoeomit_l_b)
p3_apoeomit_loo_l_b[[1]]


MR_scatter_mtDNAcn_ad <- mr(mtDNAcn_AD_Harm_Outlier_longchamps_bellenguez_apoeomit)

p4_apoeomit_scatter <- mr_scatter_plot(MR_scatter_mtDNAcn_ad, mtDNAcn_AD_Harm_Outlier_longchamps_bellenguez_apoeomit)

p4_scatter_apoeomit_format_l_b <- p4_apoeomit_scatter[[1]] + theme_bw() + 
  guides(color=guide_legend(ncol =1)) + 
  theme(
    text = element_text(size = 8), 
  )
p4_scatter_apoeomit_format_l_b

# hagg_bellenguez Plots

p1_apoeomit_Single_h_b <- mr_forest_plot(MrAD_single_mtDNAcn_apoeomit_h_b )
p1_apoeomit_Single_h_b [[1]]

P2_apoeomit_Funnel_single_h_b <- mr_funnel_plot(MrAD_single_mtDNAcn_apoeomit_h_b)
P2_apoeomit_Funnel_single_h_b[[1]]

p3_apoeomit_loo_h_b <- mr_leaveoneout_plot(MrAD_loo_mtDNAcn_apoeomit_h_b)
p3_apoeomit_loo_h_b[[1]]


MR_scatter_mtDNAcn_ad <- mr(mtDNAcn_AD_Harm_Outlier_hagg_bellenguez_apoeomit)

p4_apoeomit_scatter <- mr_scatter_plot(MR_scatter_mtDNAcn_ad, mtDNAcn_AD_Harm_Outlier_hagg_bellenguez_apoeomit)

p4_scatter_apoeomit_format_h_b <- p4_apoeomit_scatter[[1]] + theme_bw() + 
  guides(color=guide_legend(ncol =1)) + 
  theme(
    text = element_text(size = 8), 
  )
p4_scatter_apoeomit_format_h_b



```



``` {r combine_plots}
# Combines some plots onto one page.

leaveone_combine <- list(p3_Prime_loo_c_k, p3_Prime_loo_l_k, p3_Prime_loo_h_k, p3_Prime_loo_c_b, p3_Prime_loo_l_b, p3_Prime_loo_h_b)

leaveone_combine = lapply(leaveone_combine, ggplotGrob)

leaveoneout_plots_set<- grid.arrange(leaveone_combine,
nrow = 2 ,
top = "Leave_one_out_plots"
) 



p1_Prime_Single_c_k<- as_grob(p1_Prime_Single_c_k, device = NULL)

boxC <- list(box1,box2)
```
```
## Export Plots MR analsis


``` {r export_dataset_pairs}
# Plots for individual dataset pairs
# Single SNP plots

ggsave('Casual_Estimate_mtDNAcn_on_Alzheimer/results/Plots_all_data/mr_single_Forest_apoeomit_c_k.png', plot = p1_apoeomit_Single_c_k [[1]], units = 'in', height = 4, width = 9)

ggsave('Casual_Estimate_mtDNAcn_on_Alzheimer/results/Plots_all_data/mr_single_Forest_apoeomit_l_k.png', plot = p1_apoeomit_Single_l_k [[1]], units = 'in', height = 4, width = 9)


ggsave('Casual_Estimate_mtDNAcn_on_Alzheimer/results/Plots_all_data/mr_single_Forest_apoeomit_h_k.png', plot = p1_apoeomit_Single_h_k [[1]], units = 'in', height = 4, width = 9)


ggsave('Casual_Estimate_mtDNAcn_on_Alzheimer/results/Plots_all_data/mr_single_Forest_apoeomit_c_b.png', plot = p1_apoeomit_Single_c_b [[1]], units = 'in', height = 4, width = 9)


ggsave('Casual_Estimate_mtDNAcn_on_Alzheimer/results/Plots_all_data/mr_single_Forest_apoeomit_l_b.png', plot = p1_apoeomit_Single_l_b [[1]], units = 'in', height = 4, width = 9)


ggsave('Casual_Estimate_mtDNAcn_on_Alzheimer/results/Plots_all_data/mr_single_Forest_apoeomit_h_b.png', plot = p1_apoeomit_Single_h_b [[1]], units = 'in', height = 4, width = 9)




# Plots Leave one out for individual dataset pairs.

ggsave('Casual_Estimate_mtDNAcn_on_Alzheimer/results/Plots_all_data/mr_forest_loo_apoeomit_c_k.png', plot =p3_apoeomit_loo_c_k[[1]] + theme(text = element_text(size = 8)), units = 'in', height = 9, width = 9)


ggsave('Casual_Estimate_mtDNAcn_on_Alzheimer/results/Plots_all_data/mr_forest_loo_apoeomit_l_k.png', plot =p3_apoeomit_loo_l_k[[1]] + theme(text = element_text(size = 8)), units = 'in', height = 9, width = 9)


ggsave('Casual_Estimate_mtDNAcn_on_Alzheimer/results/Plots_all_data/mr_forest_loo_apoeomit_h_k.png', plot =p3_apoeomit_loo_h_k[[1]] + theme(text = element_text(size = 8)), units = 'in', height = 9, width = 9)



ggsave('Casual_Estimate_mtDNAcn_on_Alzheimer/results/Plots_all_data/mr_forest_loo_apoeomit_c_b.png', plot =p3_apoeomit_loo_c_b[[1]] + theme(text = element_text(size = 8)), units = 'in', height = 9, width = 9)


ggsave('results/Plots_all_data/mr_forest_loo_apoeomit_l_b.png', plot =p3_apoeomit_loo_l_b[[1]] + theme(text = element_text(size = 8)), units = 'in', height = 9, width = 9)


ggsave('Casual_Estimate_mtDNAcn_on_Alzheimer/results/Plots_all_data/mr_forest_loo_apoeomit_h_b.png', plot =p3_apoeomit_loo_h_b[[1]] + theme(text = element_text(size = 8)), units = 'in', height = 9, width = 9)

# Funnel plots for dataset pairs

ggsave('Casual_Estimate_mtDNAcn_on_Alzheimer/results/Plots_all_data/mr_funnel_Single_apoeomit_c_k.png', plot = P2_apoeomit_Funnel_single_c_k[[1]], units = 'in', height = 4, width = 9)


ggsave('Casual_Estimate_mtDNAcn_on_Alzheimer/results/Plots_all_data/mr_funnel_Single_apoeomit_l_k.png', plot = P2_apoeomit_Funnel_single_l_k[[1]], units = 'in', height = 4, width = 9)

ggsave('Casual_Estimate_mtDNAcn_on_Alzheimer/results/Plots_all_data/mr_funnel_Single_apoeomit_h_k.png', plot = P2_apoeomit_Funnel_single_h_k[[1]], units = 'in', height = 4, width = 9)

ggsave('Casual_Estimate_mtDNAcn_on_Alzheimer/results/Plots_all_data/mr_funnel_Single_apoeomit_c_b.png', plot = P2_apoeomit_Funnel_single_c_b[[1]], units = 'in', height = 4, width = 9)


ggsave('Casual_Estimate_mtDNAcn_on_Alzheimer/results/Plots_all_data/mr_funnel_Single_apoeomit_l_b.png', plot = P2_apoeomit_Funnel_single_l_b[[1]], units = 'in', height = 4, width = 9)


ggsave('Casual_Estimate_mtDNAcn_on_Alzheimer/results/Plots_all_data/mr_funnel_Single_apoeomit_h_b.png', plot = P2_apoeomit_Funnel_single_h_b[[1]], units = 'in', height = 4, width = 9)

# Plots Scatter for dataset pairs


ggsave('Casual_Estimate_mtDNAcn_on_Alzheimer/results/Plots_all_data/mr_scatter_apoeomit_c_k.png', plot = p4_scatter_apoeomit_format_c_k, units = 'in', height = 4, width = 9)


ggsave('Casual_Estimate_mtDNAcn_on_Alzheimer/results/Plots_all_data/mr_scatter_apoeomit_l_k.png', plot = p4_scatter_apoeomit_format_l_k, units = 'in', height = 4, width = 9)

ggsave('Casual_Estimate_mtDNAcn_on_Alzheimer/results/Plots_all_data/mr_scatter_apoeomit_h_k.png', plot = p4_scatter_apoeomit_format_h_k, units = 'in', height = 4, width = 9)

ggsave('Casual_Estimate_mtDNAcn_on_Alzheimer/results/Plots_all_data/mr_scatter_apoeomit_c_b.png', plot = p4_scatter_apoeomit_format_c_b, units = 'in', height = 4, width = 9)


ggsave('Casual_Estimate_mtDNAcn_on_Alzheimer/results/Plots_all_data/mr_scatter_apoeomit_l_b.png', plot = p4_scatter_apoeomit_format_l_b, units = 'in', height = 4, width = 9)


ggsave('Casual_Estimate_mtDNAcn_on_Alzheimer/results/Plots_all_data/mr_scatter_apoeomit_h_b.png', plot = p4_scatter_apoeomit_format_h_b, units = 'in', height = 4, width = 9)



```









## MR Metaanalysis Forest Plots


# 6 Pairs of MR: Forest Plot
```` {r apoe_retained} 

Mr_mtDNAcn_AD_appended_Prime_OR <- MR_Prime



MR_retained_forest_mod <- ggplot(data=Mr_mtDNAcn_AD_appended_Prime_OR, aes(x=method, y=or, ymin=or_lci95, ymax=or_uci95)) +
   facet_grid(cols = vars(outcome), rows = vars(exposure)) + 
geom_point(shape = 15, size = 2) +
geom_pointrange()+ 
  geom_hline(yintercept=1, lty=2, size =1) +  
  geom_errorbar(aes(ymin=or_lci95, ymax=or_uci95), width=0.5, cex=1)+ 
  coord_flip() + 
  geom_point(shape = 15, size = 2) +
  ggtitle("OR of AD by exposure/outcome pair:APOE Keep.") +
  xlab("Method")  +
  ylab("Odds Ratio AD due to mtDNAcn  (95% CI)")  
  

MR_retained_forest_mod


```


``` {r apoe_removed}

Mr_mtDNAcn_AD_appended_apoe_OR <- MR_apoe_omit

  
MR_apoe_omit_forest_plot_mod <- ggplot(data=Mr_mtDNAcn_AD_appended_apoe_OR, aes(x=method, y=or, ymin=or_lci95, ymax=or_uci95)) +
  facet_grid(cols = vars(outcome), rows = vars(exposure)) + 
geom_point(shape = 15, size = 2) +
geom_pointrange()+ 
  geom_hline(yintercept=1, lty=2, size =1) +  
  geom_errorbar(aes(ymin=or_lci95, ymax=or_uci95), width=0.5, cex=1)+ 
  coord_flip() + 
  geom_point(shape = 15, size = 2) +
  ggtitle("OR of AD by exposure/outcome pair:APOE Omit.") +
  xlab("Method")  +
  ylab("Odds Ratio AD due to mtDNAcn  (95% CI)") 
  
 MR_apoe_omit_forest_plot_mod
  
```

# Export forest plots of MR

``` {r forest_plots_combine_export}

ggsave('Casual_Estimate_mtDNAcn_on_Alzheimer/results/MR_Results_all_datasets/mr_retain_forestplot_all.png', plot = MR_retained_forest_mod)

ggsave('Casual_Estimate_mtDNAcn_on_Alzheimer/results/MR_Results_all_datasets/mr_omit_forestplot_all.png', plot = MR_apoe_omit_forest_plot_mod)

```

