---
title: "Mendelian Randomization mtDNA-CN on Alzheimer's Disease "
author: "Brian Alvarez"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}

# Note: Run this code block manually by highlighting and running the indivudual lines.
knitr::opts_chunk$set(echo = TRUE)

knitr::opts_knit$set(root.dir = 'C:/Users/brian/OneDrive/Documents/GitHub/Casual_Estimate_mtDNAcn_on_Alzheimer')


```


## Libraries

Loads the appropriate packages for MR analysis

``` {r eval=FALSE}
library(tidyverse) # Data wrangling 

library (usethis)

library(devtools)

library(ggman)

library(TwoSampleMR) # MR 

library(R.utils) #For the read_outcome_data step

library(RadialMR) # Radial MR sensitivity analysis 

library(phenoscanner)

library(LDlinkR) # LD and proxy snps

library(lhcMR)

library(janitor)
```


## Importing and formatting Exposure

This step imports our GWAS Exposure data sets(mtDNAcn) and formats them up for MR analysis


```{r eval=FALSE}
# Define column types for summary statistics GWAS
coltypes = cols(
  ID = col_character(),
  CHROM = col_double(),
  POS = col_double(),
  REF = col_character(),
  ALT = col_character(),
  AF = col_double(),
  TRAIT = col_character(),
  BETA = col_double(),
  SE = col_double(),
  Z = col_double(),
  P = col_double(),
  N = col_double(),
  OR = col_double(),
  OR_L95 = col_double(),
  OR_U95 = col_double(),
  DIR = col_character(),
  G1000_ID = col_character(),
  G1000_VARIANT = col_character(),
  DBSNP_ID = col_character(),
  DBSNP_VARIANT = col_character(),
  OLD_ID = col_character(),
  OLD_VARIANT = col_character()
)

```


``` {r read_in_chong}

exposure_path_chong = "Resources/Chong2022mtdnacn_eur.chrall.CPRA_b37.tsv.gz"

Exposure_mtDNAcn_chong <- read_tsv(exposure_path_chong,comment="#")


# To format the GWAS results for MR analysis Exposure. 

# Rename right-side variables to match column names of your GWAS dataset 
Format_mtDNAcn_chong <- Exposure_mtDNAcn_chong %>% format_data(.,
  type = "exposure",
  snps = NULL,
  header = TRUE,
  phenotype_col = "TRAIT",
  snp_col = "DBSNP_ID",
  beta_col = "BETA",
  se_col = "SE",
  eaf_col = "AF",
  effect_allele_col = "REF",
  other_allele_col = "ALT",
  pval_col = "P",
  units_col = "units",
  ncase_col = "N_CASES",
  ncontrol_col = "N_CTRLS",
  samplesize_col = "N",
  gene_col = "gene",
  min_pval = 1e-200,
  z_col = "Z",
  info_col = "info",
  chr_col = "CHROM",
  pos_col = "POS",
  log_pval = FALSE
)  %>%
  as_tibble()


```


``` {r read_in_longchamps}

exposure_path_longchamps = "Resources/Longchamps2021mtdnacn.chrall.CPRA_b37.tsv.gz"

Exposure_mtDNAcn_longchamps <- read_tsv(exposure_path_longchamps,comment="#")


# To format the GWAS results for MR analysis Exposure. 
# Rename right-side variables to match column names of your GWAS dataset.
Format_mtDNAcn_longchamps <- Exposure_mtDNAcn_longchamps %>% format_data(.,
  type = "exposure",
  snps = NULL,
  header = TRUE,
  phenotype_col = "TRAIT",
  snp_col = "DBSNP_ID",
  beta_col = "BETA",
  se_col = "SE",
  eaf_col = "AF",
  effect_allele_col = "REF",
  other_allele_col = "ALT",
  pval_col = "P",
  units_col = "units",
  ncase_col = "N_CASES",
  ncontrol_col = "N_CTRLS",
  samplesize_col = "N",
  gene_col = "gene",
  min_pval = 1e-200,
  z_col = "Z",
  info_col = "info",
  chr_col = "CHROM",
  pos_col = "POS",
  log_pval = FALSE
)  %>%
  as_tibble()



```



``` {r read_in_hagg}

exposure_path_hagg = "Resources/Hagg2020mtdnacn.chrall.CPRA_b37.tsv.gz"

Exposure_mtDNAcn_hagg <- read_tsv(exposure_path_hagg,comment="#")


# Format the GWAS results for MR analysis Exposure. 

# Rename right-side variables to match column names of your GWAS dataset.
Format_mtDNAcn_hagg <- Exposure_mtDNAcn_hagg %>% format_data(.,
  type = "exposure",
  snps = NULL,
  header = TRUE,
  phenotype_col = "TRAIT",
  snp_col = "DBSNP_ID",
  beta_col = "BETA",
  se_col = "SE",
  eaf_col = "AF",
  effect_allele_col = "REF",
  other_allele_col = "ALT",
  pval_col = "P",
  units_col = "units",
  ncase_col = "N_CASES",
  ncontrol_col = "N_CTRLS",
  samplesize_col = "N",
  gene_col = "gene",
  min_pval = 1e-200,
  z_col = "Z",
  info_col = "info",
  chr_col = "CHROM",
  pos_col = "POS",
  log_pval = FALSE
)  %>%
  as_tibble()

```


## Importing and formatting Outcome

This step imports our GWAS Outcome data sets and formats them up for MR analysis

``` {r read_in_Bellenguez}

outcome_path_bellenguez = "Resources/Bellenguez2022load.chrall.CPRA_b37.tsv.gz"

# Import Code for outcome GWAS 
OutcomeAD_bellenguez <- read_tsv(outcome_path_bellenguez,comment="#")

# Rename right-side variables to match column names of your GWAS dataset 

outcome_AD_Format_bellenguez <- OutcomeAD_bellenguez %>%
  format_data(.,
    type = "outcome",
    snps = NULL,
    header = TRUE,
    phenotype_col = "Phenotype",
    snp_col = "DBSNP_ID",
    beta_col = "BETA",
    se_col = "SE",
    eaf_col = "AF",
    effect_allele_col = "ALT",
    other_allele_col = "REF",
    pval_col = "P",
    samplesize_col = "N",
    z_col = "Z",
    chr_col = "CHROM",
    pos_col = "POS",
    log_pval = FALSE
) %>%
  as_tibble()

```



``` {r read_in_Kunkle}


outcome_path_kunkle = "Resources/Kunkle2019load_stage123.chrall.CPRA_b37.tsv.gz"

# Import Code for outcome GWAS 
OutcomeAD_kunkle <- read_tsv(outcome_path_kunkle,comment="#")

# Rename right-side variables to match column names of your GWAS dataset 

outcome_AD_Format_kunkle <- OutcomeAD_kunkle %>%
  format_data(.,
    type = "outcome",
    snps = NULL,
    header = TRUE,
    phenotype_col = "Phenotype",
    snp_col = "DBSNP_ID",
    beta_col = "BETA",
    se_col = "SE",
    eaf_col = "AF",
    effect_allele_col = "ALT",
    other_allele_col = "REF",
    pval_col = "P",
    samplesize_col = "N",
    z_col = "Z",
    chr_col = "CHROM",
    pos_col = "POS",
    log_pval = FALSE
) %>%
  as_tibble()



```


## Clumping Step
This step involving clumping and pruning to eliminate High linkage 
disequilibrium SNP (SNPS correlated with other SNPs). Removing High linkage 
disequilibrium SNP's are crucial to avoid "double counting" SNP's and
biasing our MR estimate.
The dataset will also be filter to remove a portion of insignficant SNP's for 
computational speed improvements.

``` {r eval=FALSE}

#Extract Instrumental variables from the exposure.
# Note that the lower Clump
#Clump R^2 is, the less linkage disequilibrium snps will
#be included. 
# Filter allows for more time efficient clumping, it 
# filters all of the significant SNPS and leaves all non significant 
# SNPs from the analysis.

# Filter Clumped SNPS will only include those SNPS whose P value is 
# Below 5e-8, Colname for P value is Pval.exposure in the filtered dataset.




# Chong Filter and Clumping
mtDNAcn_filter_chong <- filter(Format_mtDNAcn_chong , pval.exposure < 5e-8)


mtDNAcn_clump_chong <- clump_data(mtDNAcn_filter_chong, clump_kb = 10000,
                            clump_r2 = 0.001,
                            clump_p1 = 1,
                            clump_p2 = 1,
                            pop = "EUR"
)%>% 
  as_tibble()


# longchamps Filter and Clumping

mtDNAcn_filter_longchamps <- filter(Format_mtDNAcn_longchamps, pval.exposure < 5e-8)


mtDNAcn_clump_longchamps <- clump_data(mtDNAcn_filter_longchamps, clump_kb = 10000,
                            clump_r2 = 0.001,
                            clump_p1 = 1,
                            clump_p2 = 1,
                            pop = "EUR"
)%>% 
  as_tibble()


# hagg Filter and Clumping

mtDNAcn_filter_hagg <- filter(Format_mtDNAcn_hagg, pval.exposure < 5e-8)


mtDNAcn_clump_hagg <- clump_data(mtDNAcn_filter_hagg, clump_kb = 10000,
                            clump_r2 = 0.001,
                            clump_p1 = 1,
                            clump_p2 = 1,
                            pop = "EUR"
)%>% 
  as_tibble()



# Rename Data sets to be used in LD Proxy step

mtDNAcn_clump_chong_data <- mtDNAcn_clump_chong 

mtDNAcn_clump_longchamps_data <-  mtDNAcn_clump_longchamps

mtDNAcn_clump_hagg_data <-  mtDNAcn_clump_hagg

```


## LD Proxies
This attempts to retain any missing SNP's that are not both present in the 
exposure and outcome dataset by utilizing LD proxies in their place. This step will 
be done multiple times for each exposure > outcome pair.

Note: If an error is encounter in the query request, check "exp_snps_wo"
to see if there are any snps to query, if none exist (will return 0 rows), then no need to run LD proxies step.
Just run harmonzation on exposure outcome for that pair.
Example: No outcome snps were missing in the exposure for
longchamps_kunkle. So skip LD proxy step.


``` {r chong_kunkle_LDproxy}

# extract exposure SNPs present in outcome (outcome datasets,then exposure dataset)
outcome_clump <- semi_join(
  outcome_AD_Format_kunkle, mtDNAcn_clump_chong_data, by = "SNP"
)

# Exposure SNPs not present in outcome (exposure dataset, then outcome dataset)
exp_snps_wo <- anti_join(
  mtDNAcn_clump_chong_data, outcome_AD_Format_kunkle, by = "SNP"
)


# Use LDLinkR to identify proxy snps based on SNP dataset above
# that presents SNPs with no corresponding match in the other.
LDproxy_batch(exp_snps_wo$SNP, 
        pop = "CEU", 
        r2d = "r2", 
        token = 'a6deee62cc4a', 
        append = TRUE,
        genome_build = "grch37")
system("mv combined_query_snp_list_grch37.txt data/exposure_outcome_proxy_snps_chong_kunkle.txt")


# Munge proxy snp file
proxy_snps <- read_tsv("data/exposure_outcome_proxy_snps_chong_kunkle.txt", skip = 1, col_names = F) %>%
  rename(id = X1, func = X2, proxy_snp = X3, coord = X4, alleles = X5, maf = X6, 
         distance = X7, dprime = X8, rsq = X9, correlated_alleles = X10, FORGEdb = X11, RegulomeDB = X12) %>%
  separate(coord, c('chr', 'pos'), sep = ":") %>%
  mutate(snp = ifelse(id == 1, proxy_snp, NA), 
         chr = str_replace(chr, 'chr', ""), 
         chr = as.numeric(chr), 
         pos = as.numeric(pos)) %>%
  fill(snp, .direction = 'down') %>%
  relocate(snp, .before = proxy_snp) %>%
  dplyr::select(-id, -func, -FORGEdb, -RegulomeDB) %>%
  filter(rsq >= 0.8)

# Munge proxy snp and outcome data 
proxy_outcome <- left_join(
  proxy_snps, outcome_AD_Format_kunkle, by = c("proxy_snp" = "SNP")
) %>%
  separate(correlated_alleles, c("target_a1.outcome", "proxy_a1.outcome", 
                                 "target_a2.outcome", "proxy_a2.outcome"), sep = ",|=") %>%
  filter(!is.na(chr.outcome)) %>%
  arrange(snp, -rsq, abs(distance)) %>%
  group_by(snp) %>%
  slice(1) %>%
  ungroup() %>%
  mutate(
     proxy.outcome = TRUE,
     target_snp.outcome = snp,
     proxy_snp.outcome = proxy_snp, 
  ) %>% 
  mutate(
       new_effect_allele.outcome = case_when(
        proxy_a1.outcome == effect_allele.outcome & proxy_a2.outcome == other_allele.outcome ~ target_a1.outcome,
        proxy_a2.outcome == effect_allele.outcome & proxy_a1.outcome == other_allele.outcome ~ target_a2.outcome,
        TRUE ~ NA_character_
     ), 
      new_other_allele.outcome = case_when(
        proxy_a1.outcome == effect_allele.outcome & proxy_a2.outcome == other_allele.outcome ~ target_a2.outcome,
        proxy_a2.outcome == effect_allele.outcome & proxy_a1.outcome == other_allele.outcome ~ target_a1.outcome,
        TRUE ~ NA_character_
     ), 
     effect_allele.outcome = new_effect_allele.outcome, 
     other_allele.outcome = new_other_allele.outcome
  ) %>%
  dplyr::select(-proxy_snp, -chr, -pos, -alleles, -maf, -distance, -rsq, -dprime,  
         -new_effect_allele.outcome, -new_other_allele.outcome) %>%
  relocate(target_a1.outcome, proxy_a1.outcome, target_a2.outcome, proxy_a2.outcome, .after = proxy_snp.outcome) %>%
  rename(SNP = snp) %>%
  relocate(SNP, .after = samplesize.outcome)


# Merge outcome and proxy outcomes
chong_kunkle_dat_proxy <- bind_rows(
  outcome_clump, proxy_outcome
) %>% 
  arrange(chr.outcome, pos.outcome)

```

## No LD Proxies for longchamps_kunkle pair.
Code is shown for prosperity.

``` {r longchamps_kunkle_LDproxy}

# extract exposure SNPs present in outcome (outcome dataset,then exposure dataset)
outcome_clump <- semi_join(
  outcome_AD_Format_kunkle, mtDNAcn_clump_longchamps_data, by = "SNP"
)

# Exposure SNPs not present in outcome (exposure dataset, then outcome dataset)
exp_snps_wo <- anti_join(
  mtDNAcn_clump_longchamps_data, outcome_AD_Format_kunkle, by = "SNP"
)


# Use LDLinkR to identify proxy snps based on SNP dataset above
# that presents SNPs with no corresponding match in the other.
LDproxy_batch(exp_snps_wo$SNP, 
        pop = "CEU", 
        r2d = "r2", 
        token = 'a6deee62cc4a', 
        append = TRUE,
        genome_build = "grch37")
system("mv combined_query_snp_list_grch37.txt data/exposure_outcome_proxy_snps_longchamps_kunkle.txt")


# Munge proxy snp file
proxy_snps <- read_tsv("data/exposure_outcome_proxy_snps_longchamps_kunkle.txt", skip = 1, col_names = F) %>%
  rename(id = X1, func = X2, proxy_snp = X3, coord = X4, alleles = X5, maf = X6, 
         distance = X7, dprime = X8, rsq = X9, correlated_alleles = X10, FORGEdb = X11, RegulomeDB = X12) %>%
  separate(coord, c('chr', 'pos'), sep = ":") %>%
  mutate(snp = ifelse(id == 1, proxy_snp, NA), 
         chr = str_replace(chr, 'chr', ""), 
         chr = as.numeric(chr), 
         pos = as.numeric(pos)) %>%
  fill(snp, .direction = 'down') %>%
  relocate(snp, .before = proxy_snp) %>%
  dplyr::select(-id, -func, -FORGEdb, -RegulomeDB) %>%
  filter(rsq >= 0.8)

# Munge proxy snp and outcome data 
proxy_outcome <- left_join(
  proxy_snps, outcome_AD_Format_kunkle, by = c("proxy_snp" = "SNP")
) %>%
  separate(correlated_alleles, c("target_a1.outcome", "proxy_a1.outcome", 
                                 "target_a2.outcome", "proxy_a2.outcome"), sep = ",|=") %>%
  filter(!is.na(chr.outcome)) %>%
  arrange(snp, -rsq, abs(distance)) %>%
  group_by(snp) %>%
  slice(1) %>%
  ungroup() %>%
  mutate(
     proxy.outcome = TRUE,
     target_snp.outcome = snp,
     proxy_snp.outcome = proxy_snp, 
  ) %>% 
  mutate(
       new_effect_allele.outcome = case_when(
        proxy_a1.outcome == effect_allele.outcome & proxy_a2.outcome == other_allele.outcome ~ target_a1.outcome,
        proxy_a2.outcome == effect_allele.outcome & proxy_a1.outcome == other_allele.outcome ~ target_a2.outcome,
        TRUE ~ NA_character_
     ), 
      new_other_allele.outcome = case_when(
        proxy_a1.outcome == effect_allele.outcome & proxy_a2.outcome == other_allele.outcome ~ target_a2.outcome,
        proxy_a2.outcome == effect_allele.outcome & proxy_a1.outcome == other_allele.outcome ~ target_a1.outcome,
        TRUE ~ NA_character_
     ), 
     effect_allele.outcome = new_effect_allele.outcome, 
     other_allele.outcome = new_other_allele.outcome
  ) %>%
  dplyr::select(-proxy_snp, -chr, -pos, -alleles, -maf, -distance, -rsq, -dprime,  
         -new_effect_allele.outcome, -new_other_allele.outcome) %>%
  relocate(target_a1.outcome, proxy_a1.outcome, target_a2.outcome, proxy_a2.outcome, .after = proxy_snp.outcome) %>%
  rename(SNP = snp) %>%
  relocate(SNP, .after = samplesize.outcome)


# Merge outcome and proxy outcomes
longchamps_kunkle_dat_proxy <- bind_rows(
  outcome_clump, proxy_outcome
) %>% 
  arrange(chr.outcome, pos.outcome)

```



``` {r hagg_kunkle_LDproxy}


# extract exposure SNPs present in outcome (outcome datasets,then exposure dataset)
outcome_clump <- semi_join(
  outcome_AD_Format_kunkle, mtDNAcn_clump_hagg_data, by = "SNP"
)

# Exposure SNPs not present in outcome (exposure dataset, then outcome dataset)
exp_snps_wo <- anti_join(
  mtDNAcn_clump_hagg_data, outcome_AD_Format_kunkle, by = "SNP"
)


# Use LDLinkR to identify proxy snps based on SNP dataset above
# that presents SNPs with no corresponding match in the other.
LDproxy_batch(exp_snps_wo$SNP, 
        pop = "CEU", 
        r2d = "r2", 
        token = 'a6deee62cc4a', 
        append = TRUE,
        genome_build = "grch37")
system("mv combined_query_snp_list_grch37.txt data/exposure_outcome_proxy_snps_hagg_kunkle.txt")


# Munge proxy snp file
proxy_snps <- read_tsv("data/exposure_outcome_proxy_snps_hagg_kunkle.txt", skip = 1, col_names = F) %>%
  rename(id = X1, func = X2, proxy_snp = X3, coord = X4, alleles = X5, maf = X6, 
         distance = X7, dprime = X8, rsq = X9, correlated_alleles = X10, FORGEdb = X11, RegulomeDB = X12) %>%
  separate(coord, c('chr', 'pos'), sep = ":") %>%
  mutate(snp = ifelse(id == 1, proxy_snp, NA), 
         chr = str_replace(chr, 'chr', ""), 
         chr = as.numeric(chr), 
         pos = as.numeric(pos)) %>%
  fill(snp, .direction = 'down') %>%
  relocate(snp, .before = proxy_snp) %>%
  dplyr::select(-id, -func, -FORGEdb, -RegulomeDB) %>%
  filter(rsq >= 0.8)

# Munge proxy snp and outcome data 
proxy_outcome <- left_join(
  proxy_snps, outcome_AD_Format_kunkle, by = c("proxy_snp" = "SNP")
) %>%
  separate(correlated_alleles, c("target_a1.outcome", "proxy_a1.outcome", 
                                 "target_a2.outcome", "proxy_a2.outcome"), sep = ",|=") %>%
  filter(!is.na(chr.outcome)) %>%
  arrange(snp, -rsq, abs(distance)) %>%
  group_by(snp) %>%
  slice(1) %>%
  ungroup() %>%
  mutate(
     proxy.outcome = TRUE,
     target_snp.outcome = snp,
     proxy_snp.outcome = proxy_snp, 
  ) %>% 
  mutate(
       new_effect_allele.outcome = case_when(
        proxy_a1.outcome == effect_allele.outcome & proxy_a2.outcome == other_allele.outcome ~ target_a1.outcome,
        proxy_a2.outcome == effect_allele.outcome & proxy_a1.outcome == other_allele.outcome ~ target_a2.outcome,
        TRUE ~ NA_character_
     ), 
      new_other_allele.outcome = case_when(
        proxy_a1.outcome == effect_allele.outcome & proxy_a2.outcome == other_allele.outcome ~ target_a2.outcome,
        proxy_a2.outcome == effect_allele.outcome & proxy_a1.outcome == other_allele.outcome ~ target_a1.outcome,
        TRUE ~ NA_character_
     ), 
     effect_allele.outcome = new_effect_allele.outcome, 
     other_allele.outcome = new_other_allele.outcome
  ) %>%
  dplyr::select(-proxy_snp, -chr, -pos, -alleles, -maf, -distance, -rsq, -dprime,  
         -new_effect_allele.outcome, -new_other_allele.outcome) %>%
  relocate(target_a1.outcome, proxy_a1.outcome, target_a2.outcome, proxy_a2.outcome, .after = proxy_snp.outcome) %>%
  rename(SNP = snp) %>%
  relocate(SNP, .after = samplesize.outcome)


# Merge outcome and proxy outcomes
hagg_kunkle_dat_proxy <- bind_rows(
  outcome_clump, proxy_outcome
) %>% 
  arrange(chr.outcome, pos.outcome)

```


``` {r chong_bellenguez_LDproxy}


# extract exposure SNPs present in outcome (outcome datasets,then exposure dataset)
outcome_clump <- semi_join(
  outcome_AD_Format_bellenguez, mtDNAcn_clump_chong_data, by = "SNP"
)

# Exposure SNPs not present in outcome (exposure dataset, then outcome dataset)
exp_snps_wo <- anti_join(
  mtDNAcn_clump_chong_data, outcome_AD_Format_bellenguez, by = "SNP"
)


# Use LDLinkR to identify proxy snps based on SNP dataset above
# that presents SNPs with no corresponding match in the other.
LDproxy_batch(exp_snps_wo$SNP, 
        pop = "CEU", 
        r2d = "r2", 
        token = 'a6deee62cc4a', 
        append = TRUE,
        genome_build = "grch37")
system("mv combined_query_snp_list_grch37.txt data/exposure_outcome_proxy_snps_chong_bellenguez.txt")


# Munge proxy snp file
proxy_snps <- read_tsv("data/exposure_outcome_proxy_snps_chong_bellenguez.txt", skip = 1, col_names = F) %>%
  rename(id = X1, func = X2, proxy_snp = X3, coord = X4, alleles = X5, maf = X6, 
         distance = X7, dprime = X8, rsq = X9, correlated_alleles = X10, FORGEdb = X11, RegulomeDB = X12) %>%
  separate(coord, c('chr', 'pos'), sep = ":") %>%
  mutate(snp = ifelse(id == 1, proxy_snp, NA), 
         chr = str_replace(chr, 'chr', ""), 
         chr = as.numeric(chr), 
         pos = as.numeric(pos)) %>%
  fill(snp, .direction = 'down') %>%
  relocate(snp, .before = proxy_snp) %>%
  dplyr::select(-id, -func, -FORGEdb, -RegulomeDB) %>%
  filter(rsq >= 0.8)

# Munge proxy snp and outcome data 
proxy_outcome <- left_join(
  proxy_snps, outcome_AD_Format_bellenguez, by = c("proxy_snp" = "SNP")
) %>%
  separate(correlated_alleles, c("target_a1.outcome", "proxy_a1.outcome", 
                                 "target_a2.outcome", "proxy_a2.outcome"), sep = ",|=") %>%
  filter(!is.na(chr.outcome)) %>%
  arrange(snp, -rsq, abs(distance)) %>%
  group_by(snp) %>%
  slice(1) %>%
  ungroup() %>%
  mutate(
     proxy.outcome = TRUE,
     target_snp.outcome = snp,
     proxy_snp.outcome = proxy_snp, 
  ) %>% 
  mutate(
       new_effect_allele.outcome = case_when(
        proxy_a1.outcome == effect_allele.outcome & proxy_a2.outcome == other_allele.outcome ~ target_a1.outcome,
        proxy_a2.outcome == effect_allele.outcome & proxy_a1.outcome == other_allele.outcome ~ target_a2.outcome,
        TRUE ~ NA_character_
     ), 
      new_other_allele.outcome = case_when(
        proxy_a1.outcome == effect_allele.outcome & proxy_a2.outcome == other_allele.outcome ~ target_a2.outcome,
        proxy_a2.outcome == effect_allele.outcome & proxy_a1.outcome == other_allele.outcome ~ target_a1.outcome,
        TRUE ~ NA_character_
     ), 
     effect_allele.outcome = new_effect_allele.outcome, 
     other_allele.outcome = new_other_allele.outcome
  ) %>%
  dplyr::select(-proxy_snp, -chr, -pos, -alleles, -maf, -distance, -rsq, -dprime,  
         -new_effect_allele.outcome, -new_other_allele.outcome) %>%
  relocate(target_a1.outcome, proxy_a1.outcome, target_a2.outcome, proxy_a2.outcome, .after = proxy_snp.outcome) %>%
  rename(SNP = snp) %>%
  relocate(SNP, .after = samplesize.outcome)


# Merge outcome and proxy outcomes
chong_bellenguez_dat_proxy <- bind_rows(
  outcome_clump, proxy_outcome
) %>% 
  arrange(chr.outcome, pos.outcome)

```


``` {r longchamps_bellenguez_LDproxy}


# extract exposure SNPs present in outcome (outcome datasets,then exposure dataset)
outcome_clump <- semi_join(
  outcome_AD_Format_bellenguez, mtDNAcn_clump_longchamps_data, by = "SNP"
)

# Exposure SNPs not present in outcome (exposure dataset, then outcome dataset)
exp_snps_wo <- anti_join(
  mtDNAcn_clump_longchamps_data, outcome_AD_Format_bellenguez, by = "SNP"
)


# Use LDLinkR to identify proxy snps based on SNP dataset above
# that presents SNPs with no corresponding match in the other.
LDproxy_batch(exp_snps_wo$SNP, 
        pop = "CEU", 
        r2d = "r2", 
        token = 'a6deee62cc4a', 
        append = TRUE,
        genome_build = "grch37")
system("mv combined_query_snp_list_grch37.txt data/exposure_outcome_proxy_snps_longchamps_bellenguez.txt")


# Munge proxy snp file
proxy_snps <- read_tsv("data/exposure_outcome_proxy_snps_longchamps_bellenguez.txt", skip = 1, col_names = F) %>%
  rename(id = X1, func = X2, proxy_snp = X3, coord = X4, alleles = X5, maf = X6, 
         distance = X7, dprime = X8, rsq = X9, correlated_alleles = X10, FORGEdb = X11, RegulomeDB = X12) %>%
  separate(coord, c('chr', 'pos'), sep = ":") %>%
  mutate(snp = ifelse(id == 1, proxy_snp, NA), 
         chr = str_replace(chr, 'chr', ""), 
         chr = as.numeric(chr), 
         pos = as.numeric(pos)) %>%
  fill(snp, .direction = 'down') %>%
  relocate(snp, .before = proxy_snp) %>%
  dplyr::select(-id, -func, -FORGEdb, -RegulomeDB) %>%
  filter(rsq >= 0.8)

# Munge proxy snp and outcome data 
proxy_outcome <- left_join(
  proxy_snps, outcome_AD_Format_bellenguez, by = c("proxy_snp" = "SNP")
) %>%
  separate(correlated_alleles, c("target_a1.outcome", "proxy_a1.outcome", 
                                 "target_a2.outcome", "proxy_a2.outcome"), sep = ",|=") %>%
  filter(!is.na(chr.outcome)) %>%
  arrange(snp, -rsq, abs(distance)) %>%
  group_by(snp) %>%
  slice(1) %>%
  ungroup() %>%
  mutate(
     proxy.outcome = TRUE,
     target_snp.outcome = snp,
     proxy_snp.outcome = proxy_snp, 
  ) %>% 
  mutate(
       new_effect_allele.outcome = case_when(
        proxy_a1.outcome == effect_allele.outcome & proxy_a2.outcome == other_allele.outcome ~ target_a1.outcome,
        proxy_a2.outcome == effect_allele.outcome & proxy_a1.outcome == other_allele.outcome ~ target_a2.outcome,
        TRUE ~ NA_character_
     ), 
      new_other_allele.outcome = case_when(
        proxy_a1.outcome == effect_allele.outcome & proxy_a2.outcome == other_allele.outcome ~ target_a2.outcome,
        proxy_a2.outcome == effect_allele.outcome & proxy_a1.outcome == other_allele.outcome ~ target_a1.outcome,
        TRUE ~ NA_character_
     ), 
     effect_allele.outcome = new_effect_allele.outcome, 
     other_allele.outcome = new_other_allele.outcome
  ) %>%
  dplyr::select(-proxy_snp, -chr, -pos, -alleles, -maf, -distance, -rsq, -dprime,  
         -new_effect_allele.outcome, -new_other_allele.outcome) %>%
  relocate(target_a1.outcome, proxy_a1.outcome, target_a2.outcome, proxy_a2.outcome, .after = proxy_snp.outcome) %>%
  rename(SNP = snp) %>%
  relocate(SNP, .after = samplesize.outcome)


# Merge outcome and proxy outcomes
longchamps_bellenguez_dat_proxy <- bind_rows(
  outcome_clump, proxy_outcome
) %>% 
  arrange(chr.outcome, pos.outcome)

```


``` {r hagg_bellenguez_LDproxy}


# extract exposure SNPs present in outcome (outcome datasets,then exposure dataset)
outcome_clump <- semi_join(
  outcome_AD_Format_bellenguez, mtDNAcn_clump_hagg_data, by = "SNP"
)

# Exposure SNPs not present in outcome (exposure dataset, then outcome dataset)
exp_snps_wo <- anti_join(
  mtDNAcn_clump_hagg_data, outcome_AD_Format_bellenguez, by = "SNP"
)


# Use LDLinkR to identify proxy snps based on SNP dataset above
# that presents SNPs with no corresponding match in the other.
LDproxy_batch(exp_snps_wo$SNP, 
        pop = "CEU", 
        r2d = "r2", 
        token = 'a6deee62cc4a', 
        append = TRUE,
        genome_build = "grch37")
system("mv combined_query_snp_list_grch37.txt data/exposure_outcome_proxy_snps_hagg_bellenguez.txt")


# Munge proxy snp file
proxy_snps <- read_tsv("data/exposure_outcome_proxy_snps_hagg_bellenguez.txt", skip = 1, col_names = F) %>%
  rename(id = X1, func = X2, proxy_snp = X3, coord = X4, alleles = X5, maf = X6, 
         distance = X7, dprime = X8, rsq = X9, correlated_alleles = X10, FORGEdb = X11, RegulomeDB = X12) %>%
  separate(coord, c('chr', 'pos'), sep = ":") %>%
  mutate(snp = ifelse(id == 1, proxy_snp, NA), 
         chr = str_replace(chr, 'chr', ""), 
         chr = as.numeric(chr), 
         pos = as.numeric(pos)) %>%
  fill(snp, .direction = 'down') %>%
  relocate(snp, .before = proxy_snp) %>%
  dplyr::select(-id, -func, -FORGEdb, -RegulomeDB) %>%
  filter(rsq >= 0.8)

# Munge proxy snp and outcome data 
proxy_outcome <- left_join(
  proxy_snps, outcome_AD_Format_bellenguez, by = c("proxy_snp" = "SNP")
) %>%
  separate(correlated_alleles, c("target_a1.outcome", "proxy_a1.outcome", 
                                 "target_a2.outcome", "proxy_a2.outcome"), sep = ",|=") %>%
  filter(!is.na(chr.outcome)) %>%
  arrange(snp, -rsq, abs(distance)) %>%
  group_by(snp) %>%
  slice(1) %>%
  ungroup() %>%
  mutate(
     proxy.outcome = TRUE,
     target_snp.outcome = snp,
     proxy_snp.outcome = proxy_snp, 
  ) %>% 
  mutate(
       new_effect_allele.outcome = case_when(
        proxy_a1.outcome == effect_allele.outcome & proxy_a2.outcome == other_allele.outcome ~ target_a1.outcome,
        proxy_a2.outcome == effect_allele.outcome & proxy_a1.outcome == other_allele.outcome ~ target_a2.outcome,
        TRUE ~ NA_character_
     ), 
      new_other_allele.outcome = case_when(
        proxy_a1.outcome == effect_allele.outcome & proxy_a2.outcome == other_allele.outcome ~ target_a2.outcome,
        proxy_a2.outcome == effect_allele.outcome & proxy_a1.outcome == other_allele.outcome ~ target_a1.outcome,
        TRUE ~ NA_character_
     ), 
     effect_allele.outcome = new_effect_allele.outcome, 
     other_allele.outcome = new_other_allele.outcome
  ) %>%
  dplyr::select(-proxy_snp, -chr, -pos, -alleles, -maf, -distance, -rsq, -dprime,  
         -new_effect_allele.outcome, -new_other_allele.outcome) %>%
  relocate(target_a1.outcome, proxy_a1.outcome, target_a2.outcome, proxy_a2.outcome, .after = proxy_snp.outcome) %>%
  rename(SNP = snp) %>%
  relocate(SNP, .after = samplesize.outcome)


# Merge outcome and proxy outcomes
hagg_bellenguez_dat_proxy <- bind_rows(
  outcome_clump, proxy_outcome
) %>% 
  arrange(chr.outcome, pos.outcome)

```



## Harmonization and appending Step

Harmonze and combine all six pairs of analysis.

Chong_Kunkle
Longchamps_Kunkle
Hagg_Kunkle
Chong_Bellenguez
Longchamps_Bellenguez
Hagg_Bellenguez


``` {r eval=FALSE}
#  Change the two datasets names to ones created
# while performing clumping and LD proxies. Put proxy dataset in outcome slot (right)

# chong_kunkle_harmonization
mtDNAcn_AD_Harm_chong_kunkle <- harmonise_data(mtDNAcn_clump_chong, chong_kunkle_dat_proxy,action=2)


# longchamps_kunkle_harmonization (harmonization done with formatting dataset, no LD proxies found.)

mtDNAcn_AD_Harm_longchamps_kunkle <- harmonise_data(mtDNAcn_clump_longchamps, outcome_AD_Format_kunkle,action=2)

# hagg_kunkle_harmonization


mtDNAcn_AD_Harm_hagg_kunkle <- harmonise_data(mtDNAcn_clump_hagg, hagg_kunkle_dat_proxy,action=2)


# chong_bellenguez_harmonization

mtDNAcn_AD_Harm_chong_bellenguez <- harmonise_data(mtDNAcn_clump_chong, chong_bellenguez_dat_proxy,action=2)

# longchamps_bellenguez_harmonization

mtDNAcn_AD_Harm_longchamps_bellenguez <- harmonise_data(mtDNAcn_clump_longchamps, chong_bellenguez_dat_proxy,action=2)

# hagg_bellenguez_harmonization

mtDNAcn_AD_Harm_hagg_bellenguez <- harmonise_data(mtDNAcn_clump_hagg, hagg_bellenguez_dat_proxy,action=2)
```




## Radial MR to Remove Outliers in each dataset pair.

Radial MR will be ran first to Remove the presence of outliers.

``` {r}
#  Correction for Multiple testing bonferroni correction .05/nrow

# Chong Kunkle pair

AD_radial_ivw_mtDNAcn_chong_kunkle <- ivw_radial(mtDNAcn_AD_Harm_chong_kunkle, .05/nrow(mtDNAcn_AD_Harm_chong_kunkle), 2, .0001, TRUE)




#To view radial MR

AD_radial_ivw_mtDNAcn_chong_kunkle 


# Static plot results of Radial MR

radial_p_chong_kunkle_ivw <- plot_radial(AD_radial_ivw_mtDNAcn_chong_kunkle, radial_scale = F, show_outliers = T)

radial_p_chong_kunkle_ivw


# longchamps Kunkle pair

AD_radial_ivw_mtDNAcn_longchamps_kunkle <- ivw_radial(mtDNAcn_AD_Harm_longchamps_kunkle, .05/nrow(mtDNAcn_AD_Harm_longchamps_kunkle), 2, .0001, TRUE)



AD_radial_ivw_mtDNAcn_longchamps_kunkle 


radial_p_longchamps_kunkle_ivw <- plot_radial(AD_radial_ivw_mtDNAcn_longchamps_kunkle, radial_scale = F, show_outliers = T)

radial_p_longchamps_kunkle_ivw


# Hagg Kunkle pair

AD_radial_ivw_mtDNAcn_hagg_kunkle <- ivw_radial(mtDNAcn_AD_Harm_hagg_kunkle, .05/nrow(mtDNAcn_AD_Harm_hagg_kunkle), 2, .0001, TRUE)



AD_radial_ivw_mtDNAcn_hagg_kunkle 


radial_p_hagg_kunkle_ivw <- plot_radial(AD_radial_ivw_mtDNAcn_hagg_kunkle, radial_scale = F, show_outliers = T)

radial_p_hagg_kunkle_ivw



# Chong Bellenguez pair

AD_radial_ivw_mtDNAcn_chong_bellenguez <- ivw_radial(mtDNAcn_AD_Harm_chong_bellenguez, .05/nrow(mtDNAcn_AD_Harm_chong_bellenguez), 2, .0001, TRUE)



AD_radial_ivw_mtDNAcn_chong_bellenguez 


radial_p_chong_bellenguez_ivw <- plot_radial(AD_radial_ivw_mtDNAcn_chong_bellenguez, radial_scale = F, show_outliers = T)

radial_p_chong_bellenguez_ivw


# longchamps Bellenguez pair

AD_radial_ivw_mtDNAcn_longchamps_bellenguez <- ivw_radial(mtDNAcn_AD_Harm_longchamps_bellenguez, .05/nrow(mtDNAcn_AD_Harm_longchamps_bellenguez), 2, .0001, TRUE)



AD_radial_ivw_mtDNAcn_longchamps_bellenguez 


radial_p_longchamps_bellenguez_ivw <- plot_radial(AD_radial_ivw_mtDNAcn_longchamps_bellenguez, radial_scale = F, show_outliers = T)

radial_p_longchamps_bellenguez_ivw


# Hagg Bellenguez pair

AD_radial_ivw_mtDNAcn_hagg_bellenguez <- ivw_radial(mtDNAcn_AD_Harm_hagg_bellenguez, .05/nrow(mtDNAcn_AD_Harm_hagg_bellenguez), 2, .0001, TRUE)



AD_radial_ivw_mtDNAcn_hagg_bellenguez 


radial_p_hagg_bellenguez_ivw <- plot_radial(AD_radial_ivw_mtDNAcn_hagg_bellenguez, radial_scale = F, show_outliers = T)

radial_p_hagg_bellenguez_ivw

```


## Export Radial Plots for all pairs

``` {r radial_outliers_export}


ggsave('results/Outliers_Plots_All/mr_radial_IVW_chong_kunkle.png', plot = radial_p_chong_kunkle_ivw, units = 'in', height = 4, width = 9)

ggsave('results/Outliers_Plots_All/mr_radial_IVW_longchamps_kunkle.png', plot = radial_p_longchamps_kunkle_ivw, units = 'in', height = 4, width = 9)

ggsave('results/Outliers_Plots_All/mr_radial_IVW_hagg_kunkle.png', plot = radial_p_hagg_kunkle_ivw, units = 'in', height = 4, width = 9)

ggsave('results/Outliers_Plots_All/mr_radial_IVW_chong_bellenguez.png', plot = radial_p_chong_bellenguez_ivw, units = 'in', height = 4, width = 9)

ggsave('results/Outliers_Plots_All/mr_radial_IVW_longchamps_bellenguez.png', plot = radial_p_longchamps_bellenguez_ivw, units = 'in', height = 4, width = 9)

ggsave('results/Outliers_Plots_All/mr_radial_IVW_hagg_bellenguez.png', plot = radial_p_hagg_bellenguez_ivw, units = 'in', height = 4, width = 9)

```


## Combining Radial data and harmonized
``` {r radial_combination}

# This step creates a new dataset combining our radial data with our harmonized
# data, creating a dichotomous variable (Columns) of "outlier",
# so when conducting  MR on this dataset, omission of outlier 
# SNPS that may violate MR assumptions is possible.

# chong_kunkle

mtDNAcn_AD_Harm_Outlier_chong_kunkle <- mtDNAcn_AD_Harm_chong_kunkle %>%
  left_join(AD_radial_ivw_mtDNAcn_chong_kunkle$data) %>%
  mutate( gws.outcome = ifelse(pval.outcome <5e-8, TRUE, FALSE),
    mr_keep = ifelse(mr_keep == FALSE | Outliers == "Outlier" | gws.outcome == TRUE, FALSE, TRUE)
  )


# longchamps_kunkle

mtDNAcn_AD_Harm_Outlier_longchamps_kunkle <- mtDNAcn_AD_Harm_longchamps_kunkle %>%
  left_join(AD_radial_ivw_mtDNAcn_longchamps_kunkle$data) %>%
  mutate( gws.outcome = ifelse(pval.outcome <5e-8, TRUE, FALSE),
    mr_keep = ifelse(mr_keep == FALSE | Outliers == "Outlier" | gws.outcome == TRUE, FALSE, TRUE)
  )

# hagg_kunkle

mtDNAcn_AD_Harm_Outlier_hagg_kunkle <- mtDNAcn_AD_Harm_hagg_kunkle %>%
  left_join(AD_radial_ivw_mtDNAcn_hagg_kunkle$data) %>%
  mutate( gws.outcome = ifelse(pval.outcome <5e-8, TRUE, FALSE),
    mr_keep = ifelse(mr_keep == FALSE | Outliers == "Outlier" | gws.outcome == TRUE, FALSE, TRUE)
  )


# chong_bellenguez

mtDNAcn_AD_Harm_Outlier_chong_bellenguez <- mtDNAcn_AD_Harm_chong_bellenguez %>%
  left_join(AD_radial_ivw_mtDNAcn_chong_bellenguez$data) %>%
  mutate( gws.outcome = ifelse(pval.outcome <5e-8, TRUE, FALSE),
    mr_keep = ifelse(mr_keep == FALSE | Outliers == "Outlier" | gws.outcome == TRUE, FALSE, TRUE)
  )

# longchamps_bellenguez

mtDNAcn_AD_Harm_Outlier_longchamps_bellenguez <- mtDNAcn_AD_Harm_longchamps_bellenguez %>%
  left_join(AD_radial_ivw_mtDNAcn_longchamps_bellenguez$data) %>%
  mutate( gws.outcome = ifelse(pval.outcome <5e-8, TRUE, FALSE),
    mr_keep = ifelse(mr_keep == FALSE | Outliers == "Outlier" | gws.outcome == TRUE, FALSE, TRUE)
  )

# hagg_bellenguez

mtDNAcn_AD_Harm_Outlier_hagg_bellenguez <- mtDNAcn_AD_Harm_hagg_bellenguez %>%
  left_join(AD_radial_ivw_mtDNAcn_hagg_bellenguez$data) %>%
  mutate( gws.outcome = ifelse(pval.outcome <5e-8, TRUE, FALSE),
    mr_keep = ifelse(mr_keep == FALSE | Outliers == "Outlier" | gws.outcome == TRUE, FALSE, TRUE)
  )

```




# Appending Step

``` {r append_all_data}

mtDNAcn_AD_harm_appended <- bind_rows(mtDNAcn_AD_Harm_Outlier_chong_kunkle, mtDNAcn_AD_Harm_Outlier_longchamps_kunkle, mtDNAcn_AD_Harm_Outlier_hagg_kunkle, mtDNAcn_AD_Harm_Outlier_chong_bellenguez, mtDNAcn_AD_Harm_Outlier_longchamps_bellenguez, mtDNAcn_AD_Harm_Outlier_hagg_bellenguez)


view(mtDNAcn_AD_harm_appended)

```

```{r export_harmonized_data}
write_csv(mtDNAcn_AD_harm_appended, 'data/Harmonized_Datasets/harmonized_data_appended.csv')
# mtDNAcn_AD_harm_appended <- read_csv('data/Harmonized_Datasets/harmonized_data_appended.csv')


```


## Preliminary MR 

We run a Initial MR analysis for at a glance analysis on the SNPS being used in
MR analysis. Potential outliers not removed from this analysis

```{r MR_analysis_all_dataset_pairs}

Mr_mtDNAcn_AD_appended_Prime <- mr(mtDNAcn_AD_harm_appended, method_list = c("mr_ivw_fe", "mr_egger_regression", "mr_weighted_median", "mr_weighted_mode"))

Mr_mtDNAcn_AD_appended_Prime_OR <- generate_odds_ratios(Mr_mtDNAcn_AD_appended_Prime)

view (Mr_mtDNAcn_AD_appended_Prime_OR )



```

## Perform Outlier Removed Sensentivity analysis

``` {r sensitivity}
# Heterogeneity (Test for Overall Violations of MR assumptions)
Mr_H_Test_prime <- mr_heterogeneity(mtDNAcn_AD_harm_appended)

View(Mr_H_Test_prime)

# (Test for Violations of MR assumptions via Horizontal pleiotrophy)
Mr_Horizontal_Test_prime <- mr_pleiotropy_test(mtDNAcn_AD_harm_appended)

View(Mr_Horizontal_Test_prime)

# Directionality test (testing whether Exposure >Outcome direction is true casual direction)


#This may no longer be necessary
AD_direction_Outlier_mtDNAcn_chong <- directionality_test(mtDNAcn_AD_Harm_Outlier_chong)

View(AD_direction_Outlier_mtDNAcn_chong)

```


## Preliminary Plots
``` {r Preliminary_plots}

MrAD_single_mtDNAcn_Prime <- mr_singlesnp(mtDNAcn_AD_harm_appended)

MrAD_loo_mtDNAcn_Prime <- mr_leaveoneout(mtDNAcn_AD_harm_appended)

# Generates Plots of single and leave one out anlysis MR.

p1_Prime_Single <- mr_forest_plot(MrAD_single_mtDNAcn_Prime )
p1_Prime_Single [[1]]

P2_Prime_Funnel_single <- mr_funnel_plot(MrAD_single_mtDNAcn_Prime)
P2_Prime_Funnel_single[[1]]

p3_Prime_loo <- mr_leaveoneout_plot(MrAD_loo_mtDNAcn_Prime)
p3_Prime_loo[[1]]

# To plot preliminary MR analysis scatter plot, must rerun and rename.

MR_scatter_mtDNAcn_ad <- mr(mtDNAcn_AD_harm_appended)

p4_Prime_scatter <- mr_scatter_plot(MR_scatter_mtDNAcn_ad, mtDNAcn_AD_harm_appended)

p4_scatter_prime_format <- p4_Prime_scatter[[1]] + theme_bw() + 
  guides(color=guide_legend(ncol =1)) + 
  theme(
    text = element_text(size = 8), 
  )
p4_scatter_prime_format

```


## Export preliminary MR Results

```{r preliminary_results}
# MR Analysis
write_csv(Mr_mtDNAcn_AD_appended_Prime_OR , 'results/MR_Results_all datasets/MR_Prime.csv')

# Sensitivity Tests

write_csv(Mr_H_Test_prime , 'results/Sensitivity_Tests_all_datasets/mr_het_res_Prime.csv')
write_csv(Mr_Horizontal_Test_prime , 'results/Sensitivity_Tests_all_datasets/mr_hor_plt_res_Prime.csv')


#Code for directional effects may be removed (i did not run this code)
write_csv(AD_direction_Prime_mtDNAcn_chong , 'results/Sensitivity_Tests_Chong/Directional_Chong_Prime.csv')

# Plots

ggsave('results/Plots_all_data/mr_single_Forest_Prime.png', plot = p1_Prime_Single [[1]], units = 'in', height = 4, width = 9)

ggsave('results/Plots_all_data/mr_funnel_Single_Prime.png', plot = P2_Prime_Funnel_single[[1]], units = 'in', height = 4, width = 9)
ggsave('results/Plots_all_data/mr_forest_loo_Prime.png', plot =p3_Prime_loo[[1]] + theme(text = element_text(size = 8)), units = 'in', height = 9, width = 9)
ggsave('results/Plots_all_data/mr_scatter_Prime.png', plot = p4_scatter_prime_format, units = 'in', height = 4, width = 9)

```




## Filter APOE Region.
Removal of the Apoe region from both the Exposure and Outcome datasets. Creates
a new Binary variable apoe_region and filters dataset based on that column.

``` {r filter_exposures}

# Run the these to add binary columns marking SNP's as existing in the apoe region if they
# meet the specified creteria. These Apoe SNP's will be removed in the subsquent MR analysis.

# For Chong Kunkle

chong_kunkle_omit_apoe <- mtDNAcn_AD_Harm_chong_kunkle %>%
  left_join(AD_radial_ivw_mtDNAcn_chong_kunkle$data) %>%
  mutate(
    apoe_region = case_when(
      chr.outcome == 19 & between(pos.outcome,44912079,45912079) ~ TRUE,
      TRUE ~ FALSE
    ),
    gws.outcome = ifelse(pval.outcome <5e-8, TRUE, FALSE),
    mr_keep = ifelse(mr_keep == FALSE | Outliers == "Outlier" | apoe_region == TRUE | gws.outcome == TRUE, FALSE, TRUE)
  )



# For Longchamps Kunkle

longchamps_kunkle_omit_apoe <- mtDNAcn_AD_Harm_longchamps_kunkle %>%
  left_join(AD_radial_ivw_mtDNAcn_longchamps_kunkle$data) %>%
  mutate(
    apoe_region = case_when(
      chr.outcome == 19 & between(pos.outcome,44912079,45912079) ~ TRUE,
      TRUE ~ FALSE
    ),
    gws.outcome = ifelse(pval.outcome <5e-8, TRUE, FALSE),
    mr_keep = ifelse(mr_keep == FALSE | Outliers == "Outlier" | apoe_region == TRUE | gws.outcome == TRUE, FALSE, TRUE)
  )

# For Hagg Kunkle

hagg_kunkle_omit_apoe <- mtDNAcn_AD_Harm_hagg_kunkle %>%
  left_join(AD_radial_ivw_mtDNAcn_hagg_kunkle$data) %>%
  mutate(
    apoe_region = case_when(
      chr.outcome == 19 & between(pos.outcome,44912079,45912079) ~ TRUE,
      TRUE ~ FALSE
    ),
    gws.outcome = ifelse(pval.outcome <5e-8, TRUE, FALSE),
    mr_keep = ifelse(mr_keep == FALSE | Outliers == "Outlier" | apoe_region == TRUE | gws.outcome == TRUE, FALSE, TRUE)
  )


# For Chong Bellenguez

chong_bellenguez_omit_apoe <- mtDNAcn_AD_Harm_chong_bellenguez %>%
  left_join(AD_radial_ivw_mtDNAcn_chong_bellenguez$data) %>%
  mutate(
    apoe_region = case_when(
      chr.outcome == 19 & between(pos.outcome,44912079,45912079) ~ TRUE,
      TRUE ~ FALSE
    ),
    gws.outcome = ifelse(pval.outcome <5e-8, TRUE, FALSE),
    mr_keep = ifelse(mr_keep == FALSE | Outliers == "Outlier" | apoe_region == TRUE | gws.outcome == TRUE, FALSE, TRUE)
  )

# For Longchamps Bellenguez

longchamps_bellenguez_omit_apoe <- mtDNAcn_AD_Harm_longchamps_bellenguez %>%
  left_join(AD_radial_ivw_mtDNAcn_longchamps_bellenguez$data) %>%
  mutate(
    apoe_region = case_when(
      chr.outcome == 19 & between(pos.outcome,44912079,45912079) ~ TRUE,
      TRUE ~ FALSE
    ),
    gws.outcome = ifelse(pval.outcome <5e-8, TRUE, FALSE),
    mr_keep = ifelse(mr_keep == FALSE | Outliers == "Outlier" | apoe_region == TRUE | gws.outcome == TRUE, FALSE, TRUE)
  )

# For Hagg Bellenguez

hagg_bellenguez_omit_apoe <- mtDNAcn_AD_Harm_hagg_bellenguez %>%
  left_join(AD_radial_ivw_mtDNAcn_hagg_bellenguez$data) %>%
  mutate(
    apoe_region = case_when(
      chr.outcome == 19 & between(pos.outcome,44912079,45912079) ~ TRUE,
      TRUE ~ FALSE
    ),
    gws.outcome = ifelse(pval.outcome <5e-8, TRUE, FALSE),
    mr_keep = ifelse(mr_keep == FALSE | Outliers == "Outlier" | apoe_region == TRUE | gws.outcome == TRUE, FALSE, TRUE)
  )


```


```{r export_harmonized_data}
write_csv(mtDNAcn_AD_harm_appended_apoe, 'data/Harmonized_Datasets/harmonized_data_appended_apoe_omit.csv')
# mtDNAcn_AD_harm_appended_apoe <- read_csv('data/Harmonized_Datasets/harmonized_data_appended_apoe_omit.csv')


```

## Apoe MR 

We run a Initial MR analysis for at a glance analysis on the SNPS being used in
MR analysis. Potential outliers not removed from this analysis

```{r MR_analysis_all_dataset_pairs}

Mr_mtDNAcn_AD_appended_apoe <- mr(mtDNAcn_AD_harm_appended_apoe, method_list = c("mr_ivw_fe", "mr_egger_regression", "mr_weighted_median", "mr_weighted_mode"))

Mr_mtDNAcn_AD_appended_apoe_OR <- generate_odds_ratios(Mr_mtDNAcn_AD_appended_apoe)

view (Mr_mtDNAcn_AD_appended_apoe_OR )



```

## Perform apoe Removed Sensentivity analysis

``` {r sensitivity}
# Heterogeneity (Test for Overall Violations of MR assumptions)
Mr_H_Test_apoe <- mr_heterogeneity(mtDNAcn_AD_harm_appended_apoe)

View(Mr_H_Test_apoe)

# (Test for Violations of MR assumptions via Horizontal pleiotrophy)
Mr_Horizontal_Test_apoe <- mr_pleiotropy_test(mtDNAcn_AD_harm_appended_apoe)

View(Mr_Horizontal_Test_apoe)

# Directional test (testing whether Exposure >Outcome direction is true casual direction)


#This may no longer be necessary
AD_direction_apoe <- directionality_test(Mr_mtDNAcn_AD_appended_apoe)

View(AD_direction_apoe)

```


## Apoe removed Plots
``` {r Preliminary_plots}

MrAD_single_mtDNAcn_apoe <- mr_singlesnp(mtDNAcn_AD_harm_appended_apoe)

MrAD_loo_mtDNAcn_apoe <- mr_leaveoneout(mtDNAcn_AD_harm_appended_apoe)

# Generates Plots of single and leave one out anlysis MR.

p1_apoe_Single <- mr_forest_plot(MrAD_single_mtDNAcn_apoe )
p1_apoe_Single [[1]]

P2_apoe_Funnel_single <- mr_funnel_plot(MrAD_single_mtDNAcn_apoe)
P2_apoe_Funnel_single[[1]]

p3_apoe_loo <- mr_leaveoneout_plot(MrAD_loo_mtDNAcn_apoe)
p3_apoe_loo[[1]]

# To plot preliminary MR analysis scatter plot, must rerun and rename.

MR_scatter_mtDNAcn_ad_apoe <- mr(mtDNAcn_AD_harm_appended_apoe)

p4_apoe_scatter <- mr_scatter_plot(MR_scatter_mtDNAcn_ad_apoe, mtDNAcn_AD_harm_appended_apoe)

p4_scatter_apoe_format <- p4_apoe_scatter[[1]] + theme_bw() + 
  guides(color=guide_legend(ncol =1)) + 
  theme(
    text = element_text(size = 8), 
  )
p4_scatter_apoe_format

```


## Export apoe removed MR Results

```{r preliminary_results}
# MR Analysis
write_csv(Mr_mtDNAcn_AD_appended_apoe_OR , 'results/MR_Results_all datasets/MR_apoe_omit.csv')

# Sensitivity Tests

write_csv(Mr_H_Test_apoe , 'results/Sensitivity_Tests_all_datasets/mr_het_res_apoe_omit.csv')
write_csv(Mr_Horizontal_Test_apoe , 'results/Sensitivity_Tests_all_datasets/mr_hor_plt_res_apoe_omit.csv')


#Code for directional effects may be removed (i did not run this code)
write_csv(AD_direction_Prime_mtDNAcn_chong , 'results/Sensitivity_Tests_Chong/Directional_Chong_Prime.csv')

# Plots

ggsave('results/Plots_all_data/mr_single_Forest_apoe_omit.png', plot = p1_apoe_Single [[1]], units = 'in', height = 4, width = 9)

ggsave('results/Plots_all_data/mr_funnel_Single_apoe_omit.png', plot = P2_apoe_Funnel_single[[1]], units = 'in', height = 4, width = 9)
ggsave('results/Plots_all_data/mr_forest_loo_apoe_omit.png', plot =p3_apoe_loo[[1]] + theme(text = element_text(size = 8)), units = 'in', height = 9, width = 9)
ggsave('results/Plots_all_data/mr_scatter_apoe_omit.png', plot = p4_scatter_apoe_format, units = 'in', height = 4, width = 9)

```





