---
title: "MVMR_chong_kunkle"
output: html_document
date: "2024-08-22"
---

## Libraries

Loads the appropriate packages for MR analysis

``` {r eval=FALSE}
library(tidyverse) # Data wrangling 

library(cowplot)

library(ggman)

library(TwoSampleMR) # MR 

library(RadialMR) # Radial MR sensitivity analysis 

library(phenoscanner)

library(LDlinkR) # LD and proxy snps

library(janitor)

library(MVMR)         # Multivarble MR

#library(RMVMR)

#MVMR Munge proxies
source("/wynton/home/andrews/achatterjee/scripts/mvmr_munge_proxies.R")

#Munge proxies
source("/wynton/home/andrews/achatterjee/scripts/munge_proxies.R")

# Define column types for summary statistics (MDNA and AD GWAS)
coltypes = cols(
  ID = col_character(),
  CHROM = col_double(),
  POS = col_double(),
  REF = col_character(),
  ALT = col_character(),
  AF = col_double(),
  TRAIT = col_character(),
  BETA = col_double(),
  SE = col_double(),
  Z = col_double(),
  P = col_double(),
  N = col_double(),
  OR = col_double(),
  OR_L95 = col_double(),
  OR_U95 = col_double(),
  DIR = col_character(),
  G1000_ID = col_character(),
  G1000_VARIANT = col_character(),
  DBSNP_ID = col_character(),
  DBSNP_VARIANT = col_character(),
  OLD_ID = col_character(),
  OLD_VARIANT = col_character()
)

```

## Exposure 1: Chong
``` {r read_in_chong}

exposure_path_chong = "/wynton/group/andrews/data/summary_stats/results/Chong2022mtdnacn_eur.chrall.CPRA_b37.tsv"

mtDNAcn_chong <- read_tsv(exposure_path_chong,comment="#")


# To format the GWAS results for MR analysis Exposure. 
# Rename right-side variables to match column names of your GWAS dataset.
Format_mtDNAcn_chong <- mtDNAcn_chong %>% format_data(.,
  type = "exposure",
  snps = NULL,
  header = TRUE,
  phenotype_col = "TRAIT",
  snp_col = "DBSNP_ID",
  beta_col = "BETA",
  se_col = "SE",
  eaf_col = "AF",
  effect_allele_col = "REF",
  other_allele_col = "ALT",
  pval_col = "P",
  units_col = "units",
  ncase_col = "N_CASES",
  ncontrol_col = "N_CTRLS",
  samplesize_col = "N",
  gene_col = "gene",
  min_pval = 1e-200,
  z_col = "Z",
  info_col = "info",
  chr_col = "CHROM",
  pos_col = "POS",
  log_pval = FALSE
)  %>%
  as_tibble()


Format_mtDNAcn_chong <- Format_mtDNAcn_chong %>%
mutate(exposure = "mtDNA_chong")


# longchamps Filter and Clumping
options(ieugwasr_api = 'gwas-api.mrcieu.ac.uk/')

mtDNAcn_chong_gws <- filter(Format_mtDNAcn_chong, pval.exposure < 5e-8)


mtDNAcn_clump_chong <- clump_data(mtDNAcn_chong_gws, clump_kb = 10000,
                            clump_r2 = 0.001,
                            clump_p1 = 1,
                            clump_p2 = 1,
                            pop = "EUR"
)%>% 
  as_tibble()

#Remove APOE region
mtDNAcn_clump_chong <- filter(mtDNAcn_clump_chong, !(chr.exposure == 19 & between(pos.exposure, 44912079, 45912079)))
```

## Exposure 2: Blood Cell Composition 

This will be adjusting for differences in relative mtDNA count number in different
cell types in our exposures.

```{r exposure2}

exposure2_path_mod = "/wynton/group/andrews/data/summary_stats/results/Astle2016platelets"

platelets <- read_tsv(exposure2_path_mod, comment = "##")

platelets_renamed <- platelets %>%
rename(CHROM = chromosome, 
        POS = hm_pos,
        REF = effect_allele,
        ALT = other_allele, 
        BETA = beta, 
        SE = standard_error,
        P = p_value, 
        AF = effect_allele_frequency,
        OR = odds_ratio, 
        OR_L95 = ci_lower, 
        OR_U95 = ci_upper,
        DBSNP_ID = hm_rsid  )
           

         
# Format data to TwoSampleMR format
platelets_formatted <- platelets_renamed %>%
  format_data(.,
    type = "exposure",
    snps = NULL,
    header = TRUE,
    snp_col = "DBSNP_ID",
    beta_col = "BETA",
    se_col = "SE",
    eaf_col = "AF",
    effect_allele_col = "ALT",
    other_allele_col = "REF",
    pval_col = "P",
    chr_col = "CHROM",
    pos_col = "POS",
    log_pval = FALSE
) %>%
  as_tibble()

platelets_formatted <- platelets_formatted %>%
mutate(exposure = "Platlet_cn")

platelets_gws <- filter(platelets_formatted, pval.exposure < 5e-8)

# Perform LD clumping on SNP data, filter SNPs to make it run faster
platelets_clump <- platelets_gws %>% 
  filter(pval.exposure < 0.01) %>%
  clump_data(.,
  clump_kb = 10000,
  clump_r2 = 0.001,
  clump_p1 = 1,
  clump_p2 = 1,
  pop = "EUR"
)


platelets_clump <- filter(platelets_clump, !(chr.exposure == 19 & between(pos.exposure, 44912079, 45912079)))

# Export Clumped and Filtered dataset.

#write_csv(exposure2_dat, 'Resources/platelets_clumped_dataset.csv')

#Read in exposure2_dat 

#exposure2_dat <- read_csv("/Users/achatterjee/gitcode/Brian-Estimating-Causality-mtDNAcn-on-Alzheimer/Resources/platelets_clumped_dataset.csv")

```

## Outcome: Kunkle
``` {r read_in_Kunkle}
outcome_path_kunkle = "/wynton/group/andrews/data/summary_stats/results/Kunkle2019load_stage123.chrall.CPRA_b37.tsv.gz"

# Import Code for outcome GWAS 
OutcomeAD_kunkle <- read_tsv(outcome_path_kunkle,comment="#")

# Rename right-side variables to match column names of your GWAS dataset 

outcome_AD_Format_kunkle <- OutcomeAD_kunkle %>%
  format_data(.,
    type = "outcome",
    snps = NULL,
    header = TRUE,
    phenotype_col = "Phenotype",
    snp_col = "DBSNP_ID",
    beta_col = "BETA",
    se_col = "SE",
    eaf_col = "AF",
    effect_allele_col = "ALT",
    other_allele_col = "REF",
    pval_col = "P",
    samplesize_col = "N",
    z_col = "Z",
    chr_col = "CHROM",
    pos_col = "POS",
    log_pval = FALSE
) %>%
  as_tibble()

outcome_AD_Format_kunkle <- outcome_AD_Format_kunkle %>%
mutate(outcome = "kunkle")
```

##Combine exposures and extract from exposure datasets
```{r}
## Make list of distinct SNPs from the exposures independent GWS snps 
mvmr_snps <- bind_rows(
  select(mtDNAcn_clump_chong, SNP), 
  select(platelets_clump, SNP), 
) %>%
  distinct(SNP) %>%
  pull(SNP)

## extract combined SNP list from exposure datasets 
comb_exp <- bind_rows(
  Format_mtDNAcn_chong %>% filter(SNP %in% mvmr_snps),
  platelets_formatted %>% filter(SNP %in% mvmr_snps),
)
```


##Combine the exposure SNP lists and extract all the SNPs from each exposure
```{r}
exp_snp_list <- full_join(
  select(mtDNAcn_clump_chong, SNP, exposure), 
  select(platelets_clump, SNP, exposure), 
  by = 'SNP'
)

exp1_unq_snps <- filter(exp_snp_list, !is.na(exposure.x) & is.na(exposure.y)) %>% nrow()
exp2_unq_snps <- filter(exp_snp_list, !is.na(exposure.y) & is.na(exposure.x)) %>% nrow()
both_unq_snps <- filter(exp_snp_list, !is.na(exposure.y) & !is.na(exposure.x)) %>% nrow()


# Combined exposures SNPs

## exposure and outcome snps lists 
comb_exp_out <- bind_rows(
  Format_mtDNAcn_chong %>% filter(SNP %in% mvmr_snps) %>% 
    rename(pval = pval.exposure, trait = exposure, chr = chr.exposure, pos = pos.exposure), 
  platelets_formatted %>% filter(SNP %in% mvmr_snps) %>% 
    rename(pval = pval.exposure, trait = exposure, chr = chr.exposure, pos = pos.exposure), 
  outcome_AD_Format_kunkle %>% filter(SNP %in% mvmr_snps) %>% 
    rename(pval = pval.outcome, trait = outcome, chr = chr.outcome, pos = pos.outcome)
)

## How many SNPs are GWS across the exposures 
comb_exp_wide <- comb_exp_out %>%
  arrange(chr, pos) %>%
  select(SNP, trait, pval) %>%
  pivot_wider(names_from = trait, values_from = pval) %>%
  mutate(pval.exposure = pmin(mtDNA_chong, Platlet_cn, na.rm = T))

comb_exp_wide %>% count(mtDNA_chong < 5e-8, Platlet_cn < 5e-8)  %>%
  gt() 
```


##Proxy SNPs for chong
```{r identify proxy SNPs and munge}

# extract platelet SNPs present in chong
exposure_clump_chong <- semi_join(
 Format_mtDNAcn_chong, platelets_clump, by = "SNP"
)

# Platelet SNPs not present in chong
exp_snps_wo <- anti_join(
  platelets_clump, Format_mtDNAcn_chong, by = "SNP"
) #0 SNPs

# Use LDLinkR to identify proxy snps
# LDproxy_batch(exp_snps_wo$SNP, 
#         pop = "CEU", 
#         r2d = "r2", 
#         token = '7dbc1afe7454', 
#         append = TRUE,
#         genome_build = "grch37")
# system("mv combined_query_snp_list_grch37.txt exposure_outcome_proxy_snps_longchamps_kunkle_plate.txt")
# 
# # Munge proxy snp file
# longchamps_proxy <- mvmr_munge_proxies("exposure_outcome_proxy_snps_longchamps_kunkle_plate.txt", Format_mtDNAcn_longchamps) 
```


##Proxy SNPs for platelets
```{r identify proxy SNPs and munge}

# extract chong SNPs present in platelets
exposure_clump_platelets <- semi_join(
 platelets_formatted, mtDNAcn_clump_chong, by = "SNP"
)

# Chong SNPs not present in platelets
exp_snps_wo <- anti_join(
  mtDNAcn_clump_chong, platelets_formatted, by = "SNP"
) #0 SNPs

# Use LDLinkR to identify proxy snps
# LDproxy_batch(exp_snps_wo$SNP, 
#         pop = "CEU", 
#         r2d = "r2", 
#         token = '7dbc1afe7454', 
#         append = TRUE,
#         genome_build = "grch37")
# system("mv combined_query_snp_list_grch37.txt exposure_outcome_proxy_snps_longchamps_kunkle_plate.txt")
# 
# # Munge proxy snp file
# longchamps_proxy <- mvmr_munge_proxies("exposure_outcome_proxy_snps_longchamps_kunkle_plate.txt", Format_mtDNAcn_longchamps) 
```

#Clump combined exposures
```{r}
clump_comb_exp <- comb_exp_wide %>% 
  clump_data(.,
  clump_kb = 10000,
  clump_r2 = 0.001,
  clump_p1 = 1,
  clump_p2 = 1,
  pop = "EUR"
)

comb_exp_independent <- comb_exp %>% filter(SNP %in% clump_comb_exp$SNP) 
```

#Extract exposure SNPs from outcome
```{r}
# extract exposure SNPs present in outcome
outcome_clump <- semi_join(
  outcome_AD_Format_kunkle, comb_exp_independent, by = "SNP"
)

# Exposure SNPs not present in outomce
outcome_wo_snps <- anti_join(
  comb_exp_independent, outcome_AD_Format_kunkle, by = "SNP"
) %>%
  distinct(SNP) #13 SNPs
```

#LD proxy SNP for outcome
```{r}
# Use LDLinkR to identify proxy snps
LDproxy_batch(outcome_wo_snps$SNP, 
        pop = "CEU", 
        r2d = "r2", 
        token = 'a6deee62cc4a', 
        append = TRUE,
        genome_build = "grch37")
system("mv combined_query_snp_list_grch37.txt exposure_outcome_proxy_snps_chong_plate_kunkle.txt")

# Munge proxy snp file
outcome_dat <- munge_proxies("exposure_outcome_proxy_snps_chong_plate_kunkle.txt", outcome_AD_Format_kunkle, outcome_clump)
```

#Harmonize
```{r}
# Univariate MR 
mrdat <- harmonise_data(comb_exp_independent, outcome_dat) %>%
  as_tibble() %>%
  filter(pval.exposure < 5e-8)

## MVMR
mvdat <- mv_harmonise_data(comb_exp_independent, outcome_dat)

```

#export 
```{r}
write_rds(mvdat, 'harmonized_chong_plate_kunkle_mvmr.rds')
```

#Read in mvdat
```{r}
mvdat <- read_rds('harmonized_longchamps_plate_kunkle_mvmr.rds')

mvmr_res <- mv_multiple(mvdat, plots = F)

```


#Using MVMR package
```{r}
mvmr_dat <- cbind(mvdat$exposure_beta,
                  mvdat$exposure_se,
                  mvdat$outcome_beta,
                  mvdat$outcome_se
                 ) %>%
  as.data.frame() 


# Rename columns
colnames(mvmr_dat) <- c("IiFFho_beta", "oEdjcb_beta", "IiFFho_se", "oEdjcb_se", "outcome_beta", "outcome_se")

#Continue with mvmr_dat
mvmr_dat <- mvmr_dat %>% rownames_to_column() %>% 
      as.matrix()

F.data <- format_mvmr(BXGs = mvmr_dat[,c(2,3)],
                    BYG = mvmr_dat[,6],
                    seBXGs = mvmr_dat[,c(4,5)],
                    seBYG = mvmr_dat[,7],
                    RSID = mvmr_dat[,1])

res_mvmr <- ivw_mvmr(r_input = F.data)

mvmr_out <- res_mvmr %>% 
    as.data.frame() %>%
    rownames_to_column() %>%
    as_tibble() %>% 
    janitor::clean_names() %>%
    rename(exposure = rowname, b = estimate, se = std_error, p = pr_t) %>%
    mutate(
      exposure = ifelse(exposure == 'exposure1', 'mtDNA_chong', 'Platlet_cn'), 
      outcome = 'AD',
      method = 'IVW-MVMR'
    ) %>%
    relocate(outcome, method, .after = exposure)
```

