---
title: "MR Mock project mtDNA-CN Markdown Chong"
author: "Brian Alvarez"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

knitr::opts_knit$set(root.dir = 'C:/Users/brian/OneDrive/Documents/GitHub/Brian-Mock-MR-Project-Chong')
```


## Libraries

Loads the appropriate packages for MR analysis

``` {r eval=FALSE}
library(tidyverse) # Data wrangling 

library (usethis)

library(devtools)

library(ggman)

library(TwoSampleMR) # MR 

library(R.utils) #To read in Gz data? 
#For the read_outcome_data step

library(RadialMR) # Radial MR sensitivity analysis 

library(phenoscanner)

library(LDlinkR) # LD and proxy snps

library(lhcMR)

library(janitor)
```


## Importing and formatting Exposure

This step imports our GWAS Exposure data sets and formats them up for MR analysis


```{r eval=FALSE}
# Define column types for summary statistics GWAS
coltypes = cols(
  ID = col_character(),
  CHROM = col_double(),
  POS = col_double(),
  REF = col_character(),
  ALT = col_character(),
  AF = col_double(),
  TRAIT = col_character(),
  BETA = col_double(),
  SE = col_double(),
  Z = col_double(),
  P = col_double(),
  N = col_double(),
  OR = col_double(),
  OR_L95 = col_double(),
  OR_U95 = col_double(),
  DIR = col_character(),
  G1000_ID = col_character(),
  G1000_VARIANT = col_character(),
  DBSNP_ID = col_character(),
  DBSNP_VARIANT = col_character(),
  OLD_ID = col_character(),
  OLD_VARIANT = col_character()
)


exposure_path = "Resources/Chong2022mtdnacn_eur.chrall.CPRA_b37.tsv.gz"

Exposure_mtDNAcn <- read_tsv(exposure_path,comment="#", guess_max=19400433)


# To format the GWAS results for MR analysis Exposure. 
# Rename right-side variables to match names of your GWAS dataset columns
Format_mtDNAcn_chong <- Exposure_mtDNAcn %>% format_data(.,
  type = "exposure",
  snps = NULL,
  header = TRUE,
  phenotype_col = "TRAIT",
  snp_col = "DBSNP_ID",
  beta_col = "BETA",
  se_col = "SE",
  eaf_col = "AF",
  effect_allele_col = "REF",
  other_allele_col = "ALT",
  pval_col = "P",
  units_col = "units",
  ncase_col = "N_CASES",
  ncontrol_col = "N_CTRLS",
  samplesize_col = "N",
  gene_col = "gene",
  min_pval = 1e-200,
  z_col = "Z",
  info_col = "info",
  chr_col = "CHROM",
  pos_col = "POS",
  log_pval = FALSE
)  %>%
  as_tibble()


```

## Importing and formatting Outcome

This step imports our GWAS Outcome data sets and formats them up for MR analysis

``` {r eval=FALSE}

outcome_path = "Resources/Kunkle2019load_stage123.chrall.CPRA_b37.tsv.gz"

# Import Code for outcome GWAS 
OutcomeAD <- read_tsv(outcome_path,comment="#",guess_max=10528608)

# Rename right-side variables to match names of your GWAS dataset columns

outcome_AD_Format_Chong <- OutcomeAD %>%
  format_data(.,
    type = "outcome",
    snps = NULL,
    header = TRUE,
    phenotype_col = "Phenotype",
    snp_col = "DBSNP_ID",
    beta_col = "BETA",
    se_col = "SE",
    eaf_col = "AF",
    effect_allele_col = "ALT",
    other_allele_col = "REF",
    pval_col = "P",
    samplesize_col = "N",
    z_col = "Z",
    chr_col = "CHROM",
    pos_col = "POS",
    log_pval = FALSE
) %>%
  as_tibble()

# Above dataset to be used in LD Proxy step

```

## Filter Exposure data
This allows for more time efficient clumping.

``` {r eval=FALSE}



# The filter statements makes our dataset small in 
#reference to a pval.
#We then use the clump_data command.
#Extract Instrumental variables from the exposure.
# Note that the lower R^2 is, the less linkage disequilibrium snps will
#be included. Use clump_data. Filter will take all
#Observations below p value of 1e-05, change to
#8e-08
# This filters all of the significant SNPS and leaves all non significant 
# SNPs from the analysis.

mtDNAcn_Format_chong <- filter(Format_mtDNAcn_chong, pval.exposure<1e-05)

```

## Clumping and Harmonization
This step involving clumping and pruning to eliminate High linkage 
disequilibrium SNP (SNPS correlated with other SNPs) and harmonizes SNPS in the 
exposure and the outcome GWAS datasets. 

``` {r eval=FALSE}

#Extract Instrumental variables from the exposure.
# Note that the lower Clump
#Clump R^2 is, the less linkage disequilibrium snps will
#be included. Use clump_data instead.
# Filter allows for more time efficient clumping

mtDNAcn_Clump_chong <- clump_data(mtDNAcn_Format_chong, clump_kb = 10000,
                            clump_r2 = 0.001,
                            clump_p1 = 1,
                            clump_p2 = 1,
                            pop = "EUR"
)%>% 
  as_tibble()

# The filter statements makes our dataset small in 
#reference to a pval.
#We then use the clump_data command.
#Extract Instrumental variables from the exposure.
# Note that the lower R^2 is, the less linkage disequilibrium snps will
#be included. Use clump_data. Filter will take all
#Observations below p value of 1e-05, change to
#8e-08
# This filters all of the significant SNPS and leaves all non significant 
# SNPs from the analysis.




# Filter Clumped SNPS to include only those SNPS whose P value is 
# Below 5e-8, Colname for P value is Pval.exposure in the filtered dataset.

mtDNAcn_Clump_chong <- filter(mtDNAcn_Clump_chong , pval.exposure < 5e-8)

#For outcome data I believe you use this code. 
#It places the SNPS
#From our clumped and pruned dataset, and 
#references the outcome
#dataset to pull out those same SNPS.
#Makes sure to write in quotations the columns names for
#,"RDatasetsb/Kunkle2019load_stage123.chrall.CPRA_b37.tsv.gz"
#Put above name in file name statement.
#Put \t in "sep" statement.
#Change name from Cholesadcombine to Outcome_AD_2

# Added code for Phenotype name="TRAIT" 
# phenotype_col = "TRAIT"
# Did not work :( Must be missing something!

# Data set to be used in LD Proxy step
mtDNAcn_Clump_chong_data <- mtDNAcn_Clump_chong 
```

## LD Proxies
This attempts to retain any missing SNP's that are not both present in the 
exposure and outcome dataset by utilizing LD proxies in their place.

``` {r eval=FALSE}

# extract exposure SNPs present in outcome
outcome_clump <- semi_join(
  outcome_AD_Format_Chong, mtDNAcn_Clump_chong_data, by = "SNP"
)

# Exposure SNPs not present in outcome
exp_snps_wo <- anti_join(
  mtDNAcn_Clump_chong_data, outcome_AD_Format_Chong, by = "SNP"
)


# Use LDLinkR to identify proxy snps based on SNP dataset above
# that presents SNPs with no corresponding match in the other.
LDproxy_batch(exp_snps_wo$SNP, 
        pop = "CEU", 
        r2d = "r2", 
        token = 'a6deee62cc4a', 
        append = TRUE,
        genome_build = "grch37")
system("mv combined_query_snp_list_grch37.txt exposure_outcome_proxy_snps.txt")


# Munge proxy snp file
proxy_snps <- read_tsv("exposure_outcome_proxy_snps.txt", skip = 1, col_names = F) %>%
  rename(id = X1, func = X2, proxy_snp = X3, coord = X4, alleles = X5, maf = X6, 
         distance = X7, dprime = X8, rsq = X9, correlated_alleles = X10, FORGEdb = X11, RegulomeDB = X12) %>%
  separate(coord, c('chr', 'pos'), sep = ":") %>%
  mutate(snp = ifelse(id == 1, proxy_snp, NA), 
         chr = str_replace(chr, 'chr', ""), 
         chr = as.numeric(chr), 
         pos = as.numeric(pos)) %>%
  fill(snp, .direction = 'down') %>%
  relocate(snp, .before = proxy_snp) %>%
  dplyr::select(-id, -func, -FORGEdb, -RegulomeDB) %>%
  filter(rsq >= 0.8)

# Munge proxy snp and outcome data (Changed Outcome, to outcome_AD_Format_Chong)
proxy_outcome <- left_join(
  proxy_snps, outcome_AD_Format_Chong, by = c("proxy_snp" = "SNP")
) %>%
  separate(correlated_alleles, c("target_a1.outcome", "proxy_a1.outcome", 
                                 "target_a2.outcome", "proxy_a2.outcome"), sep = ",|=") %>%
  filter(!is.na(chr.outcome)) %>%
  arrange(snp, -rsq, abs(distance)) %>%
  group_by(snp) %>%
  slice(1) %>%
  ungroup() %>%
  mutate(
     proxy.outcome = TRUE,
     target_snp.outcome = snp,
     proxy_snp.outcome = proxy_snp, 
  ) %>% 
  mutate(
       new_effect_allele.outcome = case_when(
        proxy_a1.outcome == effect_allele.outcome & proxy_a2.outcome == other_allele.outcome ~ target_a1.outcome,
        proxy_a2.outcome == effect_allele.outcome & proxy_a1.outcome == other_allele.outcome ~ target_a2.outcome,
        TRUE ~ NA_character_
     ), 
      new_other_allele.outcome = case_when(
        proxy_a1.outcome == effect_allele.outcome & proxy_a2.outcome == other_allele.outcome ~ target_a2.outcome,
        proxy_a2.outcome == effect_allele.outcome & proxy_a1.outcome == other_allele.outcome ~ target_a1.outcome,
        TRUE ~ NA_character_
     ), 
     effect_allele.outcome = new_effect_allele.outcome, 
     other_allele.outcome = new_other_allele.outcome
  ) %>%
  dplyr::select(-proxy_snp, -chr, -pos, -alleles, -maf, -distance, -rsq, -dprime,  
         -new_effect_allele.outcome, -new_other_allele.outcome) %>%
  relocate(target_a1.outcome, proxy_a1.outcome, target_a2.outcome, proxy_a2.outcome, .after = proxy_snp.outcome) %>%
  rename(SNP = snp) %>%
  relocate(SNP, .after = samplesize.outcome)

# Merge outcome and proxy outcomes
outcome_dat_proxy <- bind_rows(
  outcome_clump, proxy_outcome
) %>% 
  arrange(chr.outcome, pos.outcome)

```
## Harmonization Step

``` {r eval=FALSE}
# and now just this. Change the two datasets names to ones created
# doing LD proxies. Put proxy dataset in outcome slot
mtDNAcn_AD_Harm_chong <- harmonise_data(mtDNAcn_Clump_chong, outcome_dat_proxy,action=2)
```

```{r export_harmonized_data}
write_csv(mtDNAcn_AD_Harm_chong, 'data/harmonized_data_Chong.csv')
# mtDNAcn_AD_Harm_chong <- read_csv('data/harmonized_data_Chong.csv')
```

## Duplicate SNPS
Checks and returns a count for any duplicate SNPs that are found.

```{r eval=FALSE, include=FALSE}
duplicated(mtDNAcn_AD_Harm_chong)

sum(duplicated(mtDNAcn_AD_Harm_chong))

janitor::get_dupes(mtDNAcn_AD_Harm_chong,SNP)
```

## Preliminary MR 

We run a Initial MR analysis for at a glance analysis on the SNPS being used in
MR analysis. Potential outliers not removed from this analysis

```{r MR_analysis}

Mr_mtDNAcn_AD_chong_Prime <- mr(mtDNAcn_AD_Harm_chong, method_list = c("mr_ivw_fe", "mr_egger_regression", "mr_weighted_median", "mr_weighted_mode"))

Mr_mtDNAcn_AD_chong_Prime_OR <- generate_odds_ratios(Mr_mtDNAcn_AD_chong_Prime)

view (Mr_mtDNAcn_AD_chong_Prime_OR )



```
## Perform Preliminary Sensentivity analysis
```{r sensitivity_tests}
# Heterogeneity
Mr_H_Test_Out_mtDNAcn_Prime_chong <- mr_heterogeneity(mtDNAcn_AD_Harm_chong)

View(Mr_H_Test_Out_mtDNAcn_Prime_chong)

# Horizontal pleiotropy Test
Mr_Horizontal_Test_mtDNAcn_Prime_chong <- mr_pleiotropy_test(mtDNAcn_AD_Harm_chong)

View(Mr_Horizontal_Test_mtDNAcn_Prime_chong)

# Directional Causality Test
AD_direction_Prime_mtDNAcn_chong <- directionality_test(mtDNAcn_AD_Harm_chong)

View(AD_direction_Prime_mtDNAcn_chong)

```
## Preliminary Plots
``` {r Preliminary_plots}

MrAD_single_mtDNAcn_Prime_chong <- mr_singlesnp(mtDNAcn_AD_Harm_chong)

MrAD_loo_mtDNAcn_Prime_chong <- mr_leaveoneout(mtDNAcn_AD_Harm_chong)

MrAD_single_plot_mtDNAcn_Prime_chong <- mr_singlesnp(mtDNAcn_AD_Harm_chong)
p1_Prime_Single <- mr_forest_plot(MrAD_single_plot_mtDNAcn_Prime_chong )
p1_Prime_Single [[1]]

Funnelplot_Singlesnp_Prime_mtDNAcn_chong <- mr_singlesnp(mtDNAcn_AD_Harm_chong)
P2_Prime_Funnel_single <- mr_funnel_plot(Funnelplot_Singlesnp_Prime_mtDNAcn_chong)
P2_Prime_Funnel_single[[1]]

MrAD_loo_plot_Prime_mtDNAcn_chong <- mr_leaveoneout(mtDNAcn_AD_Harm_chong)
p3_Prime_loo <- mr_leaveoneout_plot(MrAD_loo_plot_Prime_mtDNAcn_chong)
p3_Prime_loo[[1]]

# To plot preliminary MR analysis, must rerun and rename.

Mr_ADCholes_Plot_mtDNAcn_chong <- mr(mtDNAcn_AD_Harm_chong)

p4_Prime_scatter_chong <- mr_scatter_plot(Mr_ADCholes_Plot_mtDNAcn_chong, mtDNAcn_AD_Harm_chong)

p4_scatter_out_prime <- p4_Prime_scatter_chong[[1]] + theme_bw() + 
  guides(color=guide_legend(ncol =1)) + 
  theme(
    text = element_text(size = 8), 
  )
p4_scatter_out_prime

```
## Export preliminary MR Results

```{r preliminary_results}
# MR Analysis
write_csv(Mr_mtDNAcn_AD_chong_Prime_OR , 'results/MR_Results_Chong/MR_Prime_Chong.csv')

# Sensitivity Tests
write_csv(Mr_mtDNAcn_AD_chong_Prime_OR , 'results/MR_Results_Chong/MR_Prime_Chong.csv')
write_csv(Mr_H_Test_Out_mtDNAcn_Prime_chong , 'results/Sensitivity_Tests_Chong/mr_het_res_Chong_Prime.csv')
write_csv(Mr_Horizontal_Test_mtDNAcn_Prime_chong , 'results/Sensitivity_Tests_Chong/mr_plt_res_Chong_Prime.csv')
write_csv(AD_direction_Prime_mtDNAcn_chong , 'results/Sensitivity_Tests_Chong/Directional_Chong_Prime.csv')

# Plots

ggsave('results/Plots_Chong/mr_single_Forest_Chong_Prime.png', plot = p1_Prime_Single [[1]], units = 'in', height = 4, width = 9)
ggsave('results/Plots_Chong/mr_funnel_Single_Chong_Prime.png', plot = P2_Prime_Funnel_single[[1]], units = 'in', height = 4, width = 9)
ggsave('results/Plots_Chong/mr_forest_loo_Chong_Prime.png', plot =p3_Prime_loo[[1]] + theme(text = element_text(size = 8)), units = 'in', height = 9, width = 9)
ggsave('results/Plots_Chong/mr_scatter_Chong_Prime.png', plot = p4_scatter_out_prime, units = 'in', height = 4, width = 9)
```
## Radial MR to Remove Outliers

Radial MR will be ran first to Remove the presence of outliers and APOE SNPS.

``` {r}
#  Correction for Multiple testing bonferroni correction .05/nrow

AD_Radial_egg_mtDNAcn_chong <-  egger_radial(mtDNAcn_AD_Harm_chong, .05/nrow(mtDNAcn_AD_Harm_chong), 2, TRUE)


AD_Radial_Ivw_mtDNAcn_chong <- ivw_radial(mtDNAcn_AD_Harm_chong, .05/nrow(mtDNAcn_AD_Harm_chong), 2, .0001, TRUE)

AD_Radial_egg_mtDNAcn_chong

#Displays a Interactive radial plot of SNPS and which SNPS are outliers.

Chong_AD_Radial_egg <- plotly_radial(AD_Radial_egg_mtDNAcn_chong, TRUE)

Chong_AD_Radial_Ivw <- plotly_radial(AD_Radial_Ivw_mtDNAcn_chong , TRUE)

Chong_AD_Radial_egg

# Static plot results of Radial MR

radial_p_Chong_IVW <- plot_radial(AD_Radial_Ivw_mtDNAcn_chong, radial_scale = F, show_outliers = T)

radial_p_Chong_IVW

```
## Export Radial Plots

``` {r radial_outliers_export}

ggsave('results/Plots_Chong/mr_radial_IVW.png', plot = radial_p_Chong_IVW, units = 'in', height = 4, width = 9)

```

## Outlier details
``` {r}

# To receive details on outliers detected in radial MR
# Tutorial, after phenoscanner (snpquery=datasetname$outliers$SNP)
# Checks your dataset and queries the outlier variable as well as SNP
# Will return Error if no outliers are present.

Outlier_Details_mtDNAcn_chong <- phenoscanner(snpquery=AD_Radial_egg_mtDNAcn_chong[[outliers]][[SNP]])

```

## Combining Radial data and harmonized
``` {r}
# This step creates a new dataset combining our radial data with our harmonized
# data, creating a dichotomous outlier and Apoe category
# (Columns), so when conducting  MR on this dataset, omission of outliers and 
# APOE SNPS that may violate MR is possible.

mtDNAcn_AD_Harm_Outlier_chong <- mtDNAcn_AD_Harm_chong %>%
  left_join(AD_Radial_Ivw_mtDNAcn_chong$data) %>%
  mutate(
    apoe_region = case_when(
      chr.outcome == 19 & between(pos.outcome,44912079,45912079) ~ TRUE,
      TRUE ~ FALSE
    ),
    gws.outcome = ifelse(pval.outcome <5e-8, TRUE, FALSE),
    mr_keep = ifelse(mr_keep == FALSE | Outliers == "Outlier" | apoe_region == TRUE | gws.outcome == TRUE, FALSE, TRUE)
  )
```

## MR analysis on our data tidyed data set.

Our data set is now clean, now we rerun Mr analysis with these outliers removed and LD Proxies added in.

``` {r}
#Run MR analysis on new data-set using this code.

Mr_mtDNAcn_AD_Outlier_chong <- mr(mtDNAcn_AD_Harm_Outlier_chong, method_list = c("mr_ivw_fe", "mr_egger_regression", "mr_weighted_median", "mr_weighted_mode"))

Mr_mtDNAcn_AD_oratio_Outlier_chong <- generate_odds_ratios(Mr_mtDNAcn_AD_Outlier_chong)

view (Mr_mtDNAcn_AD_oratio_Outlier_chong)

```

## Perform Outlier Removed Sensentivity analysis

``` {r sensitivity}
# Heterogeneity (Test for Overall Violations of MR assumptions)
Mr_H_Test_Out_mtDNAcn_outlier_chong <- mr_heterogeneity(mtDNAcn_AD_Harm_Outlier_chong)

View(Mr_H_Test_Out_mtDNAcn_outlier_chong)

# (Test for Violations of MR assumptions via Horizontal pleiotrophy)
Mr_Horizontal_Test_mtDNAcn_outlier_chong <- mr_pleiotropy_test(mtDNAcn_AD_Harm_Outlier_chong)

View(Mr_Horizontal_Test_mtDNAcn_outlier_chong)

# Directionality test (testing whether Exposure >Outcome direction is true casual direction)

AD_direction_Outlier_mtDNAcn_chong <- directionality_test(mtDNAcn_AD_Harm_Outlier_chong)

View(AD_direction_Outlier_mtDNAcn_chong)

```

## MR Outlier Plots

``` {r}
MrAD_single_mtDNAcn_outlier_chong <- mr_singlesnp(mtDNAcn_AD_Harm_Outlier_chong)

MrAD_loo_mtDNAcn_outlier_chong <- mr_leaveoneout(mtDNAcn_AD_Harm_Outlier_chong)

MrAD_single_plot_mtDNAcn_outlier_chong <- mr_singlesnp(mtDNAcn_AD_Harm_Outlier_chong)
p5_Out_Removed_Forest_Single_Chong <- mr_forest_plot(MrAD_single_plot_mtDNAcn_outlier_chong)
p5_Out_Removed_Forest_Single_Chong[[1]]



Funnelplot_Singlesnp_Outlier_mtDNAcn_chong <- mr_singlesnp(mtDNAcn_AD_Harm_Outlier_chong)
p6_Out_Removed_Funnel_Single_Chong <- mr_funnel_plot(Funnelplot_Singlesnp_Outlier_mtDNAcn_chong)
p6_Out_Removed_Funnel_Single_Chong[[1]]

#This plots all of the leave one out analysis.

MrAD_loo_plot_Out_mtDNAcn_chong <- mr_leaveoneout(mtDNAcn_AD_Harm_Outlier_chong)
p7_Out_Removed_Forest_loo_Chong <- mr_leaveoneout_plot(MrAD_loo_plot_Out_mtDNAcn_chong)
p7_Out_Removed_Forest_loo_Chong[[1]]



# All MR analysis with outliers removed by plot
# To plot preliminary MR analysis, must rerun and rename.

Mr_AD_Plot_Out_mtDNAcn_chong <- mr(mtDNAcn_AD_Harm_Outlier_chong)

p8_Out_Removed_Scatter_Chong <- mr_scatter_plot(Mr_AD_Plot_Out_mtDNAcn_chong, mtDNAcn_AD_Harm_Outlier_chong)


p8_scatter_out_Chong <- p8_Out_Removed_Scatter_Chong[[1]] + theme_bw() + 
  guides(color=guide_legend(ncol =1)) + 
  theme(
    text = element_text(size = 8), 
  )
p8_scatter_out_Chong


```

## Export Code to Save Results 
```{r export_results_outliers}
# MR Analysis
write_csv(Mr_mtDNAcn_AD_Outlier_chong , 'results/MR_Results_Chong/MR_Out_Chong.csv')

# Sensitivity Tests

write_csv(Mr_H_Test_Out_mtDNAcn_outlier_chong, 'results/Sensitivity_Tests_Chong/mr_het_res_Chong_Out.csv')
write_csv(Mr_Horizontal_Test_mtDNAcn_outlier_chong , 'results/Sensitivity_Tests_Chong/mr_plt_res_Chong_Out.csv')
write_csv(AD_direction_Outlier_mtDNAcn_chong , 'results/Sensitivity_Tests_Chong/Directional_Chong_Out.csv')


# MR Outlier removed Plots

ggsave('results/Plots_Chong/mr_single_Forest_Chong_Out.png', plot = p5_Out_Removed_Forest_Single_Chong [[1]], units = 'in', height = 4, width = 9)
ggsave('results/Plots_Chong/mr_funnel_Single_Chong_Out.png', plot = p6_Out_Removed_Funnel_Single_Chong[[1]], units = 'in', height = 4, width = 9)
ggsave('results/Plots_Chong/mr_forest_loo_Chong_Out.png', plot =p7_Out_Removed_Forest_loo_Chong[[1]] + theme(text = element_text(size = 8)), units = 'in', height = 9, width = 9)
ggsave('results/Plots_Chong/mr_scatter_Chong_Out.png', plot = p8_scatter_out_Chong, units = 'in', height = 4, width = 9)

```

```{r export_outlier_results} 

```


``` {r render}

rmarkdown::render()

```