---
title: "R Notebook"
output: html_notebook
editor_options: 
  chunk_output_type: console
---

# Libraries
```{r}
 library(dplyr)
library(readr)
library(broom)
library(ggplot2)
library(tidyr)
```

#CHONG mtdnacn

## Data Preparation
```{r}
fam_file <- "/wynton/group/andrews/users/rakshyasharma/PRS/PRS-CS-Auto/resources/ADGC/adgc.fam"
covar_file <- "/wynton/group/andrews/data/adgc/ADGCdatasets/covar_covariates.tsv"
base_dir <- "/wynton/group/andrews/users/rakshyasharma/pgsc_calc/adgc/results/Aadrita/Chong2022mtdnacn_eur/adgc"

pca_file <- file.path(base_dir, "score", "adgc_popsimilarity.txt.gz")
prs_file <- file.path(base_dir, "score", "adgc_pgs.txt.gz")
```

## Read in the data
```{r}
fam_data <- read.table(fam_file, header = FALSE,
                       col.names = c("FID", "IID", "PID", "MID", "Sex", "P"))
covar_data <- read.table(covar_file, header = TRUE, sep = "\t")
pca_data <- read_tsv(pca_file, show_col_types = FALSE)
prs_data <- read.table(prs_file, header = TRUE)

# Clean and prep
fam_data <- fam_data %>% distinct(IID, .keep_all = TRUE)
covar_data <- covar_data %>% distinct(ID_2, .keep_all = TRUE)
pca_data <- pca_data %>% distinct(IID, .keep_all = TRUE)
prs_data <- prs_data %>% distinct(IID, .keep_all = TRUE)

fam_data$IID <- sub("^0_", "", fam_data$IID)
fam_data$IID <- ifelse(grepl("^[0]+[0-9]+$", fam_data$IID),
                       sub("^0+", "", fam_data$IID),
                       fam_data$IID)

covar_data <- covar_data %>% mutate(ID_2 = as.character(ID_2))
pca_data <- pca_data %>% mutate(IID = as.character(IID))
prs_data <- prs_data %>% mutate(IID = as.character(IID))

# Merge data
merged_data <- covar_data %>%
  left_join(fam_data, by = c("ID_2" = "IID")) %>%
  left_join(prs_data, by = c("ID_2" = "IID")) %>%
  left_join(pca_data, by = c("ID_2" = "IID")) %>%
  distinct() %>%
  mutate(
    across(where(is.numeric), ~na_if(., -9)),
    status_binary = if_else(status == 2, 1, 0),  # AD = 1, Control = 0
    status_binary = factor(status_binary, levels = c(0,1))
  ) %>%
  filter(status %in% c(1,2)) %>%
  drop_na(MostSimilarPop, aaoaae, apoe4any, sex, status) 

merged_data$Sex <- factor(merged_data$Sex, levels = c(1, 2), labels = c("Male", "Female"))
```

## Logistic regression: mtDNAcn PRS -> AD
```{r}
model_chong <- glm(status_binary ~ Z_norm2 + aaoaae + Sex + apoe4any + PC1 + PC2 + PC3 + PC4,
             data = merged_data, family = "binomial")
summary(model_chong)
```

## Extract info
```{r}
# Extract model coefficients
coefs_chong <- summary(model_chong)$coefficients

# Define a function to compute OR and 95% CI
compute_OR_CI <- function(df) {
  df %>%
    mutate(
      OR = exp(Estimate),
      CI_lower = exp(Estimate - 1.96 * `Std. Error`),
      CI_upper = exp(Estimate + 1.96 * `Std. Error`)
    )
}

# Convert coefs_chong to a dataframe first
coefs_chong_df <- as.data.frame(coefs_chong)

# Create the dataframe coefs_chong with the odds ratio and 95% CI
coefs_chong_with_OR_CI <- coefs_chong_df %>%
  rownames_to_column(var = "term") %>%
  compute_OR_CI()

# Display the results
coefs_chong_with_OR_CI

# Filter the Z_norm2 row and select relevant columns
z_norm_info_chong <- coefs_chong_with_OR_CI %>%
  filter(term == "Z_norm2") %>%
  select(term, Estimate, `Std. Error`, `z value`, `Pr(>|z|)`, OR, CI_lower, CI_upper) %>% 
  mutate(
    Exposure = "Chong",  # Exposure is Chong mtDNAcn
    Outcome = "AD"              # Outcome is AD
  )

# Display the filtered data
z_norm_info_chong
```



#LONGCHAMPS mtdnacn

## Data Preparation
```{r}
fam_file <- "/wynton/group/andrews/users/rakshyasharma/PRS/PRS-CS-Auto/resources/ADGC/adgc.fam"
covar_file <- "/wynton/group/andrews/data/adgc/ADGCdatasets/covar_covariates.tsv"
base_dir <- "/wynton/group/andrews/users/rakshyasharma/pgsc_calc/adgc/results/Aadrita/Longchamps2021mtdnacn/adgc"

pca_file <- file.path(base_dir, "score", "adgc_popsimilarity.txt.gz")
prs_file <- file.path(base_dir, "score", "adgc_pgs.txt.gz")
```

## Read in the data
```{r}
fam_data <- read.table(fam_file, header = FALSE,
                       col.names = c("FID", "IID", "PID", "MID", "Sex", "P"))
covar_data <- read.table(covar_file, header = TRUE, sep = "\t")
pca_data <- read_tsv(pca_file, show_col_types = FALSE)
prs_data <- read.table(prs_file, header = TRUE)

# Clean and prep
fam_data <- fam_data %>% distinct(IID, .keep_all = TRUE)
covar_data <- covar_data %>% distinct(ID_2, .keep_all = TRUE)
pca_data <- pca_data %>% distinct(IID, .keep_all = TRUE)
prs_data <- prs_data %>% distinct(IID, .keep_all = TRUE)

fam_data$IID <- sub("^0_", "", fam_data$IID)
fam_data$IID <- ifelse(grepl("^[0]+[0-9]+$", fam_data$IID),
                       sub("^0+", "", fam_data$IID),
                       fam_data$IID)

covar_data <- covar_data %>% mutate(ID_2 = as.character(ID_2))
pca_data <- pca_data %>% mutate(IID = as.character(IID))
prs_data <- prs_data %>% mutate(IID = as.character(IID))

# Merge data
merged_data <- covar_data %>%
  left_join(fam_data, by = c("ID_2" = "IID")) %>%
  left_join(prs_data, by = c("ID_2" = "IID")) %>%
  left_join(pca_data, by = c("ID_2" = "IID")) %>%
  distinct() %>%
  mutate(
    across(where(is.numeric), ~na_if(., -9)),
    status_binary = if_else(status == 2, 1, 0),  # AD = 1, Control = 0
    status_binary = factor(status_binary, levels = c(0,1))
  ) %>%
  filter(status %in% c(1,2)) %>%
  drop_na(MostSimilarPop, aaoaae, apoe4any, sex, status)

merged_data$Sex <- factor(merged_data$Sex, levels = c(1, 2), labels = c("Male", "Female"))
```

## Logistic regression: mtDNAcn PRS -> AD
```{r}
model_longchamps <- glm(status_binary ~ Z_norm2 + aaoaae + Sex + apoe4any + PC1 + PC2 + PC3 + PC4,
             data = merged_data, family = "binomial")
summary(model)
```


## Extract info
```{r}
# Extract model coefficients
coefs_longchamps <- summary(model_longchamps)$coefficients

# Define a function to compute OR and 95% CI
compute_OR_CI <- function(df) {
  df %>%
    mutate(
      OR = exp(Estimate),
      CI_lower = exp(Estimate - 1.96 * `Std. Error`),
      CI_upper = exp(Estimate + 1.96 * `Std. Error`)
    )
}

# Convert coefs_chong to a dataframe first
coefs_longchamps_df <- as.data.frame(coefs_longchamps)

# Create the dataframe coefs_chong with the odds ratio and 95% CI
coefs_longchamps_with_OR_CI <- coefs_longchamps_df %>%
  rownames_to_column(var = "term") %>%
  compute_OR_CI()


# Filter the Z_norm2 row and select relevant columns
z_norm_info_longchamps <- coefs_longchamps_with_OR_CI %>%
  filter(term == "Z_norm2") %>%
  select(term, Estimate, `Std. Error`, `z value`, `Pr(>|z|)`, OR, CI_lower, CI_upper) %>% 
  mutate(
    Exposure = "Longchamps",  # Exposure is longchamps
    Outcome = "AD"              # Outcome is AD
  )

# Display the filtered data
z_norm_info_longchamps
```

#HAGG mtdnacn

## Data Preparation
```{r}
fam_file <- "/wynton/group/andrews/users/rakshyasharma/PRS/PRS-CS-Auto/resources/ADGC/adgc.fam"
covar_file <- "/wynton/group/andrews/data/adgc/ADGCdatasets/covar_covariates.tsv"
base_dir <- "/wynton/group/andrews/users/rakshyasharma/pgsc_calc/adgc/results/Aadrita/Hagg2020mtdnacn/adgc"

pca_file <- file.path(base_dir, "score", "adgc_popsimilarity.txt.gz")
prs_file <- file.path(base_dir, "score", "adgc_pgs.txt.gz")
```

## Read in the data
```{r}
fam_data <- read.table(fam_file, header = FALSE,
                       col.names = c("FID", "IID", "PID", "MID", "Sex", "P"))
covar_data <- read.table(covar_file, header = TRUE, sep = "\t")
pca_data <- read_tsv(pca_file, show_col_types = FALSE)
prs_data <- read.table(prs_file, header = TRUE)

# Clean and prep
fam_data <- fam_data %>% distinct(IID, .keep_all = TRUE)
covar_data <- covar_data %>% distinct(ID_2, .keep_all = TRUE)
pca_data <- pca_data %>% distinct(IID, .keep_all = TRUE)
prs_data <- prs_data %>% distinct(IID, .keep_all = TRUE)

fam_data$IID <- sub("^0_", "", fam_data$IID)
fam_data$IID <- ifelse(grepl("^[0]+[0-9]+$", fam_data$IID),
                       sub("^0+", "", fam_data$IID),
                       fam_data$IID)

covar_data <- covar_data %>% mutate(ID_2 = as.character(ID_2))
pca_data <- pca_data %>% mutate(IID = as.character(IID))
prs_data <- prs_data %>% mutate(IID = as.character(IID))

# Merge data
merged_data <- covar_data %>%
  left_join(fam_data, by = c("ID_2" = "IID")) %>%
  left_join(prs_data, by = c("ID_2" = "IID")) %>%
  left_join(pca_data, by = c("ID_2" = "IID")) %>%
  distinct() %>%
  mutate(
    across(where(is.numeric), ~na_if(., -9)),
    status_binary = if_else(status == 2, 1, 0),  # AD = 1, Control = 0
    status_binary = factor(status_binary, levels = c(0,1))
  ) %>%
  filter(status %in% c(1,2)) %>%
  drop_na(MostSimilarPop, aaoaae, apoe4any, sex, status)

merged_data$Sex <- factor(merged_data$Sex, levels = c(1, 2), labels = c("Male", "Female"))

```

## Logistic regression: mtDNAcn PRS -> AD
```{r}
model_hagg <- glm(status_binary ~ Z_norm2 + aaoaae + Sex + apoe4any + PC1 + PC2 + PC3 + PC4,
             data = merged_data, family = "binomial")
summary(model_hagg)
```

## Extract info
```{r}
# Extract model coefficients
coefs_hagg <- summary(model_hagg)$coefficients

# Convert coefs_chong to a dataframe first
coefs_hagg_df <- as.data.frame(coefs_hagg)

# Create the dataframe coefs_chong with the odds ratio and 95% CI
coefs_hagg_with_OR_CI <- coefs_hagg_df %>%
  rownames_to_column(var = "term") %>%
  compute_OR_CI()


# Filter the Z_norm2 row and select relevant columns
z_norm_info_hagg <- coefs_hagg_with_OR_CI %>%
  filter(term == "Z_norm2") %>%
  select(term, Estimate, `Std. Error`, `z value`, `Pr(>|z|)`, OR, CI_lower, CI_upper) %>% 
  mutate(
    Exposure = "Hagg",  # Exposure is hagg
    Outcome = "AD"              # Outcome is AD
  )

# Display the filtered data
z_norm_info_hagg
```


#GUPTA mtdnacn

## Data Preparation
```{r}
fam_file <- "/wynton/group/andrews/users/rakshyasharma/PRS/PRS-CS-Auto/resources/ADGC/adgc.fam"
covar_file <- "/wynton/group/andrews/data/adgc/ADGCdatasets/covar_covariates.tsv"
base_dir <- "/wynton/group/andrews/users/rakshyasharma/pgsc_calc/adgc/results/Aadrita/Gupta2023mtDNAcn/adgc"

pca_file <- file.path(base_dir, "score", "adgc_popsimilarity.txt.gz")
prs_file <- file.path(base_dir, "score", "adgc_pgs.txt.gz")
```

## Read in the data
```{r}
fam_data <- read.table(fam_file, header = FALSE,
                       col.names = c("FID", "IID", "PID", "MID", "Sex", "P"))
covar_data <- read.table(covar_file, header = TRUE, sep = "\t")
pca_data <- read_tsv(pca_file, show_col_types = FALSE)
prs_data <- read.table(prs_file, header = TRUE)

# Clean and prep
fam_data <- fam_data %>% distinct(IID, .keep_all = TRUE)
covar_data <- covar_data %>% distinct(ID_2, .keep_all = TRUE)
pca_data <- pca_data %>% distinct(IID, .keep_all = TRUE)
prs_data <- prs_data %>% distinct(IID, .keep_all = TRUE)

fam_data$IID <- sub("^0_", "", fam_data$IID)
fam_data$IID <- ifelse(grepl("^[0]+[0-9]+$", fam_data$IID),
                       sub("^0+", "", fam_data$IID),
                       fam_data$IID)

covar_data <- covar_data %>% mutate(ID_2 = as.character(ID_2))
pca_data <- pca_data %>% mutate(IID = as.character(IID))
prs_data <- prs_data %>% mutate(IID = as.character(IID))

# Merge data
merged_data <- covar_data %>%
  left_join(fam_data, by = c("ID_2" = "IID")) %>%
  left_join(prs_data, by = c("ID_2" = "IID")) %>%
  left_join(pca_data, by = c("ID_2" = "IID")) %>%
  distinct() %>%
  mutate(
    across(where(is.numeric), ~na_if(., -9)),
    status_binary = if_else(status == 2, 1, 0),  # AD = 1, Control = 0
    status_binary = factor(status_binary, levels = c(0,1))
  ) %>%
  filter(status %in% c(1,2)) %>%
  drop_na(MostSimilarPop, aaoaae, apoe4any, sex, status)

merged_data$Sex <- factor(merged_data$Sex, levels = c(1, 2), labels = c("Male", "Female"))

```

## Logistic regression: mtDNAcn PRS -> AD
```{r}
model_gupta <- glm(status_binary ~ Z_norm2 + aaoaae + Sex + apoe4any + PC1 + PC2 + PC3 + PC4,
             data = merged_data, family = "binomial")
summary(model_gupta)
```

## Extract info
```{r}
# Extract model coefficients
coefs_gupta <- summary(model_gupta)$coefficients

# Convert coefs_chong to a dataframe first
coefs_gupta_df <- as.data.frame(coefs_gupta)

# Create the dataframe coefs_chong with the odds ratio and 95% CI
coefs_gupta_with_OR_CI <- coefs_gupta_df %>%
  rownames_to_column(var = "term") %>%
  compute_OR_CI()


# Filter the Z_norm2 row and select relevant columns
z_norm_info_gupta <- coefs_gupta_with_OR_CI %>%
  filter(term == "Z_norm2") %>%
  select(term, Estimate, `Std. Error`, `z value`, `Pr(>|z|)`, OR, CI_lower, CI_upper) %>% 
  mutate(
    Exposure = "Gupta",  # Exposure is gupta
    Outcome = "AD"              # Outcome is AD
  )

# Display the filtered data
z_norm_info_gupta
```


#Combine z_norm info for all mtdnacn gwass and export as a dataframe
```{r}
# Combine the dataframes
combined_z_norm_info <- bind_rows(z_norm_info_gupta, z_norm_info_chong, z_norm_info_hagg, z_norm_info_longchamps)

# Remove the 'term' column
combined_z_norm_info <- combined_z_norm_info %>%
  select(Exposure, Outcome, everything()) %>%
  select(-term)

# Display the combined dataframe 
head(combined_z_norm_info)

# Export the combined dataframe to a CSV file
write_csv(combined_z_norm_info, "/wynton/group/andrews/users/achatterjee/mtdnacn/Brian-Estimating-Causality-mtDNAcn-on-Alzheimer/prs/data/prs_results_mtdnacn_onto_AD.csv")
```

